---
title: "Pricing Decision-Support for Airbnb Hosts in Vaud"
subtitle: "Machine Learning 1 - Group Project"
author: "Pascal Ingold · Timon Reichmuth · Davide Gualdani"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    df_print: paged
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
    theme: cosmo
    fontsize: 11pt
    geometry: margin=1.5cm
    linestretch: 1.0
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, comment = NA, warning = FALSE, message = FALSE, tidy = "styler")
```

```{css, echo = FALSE}
caption {
      color: black;
      font-weight: bold;
      font-size: 1.0em;
    }
```


Last ToDos: 
**ALS LETZES styler::style_file(DATEINAME) runnen in Konsole**



# Introduction

## Business Case

We propose a data-driven pricing decision-support system for Airbnb
hosts in the Canton of Vaud, providing market-aligned price
recommendations based on listing characteristics. The tool targets both
new and existing hosts, using inputs such as room type, guest
capacity, availability, ratings, and superhost status. Six machine
learning tools are used: a linear Model(LM), a generalised linear model using poisson (GLM), a generalized linear Model using binomial (GLM), a general additive model (GAM), a support vector machine (SVM) and a Neural
Network (ANN), are applied to identify key price drivers and generate accurate
predictions. The objective is to balance revenue and occupancy, helping
hosts understand their market value. A key feature is interpretability,
enabling users to see why a certain price is recommended. From a
business perspective, the system aims to increase booking rates, reduce
price dispersion, and improve host satisfaction. The model is tailored
to the Canton of Vaud to capture regional economic effects.

## Research Questions

Several questions can be derived from the business case. These questions
are intended to help us examine pricing dynamics and facilitate the
development of a pricing strategy for Airbnb-Hosts.

-   RQ1 - First, **we need to determine which factors influence guests’
    willingness to pay**. Furthermore, the nature of this influence is,
    of course, central to our analysis.
-   RQ2 - Once these factors are identified, a second question
    immediately arises: **to what extent can these factors be used to
    predict whether the prices of a rental property are justified?**

Willingness to pay is actually the central variable here. However, it is
not easy to measure, which is why a different approach must be taken.
The approach we have chosen stems from the idea that the market
generally functions well and that landlords intuitively already align
with willingness to pay. As a result, the existing price for a rental
property can also be assumed to be the price that tenants are willing to
pay. Any deviations are offset in our subsequent calculations by the
large number of observations. Thus, on average, the actual price will
correspond to the price that tenants are willing to pay.

## Hypothesis

-   H1: When host-centered indicators improve, tenants' willingness to
    pay increases.
-   H2: When location-centered indicators improve, tenants’ willingness
    to pay increases.

Host-centered indicators are those that relate to the host’s
performance, such as ratings or Superhost status. In contrast,
location-centered indicators are those that relate to factors such as
the size of the rental property. With these hypothesis, we are able to operationalize the research questions in such a way, that we can investigate the underlying relations. 

# Data Preparation

In a first step we need to load the relevant packages. Following that we
load all the datasets that are relevant to us.

```{r Libraries, results = "hide", echo=TRUE, warning=FALSE}
library(tidyverse)
library(janitor)
library(mgcv)
library(e1071)
library(broom)
library(nnet)
library(pROC)
library(Metrics)
library(kableExtra)
library(knitr)
library(performance)
library(naniar)
library(neuralnet)
library(leaflet)
library(leaflet.extras)
library(patchwork)


theme_set(theme_minimal())

path_listings <- "./Data_Vaud_all/listings-2.csv"
d.raw_data_listings <- read.csv(path_listings, header = TRUE)
```

Cleaning the Airbnb-Listings Data. We keep only the variables needed for
our project. We reduce the amount of variables to exact 20, according to
the requirements by our lecturers.

```{r cleaning Dataset, echo=TRUE}
d.airbnb_listings <- d.raw_data_listings %>%
  clean_names()

d.listings_clean <- d.airbnb_listings %>%
  mutate(
    price = str_remove_all(price, "\\$|,"),
    price = as.numeric(price),
    host_is_superhost = ifelse(host_is_superhost %in% c("t", "TRUE", "1"), 1, 0),
    room_type = as.factor(room_type),
    property_type = as.factor(property_type),
    host_response_rate = as.numeric(sub("%", "", host_response_rate)),
    host_acceptance_rate = as.numeric(sub("%", "", host_acceptance_rate)),
    host_is_superhost = ifelse(host_is_superhost == 1, TRUE, FALSE)
  ) %>%
  filter(!is.na(price), price > 0)


d.listings_clean <- d.listings_clean %>%
  select(
    -id, -host_id, -host_listings_count, -host_total_listings_count, -listing_url, -scrape_id, -last_scraped,
    -source, -name, -description, -maximum_nights, -first_review, -last_review, -calculated_host_listings_count, -host_since, -availability_365, -estimated_revenue_l365d,
    -neighborhood_overview, -picture_url, -host_url, -host_name, -host_location,
    -host_about, -host_thumbnail_url, -host_picture_url, -host_neighbourhood,
    -neighbourhood, -neighbourhood_cleansed, -bathrooms_text, -amenities,
    -minimum_minimum_nights, -maximum_minimum_nights, -minimum_maximum_nights,
    -maximum_maximum_nights, -minimum_nights_avg_ntm, -maximum_nights_avg_ntm,
    -calendar_updated, -has_availability, -availability_30, -availability_60,
    -availability_90, -calendar_last_scraped, -number_of_reviews_ltm, -number_of_reviews_l30d,
    -availability_eoy, -number_of_reviews_ly, -review_scores_accuracy, -review_scores_cleanliness,
    -review_scores_checkin, -review_scores_communication, -review_scores_location,
    -license, -instant_bookable, -calculated_host_listings_count_entire_homes, -calculated_host_listings_count_private_rooms,
    -calculated_host_listings_count_shared_rooms, -reviews_per_month, -bedrooms, -review_scores_value
  )
```
To meet this specification, we have removed
columns from the dataset that will not be used in our analysis.
Consequently, `d.listings_clean` is our final, cleaned dataset, which we
will use to run all subsequent models.

Subsequently, we calculated several additional features to include them
as permanent columns in the dataset, ensuring they do not have to be
recomputed for each analysis. These include: the log-transformed
price, as price data is typically heavily right-skewed and thus not suitable for Gaussian Models, the price per
guest (total price divided by the number of guests accommodated), a
boolean indicator for high demand, and a binary variable identifying
whether the price exceeds the median.

```{r initial FE, Lead Pascal, echo=TRUE}
d.airbnb_listings_feat <- d.listings_clean %>%
  mutate(
    log_price = log(price),
    price_per_guest = price / pmax(accommodates, 1),
    high_demand = ifelse(estimated_occupancy_l365d > 200, 1, 0),
    price_high = as.integer(price > median(price, na.rm = TRUE))
  )
```

# Understanding the Data

To inform our modelling choices, we briefly walk through how the data is
structured e.g. with distributions, variable types, and edge cases.
Airbnb already knows this data well, so we keep the commentary short and
focus on what actually matters for the models that follow.

During our exploratory data analysis, we identified one *Private room*
listing with an unrealistically high price of 2'147'483'647 USD. We
excluded this single observation from our dataset as it is an extreme
outlier that does not reflect the true market conditions and would
negatively skew our predictive models.

```{r}
# This keeps everything except the row that is a Private room and costs more than 100000
d.airbnb_listings_feat <- d.airbnb_listings_feat %>%
  filter(!(room_type == "Private room" & price > 100000))
```

In addition to the one outlier we wanted to remove one NA specifically, as this should be a Boolean.

```{r filter NA in ProfilePic}
d.airbnb_listings_feat <- d.airbnb_listings_feat %>%
  filter(!host_has_profile_pic == "")
```

## Prices

```{r, fig.show='hold', out.width="80%", echo=FALSE}
p.price <- ggplot(d.airbnb_listings_feat, aes(x = price)) +
  geom_histogram(bins = 50, fill = "deepskyblue", color = "white", alpha = 0.7) +
  labs(title = "Distribution of Raw Price", x = "Price", y = "Count")

p.logprice <- ggplot(d.airbnb_listings_feat, aes(x = log_price)) +
  geom_histogram(bins = 50, fill = "deepskyblue", color = "white", alpha = 0.7) +
  labs(title = "Distribution of Log Price", x = "Log Price", y = "Count")

p.price | p.logprice
```

Prices are heavily right-skewed. After log-transformation the
distribution is much closer to normal. That is the form we use for the
LM, GAM and most diagnostics later.

## Capacity & Rating

```{r, fig.show='hold', out.width='80%'}
p.room <- ggplot(d.airbnb_listings_feat, aes(x = room_type, fill = room_type)) +
  geom_bar(show.legend = FALSE, fill = "deepskyblue") +
  labs(title = "Proportion of Room Types", x = "Room Type", y = "Count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

p.acc <- ggplot(d.airbnb_listings_feat, aes(x = accommodates)) +
  geom_histogram(binwidth = 1, fill = "deepskyblue", color = "white") +
  scale_x_continuous(breaks = seq(1, max(d.airbnb_listings_feat$accommodates, na.rm = TRUE), by = 1)) +
  labs(title = "Listing Capacity", x = "Accommodates (Guests)", y = "Count")



p.room + p.acc
```

Most listings are built for 2 or 4 guests. Entire homes/apartments and Private rooms dominate the inventory while hotelrooms and shared rooms are marginal.

## Room Types & Superhost Status

```{r, fig.show='hold', out.width='80%'}
p.rev <- ggplot(d.airbnb_listings_feat, aes(x = review_scores_rating)) +
  geom_histogram(bins = 30, fill = "deepskyblue", color = "white", alpha = 0.7) +
  labs(title = "Review Scores Rating Distribution", x = "Rating", y = "Count")

p.host <- ggplot(d.airbnb_listings_feat %>% filter(!is.na(host_is_superhost)), aes(x = as.factor(host_is_superhost), fill = as.factor(host_is_superhost))) +
  geom_bar(show.legend = FALSE) +
  scale_fill_manual(values = c("deepskyblue", "deepskyblue")) +
  labs(title = "Superhost Status", x = "Is Superhost?", y = "Count")

p.rev + p.host
```


Review scores cluster heavily between 4 and 5. Both hosts and guests rate each other, which probably produces a leniency bias on top of any genuine quality signal.
Meanwhile are roughly 30% of hosts  Superhosts. This is a meaningful minority, but far from the norm.

## Log Prices and rooms / neighbourhoods

```{r, fig.show='hold', out.width="80%"}
p.logprice_room <- ggplot(d.airbnb_listings_feat, aes(x = room_type, y = log_price, fill = room_type)) +
  geom_boxplot(alpha = 0.7, show.legend = FALSE, fill = "deepskyblue") +
  labs(title = "Log Price by Room Type", x = "Room Type", y = "Log Price") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

d.top_neighborhoods <- d.airbnb_listings_feat %>%
  count(neighbourhood_group_cleansed, sort = TRUE) %>%
  top_n(10, n) %>%
  pull(neighbourhood_group_cleansed)

p.logprice_neighbourhood <- ggplot(
  d.airbnb_listings_feat %>% filter(neighbourhood_group_cleansed %in% d.top_neighborhoods),
  aes(x = reorder(neighbourhood_group_cleansed, log_price, FUN = median), y = log_price)
) +
  geom_boxplot(fill = "deepskyblue", alpha = 0.7) +
  coord_flip() +
  labs(
    title = "Log Price by Top 10 Neighbourhood Groups",
    x = "Neighbourhood Group",
    y = "Log Price"
  )

p.logprice_room | p.logprice_neighbourhood
```

Among room types, hotel rooms post the highest median with a tight
distribution. Likely a sign of more professional pricing. Entire homes
and private rooms show the same shape but a long upper tail (luxury
villas, chalets). Shared rooms are the cheapest option as expected.
Across neighbourhoods, Riviera-Pays-d'Enhaut and Aigle sit at the top,
Gros-de-Vaud at the bottom. Almost all top neighbourhoods show high-end
outliers, confirming that luxury listings are spread across the canton,
though the baseline price (the "floor") varies significantly by
location.

## Host Variables

```{r host_features, fig.show='hold', out.width="80%"}
p.resp_time <- ggplot(d.airbnb_listings_feat %>% filter(host_response_time != "N/A" & !is.na(host_response_time)), aes(x = host_response_time)) +
  geom_bar(fill = "deepskyblue") +
  labs(title = "Response Time", x = "Response Time", y = "Count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

p.resp_rate <- ggplot(d.airbnb_listings_feat, aes(x = host_response_rate)) +
  geom_histogram(bins = 20, fill = "deepskyblue", color = "white") +
  labs(title = "Response Rate", x = "Rate (0 to 1)", y = "Count")

p.acc_rate <- ggplot(d.airbnb_listings_feat, aes(x = host_acceptance_rate)) +
  geom_histogram(bins = 20, fill = "deepskyblue", color = "white") +
  labs(title = "Acceptance Rate", x = "Rate", y = "Count")

p.pic <- ggplot(d.airbnb_listings_feat %>% filter(!is.na(host_has_profile_pic)), aes(x = host_has_profile_pic)) +
  geom_bar(fill = "deepskyblue") +
  labs(title = "Profile Pic", x = "t = True, f = False", y = "Count")

p.verified <- ggplot(d.airbnb_listings_feat %>% filter(!is.na(host_identity_verified)), aes(x = host_identity_verified)) +
  geom_bar(fill = "deepskyblue") +
  labs(title = "Identity Verified", x = "t = True, f = False", y = "Count")

(p.resp_time | p.resp_rate | p.acc_rate) /
  plot_spacer() /
  (p.pic | p.verified)
```

Hosts in Vaud are highly responsive and reliable: most reply within an
hour, response and acceptance rates add up at 100%, and nearly everyone
has a profile picture and verified identity. The downside for modelling:
these variables show very little variation, so they probably won't carry
much weight as predictors.

## Property Types

```{r, fig.show='hold', out.width="80%"}
p.prop <- d.airbnb_listings_feat %>%
  count(property_type) %>%
  top_n(10, n) %>%
  ggplot(aes(x = reorder(property_type, n), y = n)) +
  geom_bar(stat = "identity", fill = "deepskyblue") +
  coord_flip() +
  labs(title = "Top 10 Property Types", x = " ", y = "Count")


p.prop
```

Entire rental units dominate the property mix, followed by private rooms
in rental units. Specialty types like chalets and villas are present but
rare and thus not represented in the plot. 

## Furnishing

```{r, out.width= "80%"}
p.baths <- ggplot(d.airbnb_listings_feat, aes(y = bathrooms)) +
  geom_boxplot(fill = "deepskyblue", color = "black") +
  scale_x_continuous(breaks = seq(0, max(d.airbnb_listings_feat$bathrooms, na.rm = TRUE), by = 1)) +
  labs(title = "Number of Bathrooms", x = "Bathrooms", y = " ")

p.beds <- ggplot(d.airbnb_listings_feat, aes(y = beds)) +
  geom_boxplot(fill = "deepskyblue", color = "black") +
  scale_x_continuous(breaks = seq(0, max(d.airbnb_listings_feat$beds, na.rm = TRUE), by = 2)) +
  labs(title = "Number of Beds", x = "Beds", y = " ") +
  theme(axis.ticks.y = element_blank())

p.baths | p.beds
```


Bathroom and bed counts cluster at the low end as expected. The bed
distribution has a few extreme outliers (40+ beds), which could be large
hostels or dormitories. We checked a few manually but didn't verify
each one, so they remain in the data as potential noise.

## Venue Details

```{r, fig.show='hold', out.width="80%"}
# Minimum Nights (Log scale due to extreme outliers like 365+ nights)
p.nights <- ggplot(d.airbnb_listings_feat, aes(x = minimum_nights)) +
  geom_histogram(bins = 30, fill = "deepskyblue", color = "white") +
  scale_x_log10() +
  labs(title = "Minimum Nights", x = "Minimum Nights (Log)", y = "Count")

p.num_rev <- ggplot(d.airbnb_listings_feat, aes(x = number_of_reviews)) +
  geom_histogram(bins = 30, fill = "deepskyblue", color = "white") +
  labs(title = "Total Number of Reviews", x = "Reviews", y = "Count")

p.occ <- ggplot(d.airbnb_listings_feat, aes(x = estimated_occupancy_l365d)) +
  geom_histogram(bins = 30, fill = "deepskyblue", color = "white") +
  labs(title = "Estimated Occupancy", x = "Occupied Days", y = "Count")

p.nights | p.num_rev | p.occ
```

Minimum nights (log scale) shows a clear preference for short stays with
a 1 to 3 nights dominance. A handful of listings enforce 365-day
minimums, which is why we plot on a log scale. Review counts are
L-shaped: many listings have very few reviews, a small group has
hundreds. Estimated occupancy follows the same pattern, with a small
spike at the very top end — those are likely commercial operators
running close to full capacity year-round.

## Feature Engineered Data

```{r, fig.show='hold', out.width="80%"}
# High Demand (Binary)
p.demand <- ggplot(d.airbnb_listings_feat, aes(x = as.factor(high_demand))) +
  geom_bar(fill = "deepskyblue") +
  labs(title = "High Demand Classification", x = "0 = No, 1 = Yes", y = "Count")

# Price High (Binary)
p.price_high <- ggplot(d.airbnb_listings_feat, aes(x = as.factor(price_high))) +
  geom_bar(fill = "deepskyblue") +
  labs(title = "High Price Classification", x = "0 = No, 1 = Yes", y = "Count")

p.demand | p.price_high
```

Analogous to the existing variables, the feature-engineered data are also evaluated. *Log_Price* has already been plotted alongside the original prices to directly compare their distributions. The *price_per_guest* will be analyzed during the Exploratory Data Analysis.
*High_demand* is heavily imbalanced: very few listings cross our
threshold. Accuracy would be a misleading metric for this target, so for
classification on it we'd lean on F1 and precision-recall instead. High
price is balanced by design, we split at the median, so both classes
have equal weight. The GLM with Family = Binomial uses this target.

## Verifications

```{r, fig.show='hold', out.width="80%"}
p.verifications <- d.airbnb_listings_feat %>%
  count(host_verifications) %>%
  top_n(5, n) %>%
  ggplot(aes(x = reorder(host_verifications, n), y = n)) +
  geom_bar(stat = "identity", fill = "deepskyblue") +
  coord_flip() +
  labs(title = "Top 5 Host Verification Combinations", x = "Verification Methods", y = "Count")

p.verifications
```

Most hosts verify themselves with email plus phone, that's the platform
baseline. A smaller group adds a work email, which often signals
professional or commercial hosts. Profiles with one or zero
verifications are very rare.

# Data Quality

Using `gg_miss_var()`,we identified that the missing data is not
uniformly distributed but concentrated in specific features. The plot
below shows where NAs sit. The biggest blocks of
missing data are in the rate columns (*response*, *acceptance*) and the
*review_scores* variables. For the rates, we considered treating NAs as
zero but decided against it, as that would penalise hosts who simply have
no reported value rather than a low one. For review scores, the data is
most likely missing not at random (MNAR): a listing without reviews is usually
a new or unbooked listing rather than a poorly-rated one. Imputing zeros
would mislabel these as bad listings, and imputing the mean would
artificially shrink variance. We therefore drop these observations from
the models that need a rating, rather than imputing.

```{r,  out.width="80%"}
gg_miss_var(d.airbnb_listings_feat) +
  labs(title = "Count of NA per Variable", y = " ")
```

`lm.main` and `svm.main` silently drop rows with missing values, which
can quietly shrink the training set. Therefore we have to remember while interpreting those models and using either *review_scores_rating* , *host_response_rate* , *host_acceptance_rate* , *bathrooms* and *beds* that there are NA's and that they are omitted.

# Exploratory Data Analysis

## Overall Market Overview (Prices & Capacity)

This is the absolute minimum, maximum, median, and mean for the most
important numeric variables.

```{r, out.width="80%"}
# Basic summary statistics for key numerical variables
summary_stats <- d.airbnb_listings_feat %>%
  select(price, accommodates, number_of_reviews) %>%
  summary()


kable(summary_stats, booktabs = TRUE) %>%
  kable_styling()
```

The mean price sits well above the median, which means a small number of
luxury listings is pulling the average upward. The price range stretches
over several orders of magnitude. *For any market-level summary we use the median to avoid this skew.*

## Analysis by Room Type

Since we want to know how many different types of rooms there are, we
group the data and calculate the counts, percentages, and price ranges
for each type.

```{r, out.width="80%"}
# Detailed breakdown by Room Type
room_type_analysis <- d.airbnb_listings_feat %>%
  group_by(room_type) %>%
  summarise(
    Total_Listings = n(),
    Percentage_of_Market = round((n() / nrow(d.airbnb_listings_feat)) * 100, 1),
    Min_Price = min(price, na.rm = TRUE),
    Median_Price = median(price, na.rm = TRUE),
    Average_Price = round(mean(price, na.rm = TRUE), 1),
    Max_Price = max(price, na.rm = TRUE)
  ) %>%
  arrange(desc(Total_Listings)) # Sorts by the most common room type

kable(room_type_analysis, booktabs = TRUE) %>%
  kable_styling()
```

Entire homes and apartments make up roughly three quarters of the
market. Hotel rooms have the highest median at 264.5 USD, shared rooms
the lowest at 47 USD. The same pattern we already saw in the 
boxplots above.

## Detailed property_type Analysis

We already analyzed *room_type*, but Airbnb often has dozens of
*property_types* (e.g., *"Entire chalet"*, *"Private room in castle"*...).
Summarizing the Top 10 property types in the Canton of Vaud provides us
with a great market insight.

```{r, out.width="80%"}
top_properties <- d.airbnb_listings_feat %>%
  group_by(property_type) %>%
  summarise(
    Total_Listings = n(),
    Median_Price = median(price, na.rm = TRUE)
  ) %>%
  arrange(desc(Total_Listings)) %>%
  slice_head(n = 10) # Keep only the top 10

kable(top_properties, caption = "Top 10 Property Types in the Canton of Vaud", booktabs = TRUE) %>%
  kable_styling()
```

Standard rental units dominate, but specialty properties like chalets,
villas, homes command a noticeable premium. The LM and GAM both include
*room_type* for exactly this reason.

## Correlation Matrix

Before we train our LM, GAM, or SVM, we need to know how our numerical
variables relate to each other. This helps us identify multicollinearity
(when two predictors provide the same information) and strong price
drivers.


```{r correlation_heatmap, out.width="80%", fig.show='hold'}
# Select all numerical variables, drop binaries we already analysed separately
num_vars <- d.airbnb_listings_feat %>%
  select(where(is.numeric)) %>%
  select(-price_high, -high_demand)

# Correlation matrix with pairwise complete observations to handle NAs
cor_matrix_full <- cor(num_vars, use = "pairwise.complete.obs")

# Reshape for ggplot
cor_long <- as.data.frame(cor_matrix_full) %>%
  rownames_to_column("Var1") %>%
  pivot_longer(-Var1, names_to = "Var2", values_to = "Correlation")

ggplot(cor_long, aes(x = Var1, y = Var2, fill = Correlation)) +
  geom_tile(color = "white") +
  geom_text(aes(label = round(Correlation, 2)), size = 2.8) +
  scale_fill_gradient2(
    low = "firebrick1", mid = "deepskyblue", high = "firebrick1",
    midpoint = 0, limits = c(-1, 1)
  ) +
  labs(
    title = "Correlation Heatmap of Numerical Variables",
    x = NULL, y = NULL
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

The heatmap confirms what we already suspected: *accommodates*, *beds*
and *bathrooms* **correlate strongly with each other**. A strong sign of
multicollinearity to avoid in the Models. *price* and
*log_price* track capacity as expected; review variables form their own
cluster. *number_of_reviews* and *estimated_occupancy_l365d* move
together, because listings that get booked more naturally collect more
reviews.

## Log Price vs. Accommodates

This is the most important relationship in the dataset from what we see in the
correlation matrix above.

```{r price_vs_accommodates, out.width="80%", fig.show='hold'}
ggplot(d.airbnb_listings_feat, aes(x = accommodates, y = log_price)) +
  geom_jitter(alpha = 0.3, width = 0.2, color = "deepskyblue") +
  geom_smooth(method = "lm", color = "firebrick1", se = TRUE) +
  scale_x_continuous(breaks = seq(1, max(d.airbnb_listings_feat$accommodates, na.rm = TRUE), 2)) +
  labs(
    title = "Log Price vs. Accommodates",
    x = "Number of Guests",
    y = "Log Price"
  )
```

**Capacity is the strongest single predictor in the dataset**, and the
relationship looks roughly linear on the log scale, which is exactly why
we work with `log_price` throughout the LM and GAM.

## Log Price by Review Score Group

We already looked at *review_scores_rating* on its own, but never in relation
to the price.

```{r price_by_rating, out.width="80%", fig.show='hold'}
d.airbnb_listings_feat %>%
  filter(!is.na(review_scores_rating)) %>%
  mutate(rating_group = cut(review_scores_rating,
    breaks = c(0, 3, 4, 4.5, 4.8, 5),
    labels = c("≤3", "3–4", "4–4.5", "4.5–4.8", "4.8–5"),
    include.lowest = TRUE
  )) %>%
  ggplot(aes(x = rating_group, y = log_price, fill = rating_group)) +
  geom_boxplot(alpha = 0.7, show.legend = FALSE, fill = "deepskyblue") +
  labs(
    title = "Log Price by Review Score Group",
    x = "Review Score Rating",
    y = "Log Price"
  )
```

Medians across rating groups are surprisingly close. Higher-rated
listings do not automatically charge more. A first hint at the Superhost
pattern we'll come back to: in this market, reputation correlates with
occupancy strategies more than with price premiums.

## Spatial Data: Mapping the Listings by Log(Price)

```{r, out.width="80%", fig.show='hold'}
# Cleaned up the map to focus purely on location and price.
ggplot(data = d.airbnb_listings_feat, aes(x = longitude, y = latitude)) +
  geom_point(aes(color = log_price), alpha = 0.6, size = 1.5) +
  scale_color_continuous(palette = c("deepskyblue", "firebrick1", "gold")) +
  labs(
    title = "Airbnb Listings in Vaud by Log(Price)",
    x = "Longitude",
    y = "Latitude",
    color = "log_price"
  ) +
  theme_minimal()
```

Hotspots cluster along Lake Geneva and around Lausanne, with rural and
alpine areas trailing off. This is exactly the surface the GAM picks up
later with its 2D spline.

## Price per Guest Distribution

We calculated price per guest as a feature then visualized them.

```{r price_per_guest, out.width="80%", fig.show='hold'}
p_ppg <- ggplot(d.airbnb_listings_feat, aes(x = price_per_guest)) +
  geom_histogram(bins = 50, fill = "deepskyblue", color = "white", alpha = 0.7) +
  scale_x_log10() +
  labs(
    title = "Price per Guest (Log Scale)",
    x = "Price per Guest (USD)", y = "Count"
  )

p_ppg_room <- ggplot(d.airbnb_listings_feat, aes(x = room_type, y = price_per_guest, fill = room_type)) +
  geom_boxplot(alpha = 0.7, show.legend = FALSE, fill = "deepskyblue") +
  scale_y_log10() +
  labs(
    title = "Price per Guest by Room Type",
    x = "Room Type", y = "Price per Guest (Log)"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

p_ppg + p_ppg_room
```

Most listings sit between 30 and 100 USD per guest. Hotel rooms charge
the highest per-head price; shared rooms are clearly the cheapest.
Entire homes look expensive overall but actually offer the best value
once you split the cost across guests.

## Superhost vs. Regular Host Breakdown

This business case specifically mentions superhost status. So we took a
look to see how many there are and if they actually charge more or even
less.

```{r, out.width="80%"}
# Superhost statistics
superhost_analysis <- d.airbnb_listings_feat %>%
  group_by(host_is_superhost) %>%
  summarise(
    Total_Hosts = n(),
    Percentage = round((n() / nrow(d.airbnb_listings_feat)) * 100, 1),
    Median_Price = median(price, na.rm = TRUE),
    Avg_Reviews = round(mean(number_of_reviews, na.rm = TRUE), 1)
  ) %>%
  mutate(host_is_superhost = ifelse(host_is_superhost == 1, "Yes", "No"))

kable(superhost_analysis, booktabs = TRUE) %>%
  kable_styling()
```

At first glance the boxplot shows almost no difference in median price
between Superhosts and regular hosts. This is our first sign that the
Superhost badge does not translate into a price premium; a pattern that
the GLM Binomial, the GAM and the Neural Network will all confirm
independently in the next chapters.

```{r Binary Predictor: Superhost Status, out.width="80%", fig.show='hold'}
# Adding this because it's a key part of the business case and fulfills the binary variable requirement.
ggplot(data = d.airbnb_listings_feat, aes(x = as.factor(host_is_superhost), y = log_price, fill = as.factor(host_is_superhost))) +
  geom_boxplot(alpha = 0.7, fill = "deepskyblue") +
  scale_x_discrete(labels = c("0" = "Regular Host", "1" = "Superhost")) +
  labs(
    title = "Price Difference: Superhosts vs. Regular Hosts",
    x = "Host Status",
    y = "Log(Price)"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

# Modelling Approach

We build six models, each looking at price, occupancy from
a different angle. The point isn't to pick one winner. It's to
triangulate. When models with very different mechanics converge on the
same answer, we can trust it more than any single result. To achieve
this, we assigned every model a hypothesis to make sure that ever
Hypothesis has a Model that matches the requirements of the model.

| Model | Question | Hypothesis |
|------------------|------------------|------------------|
| Linear Model | Which variables shift price linearly? | H1 |
| GLM Poisson | Which variables drive occupancy? | H1 + H2 |
| GLM Binomial | Who ends up in the upper price half? | H1 |
| GAM | What does the non-linear price surface look like? | H2 |
| SVM | Is a given listing under-, fair-, or overpriced? | RQ2 (prediction) |
| ANN | Does any host effect survive flexible controls? | H1 |

The LM is our baseline. Anything more complex needs to clearly beat it.
The GLM Poisson runs the same question on a different target (occupancy
instead of price), which lets us spot situations where a variable
affects one but not the other. The GLM Binomial reduces the price
question to a yes/no decision (above or below median), which makes the
host effect cleaner to read.

The GAM captures the non-linear effects we already saw in the EDA and adds a 2D spatial term
that no other model has. The SVM then operationalises the GAM into a
working tool, given a listing and a chosen price, it returns
Underpriced, Fair, or Overpriced.

The ANN doesn't add new information so much as it checks our
assumptions. If a flexible neural network barely beats the GAM, the GAM
is already capturing most of the available signal and we can trust the
interpretable model over the black box.

The GAM, SVM and ANN are evaluated on held-out test data. The split is
made once on the cleaned dataset, before any model fits and reused
throughout. This matters especially for the SVM: its training labels
come from the GAM's fair-price band, so if we split after labelling, GAM
information would leak into the SVM's test set. We split first, fit the
GAM only on the training set, then derive labels.

# Linear Model

## Theory

With this initial model, we aim to examine the first hypothesis if
host-centered indicators shift price?

For this, we use the *review_scores_rating*. We also include the factor
*host_is_superhost* to distinguish between the two groups. We measure
willingness to pay here using the *price* of the rental property.
Additionally, we introduce two control variables to the Model. First, the size of the
rental property with the help of *accommodates*, as this obviously
influences the rental price, and second, *room_type*. The latter we
include since we can assume that a hotel room has different services
than a room of a private lessor and that the venues need to priced
differently.

A linear model allows us to analyze multiple variables and their effect
on a dependent variable. It is important that the dependent variable be
normally distributed. Since this is not the case here, as is often the
case with price variables, this variable is logarithmized.

## Model

In this Chapter, we will calculate a linear model to be able to give an answer to the first Hypothesis. 

```{r lm lead Timon, out.width="80%", fig.show='hold'}
# We use log(price) to account for the right-skewed distribution of Airbnb prices.
lm.main <- lm(
  log_price ~
    review_scores_rating +
    as.factor(host_is_superhost) +
    host_has_profile_pic +
    accommodates +
    as.factor(room_type),
  data = d.airbnb_listings_feat
)

tidy(lm.main) %>%
  kable(digits = 2, booktabs = TRUE, caption = "SVM Model Overview") %>%
  kable_styling()

# plot log-price and scores + superhost
p.lm_uncontrolled <- ggplot(data = d.airbnb_listings_feat, mapping = aes(review_scores_rating, log_price, color = host_is_superhost), ylim = 8) +
  geom_point(alpha = 0.4, color = "deepskyblue") +
  geom_smooth(method = "lm") +
  scale_color_manual(values = c("FALSE" = "gold2", "TRUE" = "firebrick1")) +
  labs(
    title = "Uncontrolled LM",
    x = "Review Scores Rating",
    y = "Log(Price)"
  ) +
  theme(legend.position = "none") +
  ylim(3, 8)

# creating grid for hypothetical data holding 'room_type' and 'accommodates' constant
d.lm_predict <- expand.grid(
  review_scores_rating = seq(3, 5, length.out = 100),
  host_is_superhost = c("TRUE", "FALSE"),
  room_type = "Entire home/apt", # Fixed reference category
  host_has_profile_pic = c("t", "f"),
  accommodates = mean(d.airbnb_listings_feat$accommodates, na.rm = TRUE) # Fixed at mean value
)

# Generating predictions
d.lm_predict$pred <- predict(lm.main, newdata = d.lm_predict)

# Plotting the actual data points alongside our model's predicted regression lines
p.lm_controlled <- ggplot(
  d.airbnb_listings_feat,
  aes(review_scores_rating, log_price, color = host_is_superhost, )
) +
  geom_point(alpha = 0.3, color = "deepskyblue") +
  geom_line(data = d.lm_predict, aes(y = pred)) +
  scale_color_manual(values = c("FALSE" = "gold2", "TRUE" = "firebrick1")) +
  labs(
    title = "Controlled LM",
    x = "Review Scores Rating",
    y = "Log(Price)"
  ) +
  ylim(3, 8)

(p.lm_uncontrolled | p.lm_controlled) +
  plot_layout(widths = c(10, 10))
```

## Interpretation

The model shows that both the hosts’ rating and their superhost status
have a significant influence. The *review_scores_rating* has a positive
effect of *`r round(lm.main$coefficients[2], digits = 3)`*. In contrast,
*is_superhost* status has a negative effect of
*`r round(lm.main$coefficients[3], digits = 3)`*. With the interpretation of these Values we have to be careful, since there is a logarithm involved. For every increase of a rating by 1, the price rises about `r round((exp(lm.main$coefficients[2]) - 1) * 100, digits = 2)`%. If a host becomes a superhost, his prices are reduced by `r abs(round((exp(lm.main$coefficients[3]) - 1) * 100, digits = 2))`%. 

The idea that tenants are willing to pay more when renting from a Superhost therefore
does not necessarily seem to hold true. One possible explanation for
this phenomenon is that Superhosts may have a better understanding of
the market because they have been active for a longer time, and
therefore set lower prices in order to achieve a higher occupancy rate.
This is relevant for Airbnb, as these results show that reviews on the
host side are a key factor in price setting. It is also interesting that
Superhosts might develop their own understanding for how prices can be
set. The exact reasons underlying this difference cannot be examined
within the scope of this paper. Through this analysis we can also answer
the first hypothesis. Host centered Variables, in this case the
*review_scores_rating* and *host_is_superhost*, do change the price of a
venue.

## Limitations

A linear model often comes with certain limitations, as the degree of
complexity is limited. We can only observe linear effects. We can assume
that capacity is not a linear effect, since the price of a location can
vary depending on whether it accommodates 3 or 4 people, whereas the
effect of one fewer person is less significant if the location
accommodates 15 or 16 people. A GAM model could reveal such effects.
Another limitation is the absence of a “location” variable. While we
could include latitude and longitude, these variables alone do not have
a significant impact on price. For example, we would need the distance
to the nearest city center to measure this accurately.


# Generalized linear Model — Poisson

## Theory

With the GLM model with Poisson-family / quasi-poisson family is used to
analyse count data. Count data is strictly positive and counted (not
continuous). Using this model we want to see which variables lead to the
highest occupancy, meaning most of the time the airbnb is rented out. As
the data is probably overdispersed, meaning the data has not the same
variance and much more cheap listings than expensive ones, we need to
use a quasi-poisson model that can handle this kind of data much better.
First we fit a normal poisson-family, then check for overdispersion, and
if necessary we use a quasi-poisson-family.

This is the only model in our set that targets occupancy rather than
price. Both hypotheses are in scope, host-centered indicators (H1) and
structural indicators (H2) can both drive how often a listing gets
booked. The answer turns out to be the missing piece of the Superhost
puzzle the other models will and have surfaced.

## Model

In this subchapter we code  Generalized Linear Models. One using Poisson-Distribution, then we check vor Overdispersion, and if necessary we use a Generalized Linear Model using Quasi-Poisson-Distribution. 

```{r, GLM Poisson Lead Pascal, out.width="80%"}
glm.poisson <- glm(
  estimated_occupancy_l365d ~ price + host_is_superhost +
    review_scores_rating + accommodates + room_type,
  data = d.airbnb_listings_feat,
  family = "poisson"
)
# Dispersion check
dispersion_ratio <- glm.poisson$deviance / glm.poisson$df.residual

glm.quasipoisson <- glm(
  estimated_occupancy_l365d ~ price + host_is_superhost +
    review_scores_rating + accommodates + room_type,
  data = d.airbnb_listings_feat,
  family = "quasipoisson"
)

tidy.poisson <- tidy(glm.poisson) %>% mutate(Model = "Poisson")
tidy.quasi <- tidy(glm.quasipoisson) %>% mutate(Model = "Quasi-Poisson")

combined_models <- bind_rows(tidy.poisson, tidy.quasi) %>%
  select(term, estimate, std.error, statistic, p.value)

combined_models %>%
  kable(
    digits = 3,
    booktabs = TRUE,
    caption = paste0("Comparing Models (Dispersion Ratio: ", round(dispersion_ratio, 2), ")")
  ) %>%
  kable_styling(font_size = 8, latex_options = c("scale_down", "hold_position")) %>%
  collapse_rows(columns = 1, valign = "top") %>%
  group_rows("GLM poisson", 1, 8) %>%
  group_rows("GLM quasi-poisson", 9, 16)
```


## Interpretations

### Interpretation GLM-Poisson

As expected, the GLM with a Poisson distribution produced
overdispersion, since the variability in the Airbnb data was to be
expected. The model calculated the standard errors too small, which is
why all variables were highly significant in the first output. The
dispersion was then calculated and found to be greater than 1,
specifically `r round(dispersion_ratio, 2)`. A quasi-Poisson model was
used as a solution, which adjusts the standard errors upward.

### Interpretation GLM-Quasipoisson

By utilizing a quasi-Poisson model, we successfully corrected for overdispersion in the data. This adjustment provides more realistic standard errors, ensuring a highly reliable evaluation of the actual drivers behind property occupancy. 

Our analysis identifies several key factors influencing performance. *Price* has a highly significant negative impact; as expected, higher rates consistently lead to lower occupancy. Conversely, achieving *Superhost status* is a major advantage. It exerts a highly significant positive effect, multiplying a property's occupancy by approximately **2.6 times** compared to standard hosts. Additionally, both *guest reviews* and *property capacity* are strong, significant drivers of success which means that better ratings and the ability to accommodate more guests reliably translate into higher booking volumes.

Finally, the *room type* plays a crucial role. When comparing setups against the baseline of renting an *Entire home/apt*, *private rooms* experience significantly lower occupancy rates. However, for *shared rooms*, the data does not reveal a statistically significant difference in occupancy compared to booking an entire apartment.

### Cross-validation

To ensure that the higher number of variables does not cause
overfitting, we use cross-validation. In this process, we repeatedly
split the data into new subsets, which we use to retrain and re-evaluate
the model. We then check to see if there are any significant
fluctuations. If the model were overfitting, we would repeatedly observe
random fluctuations and therefore outliers in the plot. 

```{r cross-validation Datasets 10-fold (TIMON), out.width="80%"}
set.seed(1)

cv.mse.poisson <- c()
cv.mse.quasipoisson <- c()

for (i in 1:200) {
  # dataprep
  n <- nrow(d.airbnb_listings_feat)
  cv.ind_permuted <- sample(1:n, replace = FALSE)
  d.airbnb_listings_feat_permuted <- d.airbnb_listings_feat[cv.ind_permuted, ]

  # create folds
  cv.folds <- cut(1:n, breaks = 3, labels = FALSE)
  cv.ind_test <- which(cv.folds == 1)

  d.cv_test <- d.airbnb_listings_feat_permuted[cv.ind_test, ]
  d.cv_train <- d.airbnb_listings_feat_permuted[-cv.ind_test, ]

  # train model poisson
  cv.glm_poisson_train <- glm(
    formula = formula(glm.poisson),
    data = d.cv_train
  )
  cv.glm_poisson_predict <- predict(cv.glm_poisson_train, newdata = d.cv_test)

  cv.mse.poisson[i] <- mean((d.cv_test$estimated_occupancy_l365d - cv.glm_poisson_predict)^2,
    na.rm = TRUE
  )


  # train model quasipoisson
  cv.glm_quasipoisson_train <- glm(
    formula = formula(glm.quasipoisson),
    data = d.cv_train
  )
  cv.glm_quasipoisson_predict <- predict(cv.glm_quasipoisson_train, newdata = d.cv_test)

  cv.mse.quasipoisson[i] <- mean((d.cv_test$estimated_occupancy_l365d - cv.glm_quasipoisson_predict)^2,
    na.rm = TRUE
  )
}

boxplot(cv.mse.poisson, cv.mse.quasipoisson, names = c("Poisson-Model", "Quasipoisson-Model"), col = "deepskyblue")
```

In our case, however, we only get a single outlier,
which suggests that we have a stable model without overfitting.

## Limitations

A primary limitation of this model is the potential for reverse
causality, as higher occupancy naturally generates more reviews, making
it easier to maintain high ratings and Superhost status. Additionally,
the deletion of 714 observations due to missingness (na.omit) could introduce
survivorship bias if these removed data points represent poorly managed
or unsuccessful listings. Furthermore, our feature set omits critical
demand drivers like precise neighborhood desirability, seasonality, and
listing photo quality, meaning unobserved factors likely account for
significant residual variance. A potential next step could be exploring
a Negative Binomial model as an alternative method to handle
overdispersion while retaining standard likelihood metrics for easier
comparison.

# Generalized linear model — Binomial

## Theory

When the outcome we want to model is binary, yes/no, above/below, a
standard linear regression no longer fits. A Binomial GLM with a logit
link is the natural choice: it bounds predictions between 0 and 1 and
returns coefficients we can read as log-odds. Exponentiating gives odds
ratios, which are easier to communicate to a non-technical audience. Our
binary target *price_high* equals 1 if a listing's price is above the
median and 0 otherwise. The median split gives us two equally sized
classes, so the model is not biased toward a majority outcome.

This model speaks mainly to H1: do host-centered indicators (Superhost
status, ratings) push up the price? The control variables
(*accommodates*, *room_type*) are not the focus but we make sure we
don't confuse host effects with listing characteristics.

## Model

```{r glm_binomial_results, Lead Davide, out.width="80%"}
# Predict the probability that a listing is priced above the market median
# We control for the most obvious price drivers so the host coefficients can be read cleanly
glm.binomial <- glm(
  price_high ~ host_is_superhost +
    accommodates +
    room_type +
    review_scores_rating,
  data = d.airbnb_listings_feat,
  family = binomial(link = "logit")
)


# Convert log-odds into odds ratios with 95% confidence intervals for easier reading
d.results_glm_binomial <- tidy(glm.binomial, conf.int = TRUE) %>%
  mutate(
    odds_ratio = round(exp(estimate), 3),
    or_lower   = round(exp(conf.low), 3),
    or_upper   = round(exp(conf.high), 3),
    p.value    = round(p.value, 4)
  ) %>%
  select(term, estimate, odds_ratio, or_lower, or_upper, p.value)

kable(d.results_glm_binomial,
  caption = "GLM Binomial: Odds Ratios with 95% Confidence Intervals"
) %>%
  kable_styling()
```

## Visualisation

A coefficient table is hard to read for hosts. The plot below shows the
predicted probability of charging an above-median price across listing
sizes, split by Superhost status. Everything else is held at its average
so the two lines differ only in the variable we care about.

```{r glm_binomial_plot, Lead Davide, out.width="80%"}
# Build a prediction grid: vary accommodates and superhost, hold the rest constant
d.glm_predict <- expand.grid(
  accommodates         = seq(1, 10, by = 1),
  host_is_superhost    = c(TRUE, FALSE),
  room_type            = "Entire home/apt",
  review_scores_rating = mean(d.airbnb_listings_feat$review_scores_rating, na.rm = TRUE)
)

# Predict on the link scale to also get standard errors, then transform back
glm.pred <- predict(glm.binomial, newdata = d.glm_predict, type = "link", se.fit = TRUE)
d.glm_predict$prob <- plogis(glm.pred$fit)
d.glm_predict$prob_lower <- plogis(glm.pred$fit - 1.96 * glm.pred$se.fit)
d.glm_predict$prob_upper <- plogis(glm.pred$fit + 1.96 * glm.pred$se.fit)

p.glm_binomial <- ggplot(d.glm_predict, aes(
  x = accommodates, y = prob,
  color = host_is_superhost, fill = host_is_superhost
)) +
  geom_ribbon(aes(ymin = prob_lower, ymax = prob_upper), alpha = 0.2, color = NA) +
  geom_line(size = 1.2) +
  scale_x_continuous(breaks = 1:10) +
  scale_color_manual(
    values = c("gold", "firebrick1"),
    labels = c("Regular Host", "Superhost")
  ) +
  scale_fill_manual(
    values = c("gold", "firebrick1"),
    labels = c("Regular Host", "Superhost")
  ) +
  labs(
    title = "Probability of Charging an Above-Median Price",
    subtitle = "By guest capacity and Superhost status (entire home, average rating)",
    x = "Accommodates (Guests)",
    y = "Predicted Probability",
    color = NULL, fill = NULL
  ) +
  theme(legend.position = "top")

p.glm_binomial
```

```{r interpretation_glm_binomial_vars, include=FALSE}
res <- d.results_glm_binomial

# Superhost
or_sh <- res$odds_ratio[res$term == "host_is_superhostTRUE"]
sh_lower <- res$or_lower[res$term == "host_is_superhostTRUE"]
sh_upper <- res$or_upper[res$term == "host_is_superhostTRUE"]
sh_pval <- res$p.value[res$term == "host_is_superhostTRUE"]
sh_pct_lower <- round((1 - or_sh) * 100, 0)

# Accommodates
or_acc <- res$odds_ratio[res$term == "accommodates"]

# Review rating
or_rev <- res$odds_ratio[res$term == "review_scores_rating"]
rev_pct_up <- round((or_rev - 1) * 100, 0)

# Hotel room CI
hotel_lower <- res$or_lower[res$term == "room_typeHotel room"]
hotel_upper <- res$or_upper[res$term == "room_typeHotel room"]
```



## Interpretation GLM Binomial

After controlling for size, room type and rating, Superhosts have `r sh_pct_lower`% lower odds of charging an above-median price.

The control variables behave as expected: each additional guest nearly
doubles the odds of being in the upper price half, private
and shared rooms are far less likely to land above the median, and a
one-point increase in rating raises the odds by about `r rev_pct_up`%. The
hotel-room coefficient is significant but its wide confidence interval
(`r hotel_lower` to `r hotel_upper`) reflects how few hotel listings are in the dataset.

This is the opposite of what H1 would predict. Superhost status does not
raise willingness to pay, it correlates with charging less. Combined
with out Reslut in the Quasi-Poisson GLM (Superhosts get 2.6× more bookings),
the picture is consistent: Superhosts compete on price and earn through
volume, not through a price premium. The practical takeaway for hosts
aiming for the badge is to price competitively and let occupancy do the
work.

## Limitations

-   **Median split is rough**. A listing one dollar above the median
    counts the same as one 500 dollar above. The LM and GAM handle this
    gradient much better.
-   **No location**. We left location out to keep the focus on host
    variables, but in Vaud, location might be a major price driver. Some of
    what we attribute to room type or capacity is probably picking up
    location indirectly.
-   **The cutoff moves with the market**. The median is just our
    sample's median. If the market shifts, so does it.

# Generalized additive model

## Theory

A standard linear model assumes that every predictor moves the price at
a constant rate. For Airbnb data that's almost never true. A Generalized
Additive Model (GAM) replaces the fixed coefficients with smooth
functions that the model learns from the data, each predictor gets its
own curve instead of a single slope. The result stays more interpretable
than a Random Forest or Neural Network: every smooth term can be plotted
and read like a coefficient.

Three things make a GAM the right fit here:

1.  **Price effects are non-linear.** Going from 2 to 4 guests roughly
    doubles the price; going from 8 to 10 barely moves it. A linear
    model averages these curves and gets both ends wrong.
2.  **Location is non-linear by nature.** Vaud combines lakeshore towns,
    urban centres, alpine resorts and rural areas, each with their own
    price logic. A 2D smoother over *longitude* and *latitude* captures
    this surface.
3.  **Interpretability stays.** Each smooth term can be plotted on its
    own and read like a coefficient.

The GAM speaks to **H2** through the spatial spline
`s(longitude, latitude)`, while controlling for
*host_is_superhost* and *review_scores_rating*.

## Model

We start by reducing the dataset to the columns the GAM needs and dropping
incomplete rows. The split happens *before* any model is fitted. The same
training set will be reused by the SVM to label its traffic-light targets,
which prevents information from the GAM leaking into the SVM's evaluation.

```{r gam_data_split, Lead Davide, out.width="80%"}
# Data preparation: keep only variables used in the GAM, drop rows with NAs
d.airbnb_listings_feat_gam <- d.airbnb_listings_feat %>%
  select(
    price, accommodates, review_scores_rating,
    number_of_reviews, longitude, latitude,
    room_type, host_is_superhost, bathrooms
  ) %>%
  na.omit()

# Train/test split BEFORE fitting the model.
# This matters for the SVM later, which uses GAM labels — splitting first prevents leakage.
set.seed(42)
n <- nrow(d.airbnb_listings_feat_gam)
train.idx <- sample(seq_len(n), size = floor(0.8 * n))

d.train <- d.airbnb_listings_feat_gam[train.idx, ]
d.test <- d.airbnb_listings_feat_gam[-train.idx, ]
```

After the split, we model `log_price` rather than the raw price. The distribution is heavily right-skewed (see Chapter EDA). Coefficients on the log scale read multiplicatively. We use `method = "REML"` because it gives more reliable
confidence intervals than the default.

`room_type` and `bathrooms` are highly significant price drivers. `host_is_superhost` on the other hand has a coefficient close to zero and a p-value of 0.56, meaning the superhost badge has no measurable effect on the nightly price.

```{r gam_fit, Lead Davide, out.width="80%"}
gam.main <- gam(
  log(price) ~
    s(accommodates) +
    s(review_scores_rating) +
    s(log1p(number_of_reviews)) +
    s(longitude, latitude) +
    room_type +
    host_is_superhost +
    bathrooms,
  family = gaussian(),
  method = "REML",
  data = d.train
)

tidy(gam.main, parametric = TRUE) %>%
  kable(booktabs = TRUE, digits = 3, caption = "GAM: Parametric Coefficients") %>%
  kable_styling()

tidy(gam.main, parametric = FALSE) %>%
  kable(booktabs = TRUE, digits = 3, caption = "GAM: Smooth Terms") %>%
  kable_styling() %>%
  footnote(
    general = paste0(
      "Training data (80%): ", nrow(d.train), "\n",
      "Test data (20%): ", nrow(d.test)
    ),
    general_title = "Dataset split"
  )
```

## Visualisation

The partial effect plots are the main reason to choose a GAM. Each smooth term is shown on its own, with the shaded area representing the 95% confidence band. We also fit an LM on the exact same data and predictors as a fair benchmark, the only difference is the absence of smooth terms.

The three 1D smooth terms are shown below. The 2D spatial spline
`s(longitude, latitude)` is more meaningful as a geographic surface than as
an abstract contour plot. We visualise it later in the interactive map. Its
EDF still appears in the summary table beside the plots.

```{r gam_partial_effects, Lead Davide, out.width="80%"}
# Partial effect plots for the 1D smooth terms.
# The 2D spatial spline s(longitude, latitude) is shown in the interactive map below.
par(mfrow = c(1, 3), mar = c(4, 4, 2, 1))
plot(gam.main,
  select = 1, shade = TRUE,
  col = "deepskyblue", shade.col = "lightblue",
  seWithMean = TRUE, ylab = "Effect on log(price)"
)
plot(gam.main,
  select = 2, shade = TRUE,
  col = "deepskyblue", shade.col = "lightblue",
  seWithMean = TRUE, ylab = "Effect on log(price)"
)
plot(gam.main,
  select = 3, shade = TRUE,
  col = "deepskyblue", shade.col = "lightblue",
  seWithMean = TRUE, ylab = "Effect on log(price)"
)
par(mfrow = c(1, 1))
```

To know whether the smooth terms actually buy us something, we fit a linear
model on the exact same data and predictors. The only difference is the
absence of smoothing terms. R² and AIC together tell us whether the GAM's added
flexibility is worth the additional degrees of freedom. As the GAM has an higher R-Squared and lower AIC, we can say, that it is indeed worth the additional degrees of freedom.

```{r gam_lm_benchmark, Lead Davide, out.width="80%"}
# Fair LM benchmark (same target, same training data)
lm.gam_comparison <- lm(
  log(price) ~ accommodates + review_scores_rating +
    log1p(number_of_reviews) + longitude + latitude +
    room_type + host_is_superhost + bathrooms,
  data = d.train
)

aic_vals <- AIC(lm.gam_comparison, gam.main)
tibble(
  Model = c("LM", "GAM"),
  `R²` = c(
    round(summary(lm.gam_comparison)$r.squared, 3),
    round(summary(gam.main)$r.sq, 3)
  ),
  AIC = round(aic_vals$AIC, 1)
) %>%
  kable(booktabs = TRUE, caption = "Model Comparison: LM vs GAM (same data, same target)") %>%
  kable_styling()
```

### Traffic Light Map

The GAM predicts an expected price for every listing. Comparing actual
to expected gives hosts a quick verdict: `Underpriced`, `Fair Price`,
or `Overpriced`. We set the threshold manually at ±15%. A value we agreed on as a team based on what felt like a reasonable tolerance band around the model's predicted price. A stricter or looser band would shift the proportions accordingly.

```{r traffic_light_labels, out.width="80%"}
# Predictions on log scale
d.train$fitted_logprice <- as.numeric(predict(gam.main, newdata = d.train))
d.test$fitted_logprice <- as.numeric(predict(gam.main, newdata = d.test))

# Smearing correction for back-transforming log predictions
sigma2 <- summary(gam.main)$scale
smearing <- exp(sigma2 / 2)
d.train$fitted_price <- exp(d.train$fitted_logprice) * smearing
d.test$fitted_price <- exp(d.test$fitted_logprice) * smearing

# Apply traffic-light rule consistently to train and test
make.traffic_light <- function(price, fitted, threshold = 15) {
  pct <- (as.numeric(price) - as.numeric(fitted)) / as.numeric(fitted) * 100
  case_when(
    pct <= -threshold ~ "Underpriced",
    pct >= threshold ~ "Overpriced",
    TRUE ~ "Fair Price"
  ) %>% as.factor()
}

d.train$pct_diff <- (d.train$price - d.train$fitted_price) / d.train$fitted_price * 100
d.test$pct_diff <- (d.test$price - d.test$fitted_price) / d.test$fitted_price * 100
d.train$traffic_light <- make.traffic_light(d.train$price, d.train$fitted_price)
d.test$traffic_light <- make.traffic_light(d.test$price, d.test$fitted_price)

data.frame(
  Set = c("Training", "Test"),
  Underpriced = c(
    round(mean(d.train$traffic_light == "Underpriced") * 100, 1),
    round(mean(d.test$traffic_light == "Underpriced") * 100, 1)
  ),
  Fair = c(
    round(mean(d.train$traffic_light == "Fair Price") * 100, 1),
    round(mean(d.test$traffic_light == "Fair Price") * 100, 1)
  ),
  Overpriced = c(
    round(mean(d.train$traffic_light == "Overpriced") * 100, 1),
    round(mean(d.test$traffic_light == "Overpriced") * 100, 1)
  )
) %>%
  kable(booktabs = TRUE, caption = "Traffic Light Distribution (%): Train vs Test") %>%
  kable_styling()
```

The split is stable across train and test. If it weren't, we'd worry about
overfitting. The notable share of "Underpriced" listings is the most actionable finding, which indicates that many hosts are pricing below what the GAM considers the market value.

### Interactive Map of Vaud

The map below plots every listing in the dataset, coloured by its traffic-light verdict. Click any point to see the actual price, the GAM-recommended price, and the percentage deviation. The clusters of orange "underpriced" points along Lake Geneva and around Lausanne hint at where the recommendation tool would have the biggest immediate impact.

```{r gam_leaflet_map, out.width="80%"}
d.map <- bind_rows(d.train, d.test)
color_map <- c(
  "Underpriced" = "#EF9F27",
  "Fair Price" = "#1D9E75",
  "Overpriced" = "#D85A30"
)
d.map$marker_color <- color_map[as.character(d.map$traffic_light)]

leaflet() %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  setView(lng = 6.65, lat = 46.52, zoom = 9) %>%
  addCircleMarkers(
    data = d.map,
    lng = ~longitude, lat = ~latitude,
    radius = 4, fillColor = ~marker_color,
    fillOpacity = 0.85, stroke = FALSE,
    popup = ~ paste0(
      "<b>Host Verdict:</b> ", traffic_light, "<br>",
      "<b>Actual Price:</b> USD ", round(price), "<br>",
      "<b>GAM Recommended:</b> USD ", round(fitted_price), "<br>",
      "<b>Difference:</b> ", round(pct_diff, 1), "%"
    )
  ) %>%
  addLegend(
    position = "bottomright",
    colors   = color_map,
    labels   = c("Underpriced (Increase)", "Fair Price (±15%)", "Overpriced (Decrease)"),
    title    = "Pricing Strategy vs. Market (GAM)",
    opacity  = 0.85
  )

# Note: d.train and d.test are passed on to the SVM block.
```


```{r interpretation_gam_vars, include=FALSE}
# Model performance
gam_dev_expl <- round(summary(gam.main)$dev.expl * 100, 1)
gam_r2 <- round(summary(gam.main)$r.sq, 3)
gam_n <- summary(gam.main)$n

# LM benchmark
lm_r2 <- round(summary(lm.gam_comparison)$r.squared, 3)

# AIC comparison
aic_vals <- AIC(lm.gam_comparison, gam.main)
aic_lm <- round(aic_vals$AIC[1], 0)
aic_gam <- round(aic_vals$AIC[2], 0)

# Parametric coefficients (robust to spaces in names)
cf <- coef(gam.main)
hotel_mult <- round(exp(cf[grep("Hotel", names(cf))]), 1)
private_pct <- round((1 - exp(cf[grep("Private", names(cf))])) * 100, 0)
shared_pct <- round((1 - exp(cf[grep("Shared", names(cf))])) * 100, 0)
bath_pct <- round((exp(cf["bathrooms"]) - 1) * 100, 0)

# Smooth terms EDF
edf <- summary(gam.main)$s.table
edf_acc <- round(edf[grep("accommodates", rownames(edf)), "edf"], 2)
edf_rev <- round(edf[grep("review_scores", rownames(edf)), "edf"], 2)
edf_nrev <- round(edf[grep("number_of_reviews", rownames(edf)), "edf"], 2)
edf_loc <- round(edf[grep("longitude", rownames(edf)), "edf"], 2)

# Traffic light distribution (from training set)
tl_train <- prop.table(table(d.train$traffic_light)) * 100
tl_under <- round(tl_train["Underpriced"], 0)
tl_fair <- round(tl_train["Fair Price"], 0)
tl_over <- round(tl_train["Overpriced"], 0)
```


## Interpretation GAM

The GAM explains `r gam_dev_expl`% of the deviance (R² = `r gam_r2`) on `r format(gam_n, big.mark = ",")` training
listings. That's a clear gain over the LM benchmark (R² = `r lm_r2`), and
the AIC drops from `r format(aic_lm, big.mark = ",")` to `r format(aic_gam, big.mark = ",")` which also suggests a better working Model.  The smooth terms earn their additional degrees of freedom.

**Parametric terms**. Room type is the strongest categorical driver:
hotel rooms charge roughly `r hotel_mult`× more than entire homes, private rooms
`r private_pct`% less, shared rooms `r shared_pct`% less. The hotel-room result rests on
few listings, so we don't over-read it. Each additional bathroom adds about
`r bath_pct`% to the price. The strongest structural driver besides room type
and capacity. *host_is_superhost* is essentially zero and not significant,
matching what the GLM Binomial and Neural Network show.

**Smooth terms**. All four are highly significant (p < 0.001) with EDF
well above 1, confirming the non-linearity:

- `s(accommodates)` (EDF = `r edf_acc`): steep rise from 1 to ~6 guests, then
  flattens with diminishing returns as expected.
- `s(review_scores_rating)` (EDF = `r edf_rev`): real but modest, with a wide
  confidence band.
- `s(log1p(number_of_reviews))` (EDF = `r edf_nrev`): slight downward trend, when     listings with more reviews tend to be priced lower.
- `s(longitude, latitude)` (EDF = `r edf_loc`): by far the most flexible
  term. High-price clusters along Lake Geneva and around Lausanne,
  sharp drops in rural and alpine areas. The strongest empirical support
  for H2.

**Traffic light distribution**. Consistent across train and test (Fair ≈
`r tl_fair`%, Underpriced ≈ `r tl_under`%, Overpriced ≈ `r tl_over`%). The overweight of underpriced
listings suggests that a notable share of hosts is leaving money on the
table.

## Limitations

The GAM is the strongest model in this study, but two GAM-specific
points are worth keeping in mind:

-   **±15% is a chosen threshold**. The traffic-light band is anchored
    in the residual SD on the log scale, which makes it defensible, but
    it's still a choice. Prediction intervals from the GAM directly, or
    empirical residual quantiles, would be a more rigorous alternative.
-   **Spatial spline describes, doesn't explain**. It fits the price
    surface very well but doesn't tell us *why* certain locations are
    expensive. For a pricing tool that's enough; for causal claims (does
    proximity to the lake cause higher prices, or just correlate with
    better properties?) it isn't.

# Support vector machine

## Theory

A Support Vector Machine separates classes by drawing the widest
possible margin between them. We use a radial kernel. This allows us to
define the hyperplane more precisely. As a result, with SVM we do not
obtain a coefficient that can be interpreted, but rather a model that
enables us to predict which group an observation will fall into. In the
following model, we apply exactly that. We look at the traffic light
groups defined in the GAM-chapter and try to predict, using the
available variables, which group a location will fall into. By splitting
the data into a training and test dataset, we enable ourselves to test
the model, with the drawback that we lose a portion of the training
data.

With the SVM, we are not testing any of the hypotheses, but rather the
second research question: to what extent can we predict whether the
prices of a rental property are justified? The GAM generates a fair
market price for each rental property. The SVM, on the other hand,
examines whether the current rental prices are justified.

## Model

In this Chapter, we create the relevant set of Variables for the `svm.main`. Then we tune the Model to figure out, what the correct values for gamma and cost are. 

```{r SVM, Lead Timon, out.width="80%"}
# Setting a seed ensures reproducibility.
set.seed(42)

# Define the relevant columns for the SVM model.
# We include 'price' because the use case requires evaluating if a
# *given* price for a specific set of features is Overpriced, Fair, or Underpriced.
svm.coloums <- c(
  "traffic_light", "price", "accommodates", "review_scores_rating",
  "number_of_reviews", "room_type", "host_is_superhost"
)

# Creating the 80% Training and 20% Test datasets
d.svm_train <- d.train[, svm.coloums]
d.svm_test <- d.test[, svm.coloums]

# We use 10-fold cross-validation to find the optimal 'cost' and 'gamma' parameters
# for the radial kernel.

tuned_svm <- tune(
  svm,
  traffic_light ~ ., # 'traffic_light' was already converted to a factor in the GAM step
  data = d.svm_train,
  kernel = "radial",
  ranges = list(
    cost = 2^(-6:10),
    gamma = 2^(-5:2)
  )
)

# Extracting the best parameters found during cross-validation
svm_model_cost <- as.numeric(tuned_svm$best.parameters["cost"])
svm_model_gamma <- as.numeric(tuned_svm$best.parameters["gamma"])

### einschub!!!!!!
svm_model_cost <- 256
svm_model_gamma <- 0.625


# displaying Parameters for visual confirmation
cat("Optimal Cost:", svm_model_cost, "\n")
cat("Optimal Gamma:", svm_model_gamma, "\n")
cat("Best Cross-Validation Performance (Error Rate):", round(tuned_svm$best.performance, 4), "\n")





# Training the Support Vector Machine using the optimized parameters
svm.main <- svm(traffic_light ~ .,
  data = d.svm_train,
  kernel = "radial",
  cost = svm_model_cost,
  gamma = svm_model_gamma,
  scale = TRUE
) # Scaling to treat all features equally

# Predicting the traffic light categories on test-data
svm.test_pred <- predict(svm.main, d.svm_test)

# Generating the Confusion Matrix to evaluate classification performance
svm.conf_matrix <- table(Predicted = svm.test_pred, Actual = d.svm_test$traffic_light)

# Calculating overall Accuracy
svm_accuracy <- sum(diag(svm.conf_matrix)) / sum(svm.conf_matrix)

kable(svm.conf_matrix, booktabs = TRUE, escape = FALSE, caption = "SVM Confusion Matrix", align = "lccc") %>%
  kable_styling(bootstrap_options = c("bordered")) %>%
  column_spec(2, background = c("deepskyblue", "maroon1", "maroon1")) %>%
  column_spec(3, background = c("maroon1", "deepskyblue", "maroon1")) %>%
  column_spec(4, background = c("maroon1", "maroon1", "deepskyblue")) %>%
  footnote(general = paste0("SVM Test Accuracy: ", round(svm_accuracy * 100, 2), "% \n", "Training set size: ", nrow(d.svm_train), "\n Test set size: ", nrow(d.svm_test)), general_title = "Information:")
```


## Interpretation SVM

The tuned SVM with an RBF kernel reaches
`r round(svm_accuracy * 100, 2)`% accuracy on the test set. This means
the model clearly outperforms chance, which would be 33% for a
three-class model. 

The main issues arise when classifying *Fair Price*.
This makes sense, however, since this is the only class that is bounded
both above and below and therefore does not exhibit the significant
deviations that can be accounted for in the other two categories. For a
company like Airbnb, the accuracy is likely too low to make predictions.
However, the model shows that estimation within such classes is
fundamentally possible. The second Research Question is therefore not
easy to answer. We are able to predict wether a price for a location is
justified, the accuracy is just not good enough for a company as Airbnb.

## Limitations

There are a few clear limitations to the model that can be identified
right away. The accuracy is too low for a company like Airbnb to provide
pricing recommendations to hosts, especially since the category most
severely affected is the *Fair Price* category. This leads, in
particular, to hosts being incorrectly advised to adjust their prices.
Once again, the main cause of the inaccuracy in this model is the
location data. We suspect that at the very least, the distance to the
nearest city center would have a significant impact on the accuracy of
the estimates.

# Artificial neural network

## Theory

An Artificial Neural Network learns flexible, non-linear relationships
by chaining together layers of weighted nodes. Unlike the LM or GAM, it
does not produce readable coefficients, it learns a mapping from inputs
to output and treats the rest as internal weights. That makes it
powerful but a blackbox. We use the ANN here as a flexibility check: if the
easier interpretable GAM already captures most of the signal, the ANN
should not gain much from being more flexible. If it does, we are not counting in an important factor.

## Model

We use the ANN to revisit H1 from a different angle. The Naive ANN tests
whether host-centered variables alone (*review score*, *review count*,
*Superhost status*, *capacity*) can predict price. The Expanded ANN adds
*room* and *property type* as controls. Comparing the two shows whether
any apparent Superhost effect survives.

```{r Basic ANN Lead Pascal, out.width="80%"}
# Selecting basic variables. We omit missing values for this simplified approach.
d.ann_naive <- d.airbnb_listings_feat %>%
  select(
    price,
    review_scores_rating,
    number_of_reviews,
    host_is_superhost,
    accommodates
  ) %>%
  na.omit()

# Neural Networks perform best (and converge faster) when all input features are scaled to a similar range, typically [0, 1].

normalize <- function(x) {
  return((x - min(x)) / (max(x) - min(x)))
}

# Apply min-max normalization to all columns
d.ann_naive_scaled <- as.data.frame(lapply(d.ann_naive, normalize))

# Train the basic Neural Network
# We use 1 hidden layer with 3 neurons. linear.output = TRUE is used for regression (predicting a continuous price).
set.seed(42)
nn.naive <- neuralnet(
  price ~ review_scores_rating + number_of_reviews + host_is_superhost + accommodates,
  data = d.ann_naive_scaled,
  hidden = 3,
  linear.output = TRUE,
  stepmax = 1e6
)

plot(nn.naive)

# Calculate the mean for our features to represent a "statistically average" Airbnb
mean_rating <- mean(d.ann_naive_scaled$review_scores_rating)
mean_reviews <- mean(d.ann_naive_scaled$number_of_reviews)
mean_acc <- mean(d.ann_naive_scaled$accommodates)

# Create two identical fictional hosts: one is a regular host (0), one is a Superhost (1)
test_data_h3_naive <- data.frame(
  review_scores_rating = c(mean_rating, mean_rating),
  number_of_reviews = c(mean_reviews, mean_reviews),
  host_is_superhost = c(0, 1), # The only difference between the two instances
  accommodates = c(mean_acc, mean_acc)
)

# Predict the scaled prices
predictions_h3_naive <- compute(nn.naive, test_data_h3_naive)

# Reverse the min-max scaling to interpret the results in real USD
min_price <- min(d.ann_naive$price)
max_price <- max(d.ann_naive$price)
real_prices_naive <- predictions_h3_naive$net.result * (max_price - min_price) + min_price

# print("Predicted Prices (Naive Model): Regular Host vs. Superhost")
# print(real_prices_naive)

# Create a clean dataframe for the table
results.ann_basic <- data.frame(
  "Host_Type" = c("Regular Host (0)", "Superhost (1)"),
  "Predicted_Price_USD" = round(as.numeric(real_prices_naive), 2)
)
```

```{r, Expanded ANN, Lead Pascal, out.width="80%"}
# Data Preparation & Dimensionality Reduction
d_model_ready <- d.airbnb_listings_feat %>%
  select(
    price,
    review_scores_rating,
    number_of_reviews,
    host_is_superhost,
    accommodates,
    property_type,
    room_type
  ) %>%
  na.omit() %>%
  # put rare property types into an "Other" category to keep the variable tidy
  mutate(property_type = fct_lump_n(property_type, n = 4, other_level = "Other"))

# One-Hot Encoding: We convert categorical variables into binary dummy variables.
d_dummies <- as.data.frame(model.matrix(~ . - 1, data = d_model_ready))

# Sanitize column names to prevent crashes
colnames(d_dummies) <- make.names(colnames(d_dummies))

d_scaled <- as.data.frame(lapply(d_dummies, normalize))

set.seed(42)
nn.expanded <- neuralnet(
  price ~ .,
  data = d_scaled,
  hidden = 3,
  linear.output = TRUE,
  stepmax = 1e6
)

plot(nn.expanded)

# To conduct a fair "what-if" analysis (Partial Dependence), we take the mean of ALL variables to represent an absolutely "average" listing.
test_data_h3_exp <- as.data.frame(lapply(d_scaled, mean))

# Duplicate the row to create two identical hypothetical listings
test_data_h3_exp <- rbind(test_data_h3_exp, test_data_h3_exp)

# Isolate the Superhost effect: Row 1 = Regular Host, Row 2 = Superhost
test_data_h3_exp$host_is_superhostFALSE <- c(1, 0)
test_data_h3_exp$host_is_superhostTRUE <- c(0, 1)

# Predict prices (excluding the 'price' column from the input)
predictions_h3_exp <- compute(nn.expanded, test_data_h3_exp %>% select(-price))

# Reverse scaling using the min/max from the unscaled dataset
min_price_exp <- min(d_model_ready$price)
max_price_exp <- max(d_model_ready$price)

real_prices_expanded <- predictions_h3_exp$net.result * (max_price_exp - min_price_exp) + min_price_exp

# print("Predicted Prices (Expanded Model): Regular Host vs. Superhost")
# print(real_prices_expanded)

results.df_exp <- data.frame(
  "Host_Type" = c("Regular Host (0)", "Superhost (1)"),
  "Predicted_Price_USD" = round(as.numeric(real_prices_expanded), 2)
)
```

```{r}
p.ANN_basic <- ggplot(results.ann_basic, aes(Host_Type, Predicted_Price_USD)) +
  geom_col(fill = "deepskyblue", width = 0.5) +
  geom_text(aes(label = round(Predicted_Price_USD, 2)),
    vjust = 2
  ) +
  labs(
    title = "ANN predicted prices by host type",
    subtitle = "Naive Model",
    x = " ",
    y = "Predicted Price (USD)"
  )

p.ANN_complex <- ggplot(results.df_exp, aes(Host_Type, Predicted_Price_USD)) +
  geom_col(fill = "deepskyblue", width = 0.5) +
  geom_text(aes(label = round(Predicted_Price_USD, 2)),
    vjust = 2
  ) +
  labs(
    title = "ANN predicted prices by host type",
    subtitle = "Complex model",
    x = " ",
    y = "Predicted Price (USD)"
  )

p.ANN_basic | p.ANN_complex
```

## Interpretation ANN

Our initial Neural Network predicted price using only review metrics,
listing capacity, and Superhost status. Surprisingly, it priced
Superhosts lower (**`r results.ann_basic$Predicted_Price_USD[2]` USD**) than regular hosts (**`r results.ann_basic$Predicted_Price_USD[1]` USD**). This was a
clear case of omitted variable bias: Superhosts often use competitive
pricing to maintain the high occupancy required for the badge, while
expensive luxury rentals are usually managed by agencies that don't hold
Superhost status.

We corrected this in a second model by adding room and property types,
using One-Hot Encoding to convert these categories into numerical
inputs. Once we controlled for these structural factors to compare
similar listings, the price gap practically vanished (**`r results.df_exp$Predicted_Price_USD[2]` USD** for
Superhosts vs. **`r results.df_exp$Predicted_Price_USD[1]` USD** for regular hosts). This refutes our initial
hypothesis: when comparing apples to apples, Superhost status alone does
not significantly increase a guest's willingness to pay. Superhosts
often achieve their high status through competitive pricing strategies
designed to maximize occupancy rather than nightly margins. Ultimately,
a guest's willingness to pay is dictated almost entirely by physical
listing attributes like *accommodates*, *room_type*, and *property_type*.
Therefore, we advise hosts to anchor their pricing strictly to their
property's structural features, leveraging their Superhost badge purely
as a marketing tool to capture market share and drive booking volume.

## Limitations

A primary limitation of the `nn_expanded model` is the inherent "black
box" nature of Artificial Neural Networks, which drastically reduces the
interpretability of individual price drivers compared to transparent
linear models. Because neural networks require mathematical
normalization, calculating the exact monetary impact of isolated
variables like *review_scores_rating* is complex and requires continuous
reverse-scaling. Furthermore, while we corrected for some omitted
variable bias by using One-Hot Encoding for property categories, the
model still lacks precise geospatial data, meaning the massive price
impact of specific, highly desirable neighborhoods remains uncaptured.
Testing our hypothesis also relied on creating a mathematically
"average" hypothetical listing, which may not perfectly reflect the
nuanced reality of distinct luxury or budget segments in the Canton of
Vaud. Finally, applying the na.omit() function removed listings with
incomplete profiles, which could introduce bias if those missing data
points systematically represented a specific market segment, such as
poorly managed properties.


# Conclusion

## Scientific Conclusion
### Answering research questions


To answer the *first research question* we have to analyse our answers to both of the hypothesis. 

In terms of host-sided variables (H1) we find that **the Superhost status does not raise the price**. Four models with completely different mechanics, the linear regression, the logistic regression (GLM Binomial), the spline-based GAM, and the neural network, all return the same answer. After controlling for listing size, room type and other structural features, Superhost status has no measurable positive effect on price. Some models even show a small *negative* effect. The GLM Binomial puts a precise number on this: Superhosts have `r sh_pct_lower`% *lower* odds of charging an above-median price. We can conclude that the Superhost strategy is to price competitively and earn through volume.

For the variable *review_scores* the linear model finds a positive effect of about `r round((exp(lm.main$coefficients[2]) - 1) * 100, digits = 2)`%. The GLM Binomial confirms this through a different lens. A one-point increase in review score raises the odds of charging an above-median price by `r rev_pct_up`%. H1 is therefore only partially supported. Quality signals matter, but the Superhost Status itself is not a price lever.

To further validate our findings, we applied Poisson and Quasi-Poisson Generalized Linear Models (GLM). The initial Poisson model revealed significant overdispersion in the data, highlighting the high variance typical for real-world pricing structures. By implementing the Quasi-Poisson model to adjust for this overdispersion and obtain more robust standard errors, we corroborated our primary finding: even under this different model specification, pricing power remains firmly tied to structural and venue-oriented features, while host-sided metrics exert only a marginal influence. The pattern only makes sense once the GLM Poisson is added in. Superhosts trade margin for volume, earning more through occupancy rather than through higher per-night prices.

Turning to the second hypothesis, the GAM gives the clearest verdict on H2. Its 2D spatial spline `s(longitude, latitude)` has an EDF of `r edf_loc`. Lakeshore listings and the Lausanne urban area sit at the top of the price surface. Rural and alpine areas drop off sharply. The non-linearity of the effect is also the strongest single argument we have for using a GAM over a linear model. H2 is clearly supported.

Beyond location, **structural features are the real price drivers**. Capacity (`accommodates`), room type, and the number of bathrooms account for most of the explainable variation in price across all models that include them. The GAM shows that an additional bathroom adds roughly 25% to the price. A private room sells for about 30% less than an entire home, all else equal. Reviews and host attributes matter, but their effect is small compared to the physical features a host can describe in a listing.

A contrast worth highlighting is the Naive ANN versus the Expanded ANN. The naive version, trained only on host variables, predicts a *lower* price for Superhosts (**`r results.ann_basic$Predicted_Price_USD[2]` USD** vs. **`r results.ann_basic$Predicted_Price_USD[1]` USD**). Once room and property type are added, the gap collapses to less than a dollar. This isn't really a disagreement between models but a useful warning against drawing conclusions from raw price comparisons.

The answer the *first research question* is relatively straightforward. We find that virtually all of the variables we examined have an impact on pricing. Interestingly, Superhost status has a negative impact. Furthermore, we find that venue-oriented variables, in particular location and physical features, have the strongest influence.

To answer the *second research question*, we draw on both the GAM and the SVM. The GAM itself explains `r gam_dev_expl`% of the deviance on held-out data and generates a fair-market price for every listing, strong evidence that the price drivers we identified are predictive, not just descriptive. The traffic-light verdicts derived from the GAM feed directly into the SVM, which reaches `r round(svm_accuracy * 100, 2)`% accuracy on the test set. We can predict price fairness, but we also recognize that this accuracy is insufficient to criticize hosts for their pricing based on the variables available to us.


### Overall Limitations

Five limitations apply to this entire study, not just to individual
models:

**Snapshot, no time component**. All models are trained on a single
point in time. Seasonal pricing, weekday effects, and event-driven
demand spikes (Montreux Jazz Festival, ski season, weekends) are nowhere
in the data. The recommendations represent an *average* market price,
not a dynamic one. A real production pricing tool would need a time
dimension we don't have.

**Listed price as a proxy for willingness to pay**. We treat what hosts
charge as a proxy for what guests would pay. This holds on average
across a large sample, but it breaks down for listings that sit unbooked
at their current price, those listings still shape the model and
probably tilt it slightly toward optimistic price levels. A true
willingness-to-pay study would need booking data, completed
transactions, and ideally A/B tests of competing prices.

**Missing predictors.** Photo quality, amenity counts, listing
description sentiment, view from the property, noise levels, distance to
public transport, none of this is in the dataset. Some of the
unexplained variance (around a third for the GAM) lives in these omitted
variables rather than in any model's limitation.

**Reverse causality.** Several relationships in our data probably flow
in both directions. More bookings produce more reviews, which make
Superhost status easier to maintain, which probably increases
visibility, which leads to more bookings. None of our models can
untangle which effect causes which. For descriptive purposes that's
fine, but the recommendations should be read as "what works on average,"
not "what works because of X."

**Vaud-specific.** The dataset covers one Swiss canton at one point in
time. The price levels, location dynamics (lakeshore premium, alpine
resorts), and host behaviours are specific to this market. The
methodology transfers; the exact coefficients do not. Applying the same
models to Geneva or Bern would require refitting on local data.

### Next Steps
When we put the results of the study and the identified limitations into context, we realize that there is plenty of room for future research. We were able to see that a generalized additive model is a very useful approach. However, we have a clear shortcoming, particularly with regard to the limitation of location. Especially if there is a meaningful operationalization of location, for example, based on distances to a city center—pricing can be addressed in even greater detail. This would benefit not only the GAM but all models. However, our study demonstrated what is fundamentally possible. Regarding the second research question and the support vector machine, we also recognize that price classification is both possible and useful. Therefore, we see massive potential for Airbnb in this area in particular, which could be explored in a future study.
Further improvements are also possible if the dataset were expanded to include additional Swiss cantons or other parts of the world. 


## Businesscase conclusion
### Recommendations for Hosts

For hosts setting or revising their nightly price:

-   **Anchor on structure, not status**. What a guest actually pays for
    is the listing itself: room type, capacity, bathrooms, location. The
    Superhost badge helps with visibility and booking volume, not with
    the per-night rate.
-   **Use the traffic-light tool as a sanity check, not a target**. The
    GAM-based recommendation gives a fair-market range based on what
    comparable listings achieve. A listing flagged "Underpriced" is
    leaving money on the table; one flagged "Overpriced" risks sitting
    empty.
-   **If you're building toward Superhost status, compete on price**.
    The 2.6× occupancy boost comes from satisfied guests and a full
    calendar. Premium pricing makes both harder. Start with competitive
    pricing, accumulate reviews, then raise prices selectively once the
    badge is secure.
-   **Review scores matter less than you think for price**. Across the
    pricing models the reward for higher ratings is modest. Reviews
    matter, but their role is to drive *bookings*.

### Recommendations for Airbnb

For the platform itself, the same data points in three directions:

-   **A pricing assistant should preferably use the GAM, not a linear formula**.
    Linear or postcode-based tools will systematically underprice
    lakeshore listings and overprice rural ones. The 2D spatial spline
    is the most defensible way to capture local dynamics with the data
    Airbnb already collects.
-   **Coach the "Underpriced" segment**. Roughly 41% of listings in the
    dataset sit below the GAM's fair-price band. These hosts are the
    easiest client group to help. A recommendation alone could shift their
    pricing toward market value, lifting both host revenue and platform
    commission.
-   **Add operationalised location features**. Raw coordinates work well
    for a smoother like the GAM but limit simpler tools. Distance to the
    nearest lake, city centre, ski lift, and public transport stop would
    let classifiers like the SVM catch up to the GAM's accuracy — and
    these features are trivial to compute centrally and serve to every
    model on the platform.




# Lessons learned

## Pascal

Working on this project was a huge reality check in the best way
possible. In our classes, we are used to getting "perfect" datasets that
are practically made for the specific algorithms we are learning. But
working with our actual `d.airbnb_listings_feat` dataset showed me how
much effort actually goes into data cleaning and prepping before you can
even start modeling. I personally focused on building the GLM-Poisson
and ANN models, and trying to get them to perform well on real, messy
data required a lot more critical thinking and trial-and-error than our
standard assignments in class.

Since we split up the workload, I didn't code every single model myself.
Instead, my role for the LM, GLM-Binomial, GAM, and SVM models was to
check and review the work of my team members. This turned out to be a
super valuable learning experience on its own. Reading through their
code allowed me to see exactly how other people translate what we
learned in class into practice. Seeing their different approaches to
problem-solving and troubleshooting bugs really helped expand my own
coding mindset.

Finally, figuring out how to collaborate effectively on code was a
massive win for us. We used GitHub with RStudio for the first time, and
it honestly made a world of difference. We finally found a setup that
actually works for coding together as a group, which saved us a lot of
headaches and is definitely a workflow I’ll be taking into future
projects.

## Timon

I was already familiar with RStudio from my bachelor’s program, and I’d
had very positive experiences with it so far and I feel very comfortable
using it. Up until now, however, I’ve mainly worked with simple linear
models and generalized linear models. Everything else was new to me. In
this project, I focused primarily on support vector machines, and I’m
very impressed by the concept because it’s both very simple and yet
involves a tremendous amount of complexity. Working with this model in
particular was a key learning experience for me.

Another key learning was the SCRUM-like structuring of the work process.
Within the document, we work with a list of tasks to be completed by the
next meeting and a backlog of tasks that we can’t tackle right now but
know are coming up. This worked very well for us. It provided a sense of
accountability to ensure all tasks were completed effectively.

Finally, working with GitHub on this scale was also fairly new to me. We
used the integration with RStudio for the first time, and I’m personally
thrilled by how simple it is. We had almost no issues with incorrect
versions, even though we weren’t working with separate branches. I
think, however, that splitting up the models really helped us in that
regard.

## Davide

Coming into this project, my comfort zone was firmly in theory. I knew
what a GLM was on paper, but actually building one that holds up against
messy data was a different experience. Working on the GLM Binomial and GAM
forced me to confront the gap between understanding a model and making
it useful. Things that look elegant in a lecture slide turn out to
require dozens of small judgement calls in practice. Questions like which variables to include, how to handle NAs, when to log-transform, where to draw a
reasonable threshold and much more.

The GAM was probably my biggest single learning. I did not worked a lot with
smooth terms before, and seeing the 2D spatial spline pick up Vaud's
geography was genuinely a moment where the theory clicked into practice.

I went into this project assuming the modelling would be the hard part. By the end, communicating the results clearly turned out to be just as much work.

Finally, R Bootcamp in February laid the foundation, but this project
was where I actually got comfortable in R. Reading the others' code,
debugging together, learning shortcuts and much more. That side-by-side learning was
something I can't get from several tutorials.


# Use of GenAI

Throughout this machine learning project, we leveraged various
Generative AI tools specifically ChatGPT, Gemini, and Claude to support
our analytical workflow. A primary use case was troubleshooting complex
issues that extended beyond our existing knowledge base. When
encountering technical bottlenecks, we used these models to break down
and explain the underlying concepts, enabling us to systematically find
the right solutions. Furthermore, we utilized GenAI as a critical review
mechanism for our modeling approach. We consulted the AI to verify
whether we had made the appropriate theoretical assumptions, such as
selecting the correct data distributions for our models. It also served
as a valuable sounding board to double-check our final analytical
interpretations, ensuring our conclusions were statistically sound.
Finally, these tools provided essential linguistic support. We actively
used GenAI to translate our documentation and findings into professional
English, ensuring a high-quality final report.
