---
title: "Main Group Work_ML"
author: "Pascal Ingold, Timon Reichmuth und Davide Gualdani"
date: "2026-03-22"
output:
  html_document:
    df_print: paged
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, comment = NA)

```

Next Steps: 


Pascal: 
- a Generalised Linear Model with family set to Poisson --> glm(Price ~ reviews) Achtung Price kann nicht mit Poisson gerechnet werden! Sind keine Zähldaten. Neu Auslastung genutzt, um Zähldaten zu haben. -> DONE
- NA Handling untersuchen -> DONE
- Neural Network --> Ev. auch Kategorisierung analog SVM --> DONE




Timon: 
- a Linear Model --> lm(price ~ rewview_rating + Superhost T/F) --> für unterscheidung und kontrolle. -> INPROGRESS


- NA Handling untersuchen
  |-> NA vorwiegend bei ratings und rates vor. Verdacht: wenn nie eine Vermietung gemacht wurde erhalten wir einen NA in den Ratings
  TODO: schauen ob wenn wir mit den Variablen rechnen, NA's einen negativen Einfluss haben. Wie zählen NA in ein Modell hinein?
        --> beim lm keinen Einfluss, da diese Zeilen einfach weggelassen werden. 
        --> bei anderen Modellen muss aufgepasst werden. 
        
        
– a Support Vector Machine --> Overpriced & Underpriced & Right Price (Davide und Timon)


Davide:
- – a Generalised Linear Model with family set to Binomial --> glm(Price ~ Superhost T/F) -> DONE
– a Generalised Additive Model --> gam.main <- gam(price ~  --> allenfalls aufbauend auf dem lm()
                                    s(accommodates) +
                                    s(review_scores_rating) +
                                    s(number_of_reviews) +
                                    s(longitude, latitude) +
                                    room_type +
                                    host_is_superhost +
                                    bathrooms,
                                  family = gaussian(link = "log"),
                                  data   = d.airbnb_listings_feat_gam)



Tasks ab 15.4:

Timon: 

zeitkritisch: fixieren des zweiten svm- Modells für DAVIDE
- Kommentare bei lm erweitern uns verbessern. (allenfalls translate)
- ausarbeiten des svm-Models. 
- schreiben des interpretationstextes lm
- schreiben des interpretationstextes svm


Pascal:
- ausarbeiten des ANN und vereinfachung der Grafik? (Wenn möglich)
- Pilot mit Auslastung analog zu poisson. wenn klar wird dsas auslastung besser ist als der PReis, können wir die Analyse noch shiften. 

Davide:
- nach GO von TIMON anpassen der Traffic-Lights-Benchmarks so dass das svm besser funktioniert. 
- Blick über lm und poisson


ALLE:
Limitationen im Kapitel unten ergänzen. jeweils für die eigenen Modelle. 











Infos zu bedenken: 
ABGABEDATUM: 12 Juni

- Outlier Handling definieren.
--> falsche Datenpunkte sollten gelöscht werden. Extremwerte im Zweifel drin lassen. Ansonsten kann das als p-hacking verstanden werden. 

Coding Conventions. 
• data sets are named with the prefix “d.” (e.g., d.cats).
• linear model objects are named with the prefix “lm.” (e.g., lm.cats)
• within the text R-objects are written using code formatting (e.g., lm.cats)
14
• variable names are written using italic (e.g., Body.weight)
• these conventions are arbitrary, but hopefully useful to the reader




# Research Questions

What factors, which can be modified by Hosts on AirBnb, affect guest willingness to pay more for a certain room?

What are factors that influence the guest willingness to pay (WTP)

What are factors on which the guest can adjust their pricing model?

## Concepts: Willingness to pay (WTP)

WTP is the hypothetical value, how much a customer is willing to pay for a location or a room for a certain standard. This can rise, if for example the host provides the ability to wash clothes in the apartment. So for the same location, once with and once without a washing machine, the customers WTP might increase with the washer. In a more abstract example, if the host is a super host, the customer might trust the host more and his WTP is higher. 

## Hypothesis
H1: If the review-score of a host is better, the WTP of a guest increases.
H2: If a host has more reviews, the WTP of a guest increases. 
H3: If a host is e superhost, the WTP of a guest increases. 

WTP is measured in the Price displayed. We assume, that the mean of prices is similar to the WTP and residuals even out. Therefore we can work with the effective prices displayed on AirBnb. 



# Business Case
We propose a data-driven pricing decision-support system for Airbnb hosts in the
Canton of Vaud, providing market-aligned price recommendations based on listing
characteristics. The tool targets both new and existing hosts, using inputs such
as location, room type, guest capacity, availability, ratings, and superhost status.
Machine learning models, including LM, GLM, GAM, SVM, and Neural
Networks, will be applied to identify key price drivers and generate accurate
predictions. The objective is to balance revenue and occupancy, helping hosts
understand their market value.
A key feature is interpretability, enabling users to see why a certain price is
recommended. From a business perspective, the system aims to increase booking
rates, reduce price dispersion, and improve host satisfaction.
The model is tailored to the Canton of Vaud to capture regional economic effects.


# Editing of Dataset
First step: Loading Packages

```{r Libraries, results = "hide", echo=FALSE}
library(tidyverse)
library(janitor)
library(mgcv)
library(e1071)
library(nnet)
library(pROC)
library(Metrics)
library(performance)
library(broom)
library(stringr)
library(ggplot2)
library(patchwork)
library(knitr)
library(performance)
library(naniar)
library(visdat)
library(neuralnet)
library(leaflet)
library(leaflet.extras)

```



Second step: Reading all available Datasets
```{r reading Dataset, echo=FALSE}
path_listings <- "./Data_Vaud_all/listings-2.csv"
d.raw_data_listings <- read.csv(path_listings, header = TRUE)
```

Cleaning the Airbnb-Listings Data. We keep only the variables needed for our project. We reduce the amount of variables to exact 20, according to the requirements by our lecturers.

```{r cleaning Dataset, echo=FALSE}
d.airbnb_listings <- d.raw_data_listings %>%
  clean_names()

d.listings_clean <- d.airbnb_listings %>%
  mutate(
    price = str_remove_all(price, "\\$|,"),
    price = as.numeric(price),
    host_is_superhost = ifelse(host_is_superhost %in% c("t", "TRUE", "1"), 1, 0),
    room_type = as.factor(room_type),
    property_type = as.factor(property_type)
  ) %>%
  filter(!is.na(price), price > 0)

d.listings_clean <- d.listings_clean %>%
  select(-id, -host_id, -host_listings_count, -host_total_listings_count, -listing_url, -scrape_id, -last_scraped,
         -source, -name, -description, -maximum_nights, -first_review, -last_review, -calculated_host_listings_count, -host_since, -availability_365, -estimated_revenue_l365d, 
         -neighborhood_overview, -picture_url, -host_url, -host_name, -host_location,
         -host_about, -host_thumbnail_url, -host_picture_url, -host_neighbourhood,
         -neighbourhood, -neighbourhood_cleansed, -bathrooms_text, -amenities, 
         -minimum_minimum_nights, -maximum_minimum_nights, -minimum_maximum_nights,
         -maximum_maximum_nights, -minimum_nights_avg_ntm, -maximum_nights_avg_ntm,
         -calendar_updated, -has_availability, -availability_30, -availability_60,
         -availability_90, -calendar_last_scraped, -number_of_reviews_ltm, -number_of_reviews_l30d,
         -availability_eoy, -number_of_reviews_ly, -review_scores_accuracy, -review_scores_cleanliness,
         -review_scores_checkin, -review_scores_communication, -review_scores_location,
         -license, -instant_bookable, -calculated_host_listings_count_entire_homes, -calculated_host_listings_count_private_rooms,
         -calculated_host_listings_count_shared_rooms, -reviews_per_month, -bedrooms, -review_scores_value)
```

This is our final, cleaned Dataset.
```{r final Call cleaned Dataset, echo=TRUE}
str(d.listings_clean)
```


# Feature Engineering

First Feature Engineering for the airbnb_listings data

```{r FE for EDA (Pascal), echo=FALSE}

#Für Model XXXX 
d.airbnb_listings_feat <- d.listings_clean %>%
  mutate(
    log_price = log(price),
    price_per_guest = price / pmax(accommodates, 1),
    high_demand = ifelse(estimated_occupancy_l365d > 200, 1, 0)
  )
```


```{r FE for lm (TIMON)}

#removing % from response rate and make it binomial
d.airbnb_listings_feat$host_response_rate <- as.numeric(sub("%", "", d.airbnb_listings_feat$host_response_rate)) / 100 

#removing % from response rate and make it binomial
d.airbnb_listings_feat$host_acceptance_rate <- as.numeric(sub("%", "", d.airbnb_listings_feat$host_acceptance_rate)) / 100 

#change the values of host_is_superhost to TRUE/False
d.airbnb_listings_feat$host_is_superhost <- ifelse(d.airbnb_listings_feat$host_is_superhost==1,
                                                   TRUE, FALSE
                                                   )


```


```{r FE for SVM (TIMON)}
d.airbnb_listings_feat$estimated_revenue_365 <- d.airbnb_listings_feat$price * d.airbnb_listings_feat$estimated_occupancy_l365d

summary(d.airbnb_listings_feat$estimated_revenue_365)
hist(d.airbnb_listings_feat$estimated_revenue_365)
print(quantile(d.airbnb_listings_feat$estimated_revenue_365))


```
wie gehen wir bei diesen Variablen mit den NA um?

# Data Quality

Wir prüfen ob und wieviele NA's vorhanden sind. Hierzu nutzen wir naniar und visdat, um die NA's auch visualisieren zu können. Die meisten NA-Werte sind bei den Rates, daher wird von Pascal angenommen, dass die Nullen nicht importiert wurden und deshalb es zu dieser grossen Anzahl NAs kam. Ich würde hier die NA durch 0 ersetzen.
Entscheid: Da Ratings fehlen werden sie in Modellen omitted. Daten fehlen aus einem Grund und nicht random. 

```{r}
vis_miss(d.airbnb_listings_feat)
gg_miss_var(d.airbnb_listings_feat)
gg_miss_upset(d.airbnb_listings_feat)
```


#GLM set to binomial(Davide)
1. Schritt: Zielvariable binarisieren
Aus Preis eine 0/1-Variable machen
Der Median ist der sauberste Cutpoint — er ist robust gegenüber Ausreissern und teilt den Datensatz exakt in zwei Hälften
```{r FE for glm binomial (DAVIDE)}
d.airbnb_listings_feat <- d.airbnb_listings_feat %>%
  mutate(price_high = as.integer(price > median(price, na.rm = TRUE)))

# Checken ob Split sauber ist
table(d.airbnb_listings_feat$price_high)

```


```{r FE for glm poisson (PASCAL)}


```




```{r final call of FE-dataset}
str(d.airbnb_listings_feat)
```


# Exploratory Data Analysis

```{r}
hist(d.airbnb_listings_feat$price)
hist(d.airbnb_listings_feat$log_price)
ggplot(data= d.airbnb_listings_feat, aes(x = room_type, y = price)) +
  geom_boxplot() +
  scale_y_log10() +
  labs(
    title = "Preisverteilung nach Room Type (log-Skala)",
    x = "Room Type",
    y = "Log_Preis"
  ) +
  theme_minimal()

ggplot(data = d.airbnb_listings_feat, aes(x = accommodates, y = log_price)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "loess", se = FALSE) +
  labs(
    title = "Preis vs. Kapazität (log-Skala)",
    x = "Anzahl Gäste",
    y = "Log_Preis"
  ) +
  theme_minimal()
  
ggplot(data = d.airbnb_listings_feat, aes(x = log_price, y = number_of_reviews)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "loess", se = FALSE) +
  labs(
    title = "Anzahl Reviews vs. Preis",
    x = "Log_Preis",
    y = "Anzahl Reviews"
  ) +
  theme_minimal()

ggplot(data = d.airbnb_listings_feat, aes(x = longitude, y = latitude)) +
  geom_point(aes(color = log_price), alpha = 0.6) +
  scale_color_viridis_c() +
  labs(
    title = "Airbnb Listings (Farbe = Log-Preis)",
    x = "Longitude",
    y = "Latitude",
    color = "Log_Preis"
  ) +
  theme_minimal()

ggplot(data = d.airbnb_listings_feat, aes(x = longitude, y = latitude)) +
  geom_point(aes(color = log_price, size = number_of_reviews), alpha = 0.5) +
  scale_color_viridis_c() +
  labs(
    title = "Listings: Preis und Nachfrage",
    x = "Longitude",
    y = "Latitude",
    color = "Log_Preis",
    size = "Reviews"
  ) +
  theme_minimal()
```

*Important to mention*
During our exploratory data analysis, we identified one 'Private room' listing with an unrealistically high price of 2'147'483'647 CHF. We excluded this single observation from our dataset as it is an extreme outlier that does not reflect the true market conditions and would negatively skew our predictive models.
```{r}
# This keeps everything EXCEPT the row that is a Private room AND costs more than 5000
d.airbnb_listings_feat <- d.airbnb_listings_feat %>%
  filter(!(room_type == "Private room" & price > 100000))
```


# EDA (DAVIDE)
```{r}
# --- 1. Distribution of the Target Variable (Price) ---
# Upgrading base R hist() to ggplot2 for a more professional look.
# Showing both raw and log side-by-side helps visually justify the transformation to the client.
p_raw <- ggplot(d.airbnb_listings_feat, aes(x = price)) +
  geom_histogram(bins = 50, fill = "steelblue", color = "white") +
  labs(title = "Raw Price Distribution", x = "Price", y = "Count") +
  theme_minimal()

p_log <- ggplot(d.airbnb_listings_feat, aes(x = log_price)) +
  geom_histogram(bins = 50, fill = "darkgreen", color = "white") +
  labs(title = "Log-transformed Price Distribution", x = "Log(Price)", y = "Count") +
  theme_minimal()
p_raw + p_log # Displays them side-by-side

# --- 2. Categorical Predictor: Room Type vs. Price ---
# Your original code was good, just cleaned up the aesthetics.
ggplot(data = d.airbnb_listings_feat, aes(x = room_type, y = log_price, fill = room_type)) +
  geom_boxplot(alpha = 0.7) +
  labs(
    title = "Price Distribution by Room Type",
    x = "Room Type",
    y = "Log(Price)"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

# --- 3. Continuous Predictor: Capacity vs. Price ---
ggplot(data = d.airbnb_listings_feat, aes(x = accommodates, y = log_price)) +
  geom_point(alpha = 0.3, color = "darkgray") +
  geom_smooth(method = "loess", color = "blue", se = FALSE) +
  labs(
    title = "Price vs. Capacity",
    subtitle = "Smoother applied to identify potential non-linear trends",
    x = "Number of Guests (Accommodates)",
    y = "Log(Price)"
  ) +
  theme_minimal()

# --- 4. Continuous Predictor: Reviews vs. Price ---
# Correction: The response variable (log_price) should generally be on the y-axis 
# when plotting against a predictor[cite: 255].
ggplot(data = d.airbnb_listings_feat, aes(x = number_of_reviews, y = log_price)) +
  geom_point(alpha = 0.3, color = "darkgray") +
  geom_smooth(method = "loess", color = "red", se = FALSE) +
  labs(
    title = "Does the Number of Reviews impact Price?",
    x = "Total Number of Reviews",
    y = "Log(Price)"
  ) +
  theme_minimal()

# --- 5. Binary Predictor: Superhost Status ---
# Adding this because it's a key part of your business case and fulfills the binary variable requirement.
ggplot(data = d.airbnb_listings_feat, aes(x = as.factor(host_is_superhost), y = log_price, fill = as.factor(host_is_superhost))) +
  geom_boxplot(alpha = 0.7) +
  scale_x_discrete(labels = c("0" = "Regular Host", "1" = "Superhost")) +
  labs(
    title = "Price Difference: Superhosts vs. Regular Hosts",
    x = "Host Status",
    y = "Log(Price)"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

# --- 6. Spatial Data: Mapping the Listings ---
# Cleaned up the map to focus purely on location and price.
ggplot(data = d.airbnb_listings_feat, aes(x = longitude, y = latitude)) +
  geom_point(aes(color = log_price), alpha = 0.6, size = 1.5) +
  scale_color_viridis_c(option = "plasma") +
  labs(
    title = "Airbnb Listings in Vaud by Price",
    x = "Longitude",
    y = "Latitude",
    color = "Log(Price)"
  ) +
  theme_minimal()
```
# Davide's overall Data Analysis
## 1. Overall Market Overview (Prices & Capacity)
This is the absolute minimum, maximum, median, and mean for the most important numeric variables.
```{r}
# Basic summary statistics for key numerical variables
summary_stats <- d.airbnb_listings_feat %>%
  select(price, log_price, accommodates, number_of_reviews) %>%
  summary()

print(summary_stats)
```
Client Interpretation: "The average Airbnb in Vaud costs X CHF per night, but the median is Y CHF, indicating that a few extremely expensive luxury listings might be skewing the average upward. Prices range from a minimum of A CHF to a maximum of B CHF."

## 2. Analysis by Room Type
Since we want to know how many different types of rooms there are, we group the data and calculate the counts, percentages, and price ranges for each type.
```{r}
# Detailed breakdown by Room Type
room_type_analysis <- d.airbnb_listings_feat %>%
  group_by(room_type) %>%
  summarise(
    Total_Listings = n(),
    Percentage_of_Market = round((n() / nrow(d.airbnb_listings_feat)) * 100, 1),
    Min_Price = min(price, na.rm = TRUE),
    Median_Price = median(price, na.rm = TRUE),
    Average_Price = round(mean(price, na.rm = TRUE), 1),
    Max_Price = max(price, na.rm = TRUE)
  ) %>%
  arrange(desc(Total_Listings)) # Sorts by the most common room type

print(room_type_analysis)
```
Client Interpretation: "Entire homes/apartments dominate the Vaud market, making up 75% of all listings. As expected, the Hotel room command the highest median price (264.5 CHF), compared to shared rooms (47 CHF)."

## 3. Superhost vs. Regular Host Breakdown
This business case specifically mentions superhost status. So we took a look to see how many there are and if they actually charge more or even less.
```{r}
# Superhost statistics
superhost_analysis <- d.airbnb_listings_feat %>%
  group_by(host_is_superhost) %>%
  summarise(
    Total_Hosts = n(),
    Percentage = round((n() / nrow(d.airbnb_listings_feat)) * 100, 1),
    Median_Price = median(price, na.rm = TRUE),
    Avg_Reviews = round(mean(number_of_reviews, na.rm = TRUE), 1)
  ) %>%
  mutate(host_is_superhost = ifelse(host_is_superhost == 1, "Yes", "No"))

print(superhost_analysis)
```


## 1. Correlation Matrix (Crucial before Machine Learning)
Before we train our LM, GAM, or SVM, we need to know how our numerical variables relate to each other. This helps us identify multicollinearity (when two predictors provide the same information) and strong price drivers.

```{r}
# Select relevant numerical variables (drop NAs just for the correlation calculation)
cor_data <- d.airbnb_listings_feat %>%
  select(log_price, accommodates, number_of_reviews) %>%
  drop_na()

# Calculate the correlation matrix
cor_matrix <- cor(cor_data)

# Render a nice table for your RMarkdown document
kable(round(cor_matrix, 2), caption = "Correlation Matrix of Key Numerical Drivers")
```
Client Interpretation: "As expected, accommodates shows the strongest positive correlation with price. It is also interesting to note the relationship between availability and number_of_reviews..."

## 2. Detailed property_type Analysis
We already analyzed room_type, but Airbnb often has dozens of property_types (e.g., "Entire chalet", "Private room in castle", etc.). Summarizing the Top 10 property types in the Canton of Vaud provides maybe a great market insight.
```{r}
# Top 10 Property Types by frequency and their median price
top_properties <- d.airbnb_listings_feat %>%
  group_by(property_type) %>%
  summarise(
    Total_Listings = n(),
    Median_Price = median(price, na.rm = TRUE)
  ) %>%
  arrange(desc(Total_Listings)) %>%
  slice_head(n = 10) # Keep only the top 10

kable(top_properties, caption = "Top 10 Property Types in the Canton of Vaud")
```

Client Interpretation: "While standard apartments like "entire rental unit" are the most common, specialty properties like chalets, homes and villas command a significant premium, which our model will need to capture."

## 3. Missing Values (NA) Check
This step is less for the client and essential for your ML models. Most modeling functions (like lm() or svm()) will throw errors or silently drop entire rows if NAs are present. You need to know exactly where your data is missing!

```{r}
# Count the number of NAs for each key column in your dataset
na_summary <- d.airbnb_listings_feat %>%
  select(price, log_price, accommodates, number_of_reviews, host_is_superhost) %>%
  summarise(across(everything(), ~ sum(is.na(.)))) %>%
  t() %>% # Transpose for better readability
  as.data.frame() %>%
  rename(Missing_Values = V1)

kable(na_summary, caption = "Count of Missing Values per Variable")
```



#LM (TIMON - Davide translated to english and commments)


```{r lm controlled by Timon}
# --- LM controlled by Timon ---
# Fitting a Multiple Linear Regression model to predict the log-transformed price.
# We use log(price) to account for the right-skewed distribution of Airbnb prices.
lm.main <- lm(log(price) ~ review_scores_rating + as.factor(host_is_superhost) + 
              as.factor(room_type) + accommodates, data = d.airbnb_listings_feat)

summary(lm.main)

# Checking for multicollinearity using Variance Inflation Factor (VIF).
# Values range from 1 to infinity. A value of exactly 1 indicates no collinearity.
# Values > 5-10 suggest problematic multicollinearity among the predictor variables.
multicollinearity(lm.main)
```
# LM Controlled Plots (TIMON - Davide translated to english and commments)
```{r}
# --- LM controlled plots ---
# Visualizing the raw relationship between review scores, log(price), and superhost status.
ggplot(data = d.airbnb_listings_feat, mapping = aes(review_scores_rating, log(price), color = host_is_superhost)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "lm")

# Creating a grid of hypothetical data to predict and plot the fitted LM lines cleanly.
# We hold 'room_type' and 'accommodates' constant to isolate the specific effects of reviews and superhost status.
d.lm_predict <- expand.grid(
  review_scores_rating = seq(3, 5, length.out = 100),
  host_is_superhost = c("TRUE", "FALSE"),
  room_type = "Entire home/apt",   # Fixed reference category
  accommodates = mean(d.airbnb_listings_feat$accommodates, na.rm = TRUE) # Fixed at mean value
)

# Generating predictions based on the trained LM model
d.lm_predict$pred <- predict(lm.main, newdata = d.lm_predict)

# Plotting the actual data points alongside our model's predicted regression lines
ggplot(d.airbnb_listings_feat, 
       aes(review_scores_rating, log(price), color = host_is_superhost)) +
  geom_point(alpha = 0.3) +
  geom_line(data = d.lm_predict, aes(y = pred), size = 1.2) +
  labs(title = "LM Predictions: Log(Price) vs. Review Rating",
       x = "Review Scores Rating",
       y = "Log(Price)")
```



# GLM set to poisson (PASCAL - Davide translated to english and commments)
```{r}
# --- GLM set to Poisson (Pascal) ---
# Fitting a Poisson model to predict 'estimated_occupancy_l365d' (which represents count data).
# Note: Price cannot be used as a response in a Poisson model as it is continuous, not discrete counts.

glm_poisson <- glm(estimated_occupancy_l365d ~ price + host_is_superhost + 
                     review_scores_rating + accommodates + room_type, 
                   data = d.airbnb_listings_feat, 
                   family = "poisson")

summary(glm_poisson)

# Checking for Overdispersion.
# A strict assumption of the Poisson distribution is that the mean exactly equals the variance.
# Since Airbnb bookings fluctuate heavily (many cheap listings, few very expensive ones), overdispersion is expected.
dispersion_ratio <- glm_poisson$deviance / glm_poisson$df.residual
print(paste("Dispersion Ratio:", dispersion_ratio))

# The dispersion ratio is significantly larger than 1, confirming overdispersion.
# To correct this, we switch to a Quasi-Poisson model. This properly scales up the standard errors
# to prevent the model from identifying false-positive (overly significant) variables.
glm_quasipoisson <- glm(estimated_occupancy_l365d ~ price + host_is_superhost + 
                          review_scores_rating + accommodates + room_type, 
                        data = d.airbnb_listings_feat, 
                        family = "quasipoisson")

summary(glm_quasipoisson)
```
## Interpretation 
### GLM-Poisson:

Das GLM mit Poissonverteilung hat wie erwartet eine Überdispersion produziert, da die Streuung der Airbnb-Daten zu erwarten war. Das Modell hat die Standardfehler zu klein berechnet, wweshalb im ersten Output alle Variablen hochsignifikant waren. Danach wurde die Dispersion berechnet und sie war grösser als 1, nämlich 56.53. Als Lösung wurde ein Quasipoissonmodell genutzt, welches die Standardfehler nach oben anpasst.

### GLM-Quasipoisson
Das GLM mit Quasipoissonverteilung hat die Überdispersion korrigiert. Die Standardfehler wurden nach oben angepasst (deutlich) und gibt nun ein realistisches Bild, welche Faktoren wirklich einen Einfluss auf die Auslastung haben.
Interpretationen:
- Price -> hochsignifikant negativ. Wie erwartet: Je höher der Preis, desto geringer die Auslastung
- Superhost: hochsignifikant positiv: Achtung, logarithmierte Raten, darum Effekt mit Exponentialfunktion berechnen exp(0.9577) = 2.6. Ein Superhost zu sein steigert die Auslastung um das 2,6-fache im Vergleich zum normalen Host.
- Bewertungen: Positiv und signifikant. Bessere Bewertungen führen zu mehr Buchungen.
- Kapazität: Positiv und signifikant. Je mehr Leute reinpassen, desto mehr Buchungen.
- Zimmertypen: Im Intercept ist "Entire home/apt". Private Room ist signifikant schlechter gebucht als ganze Unterkunft. Shared Room hat hohen P-Wert. Statistisch kann man nicht sagen, ob sich die Auslastung eines geteilten Zimmers von der ganzen Wochnung unterscheidet.

## GLM Quasi-Poisson Ergebnisse für Businesscase

```{r}
# --- GLM Quasi-Poisson Results Extraction (Pascal) ---
# Extract model results into a clean dataframe using broom::tidy()
d.results_glm_quasipoisson <- tidy(glm_quasipoisson) %>%
  # Calculate the percentage change for easier business interpretation.
  # Since Poisson models use a log-link function, we must exponentiate the coefficients (estimates) 
  # to interpret them as multiplicative effects on the expected count (occupancy).
  mutate(
    pct_change = (exp(estimate) - 1) * 100,
    # Round the percentage change to 2 decimal places for readability
    pct_change = round(pct_change, 2)
  ) %>%
  # Select only the columns most relevant for communicating insights to the client
  select(term, estimate, pct_change, p.value)

print(d.results_glm_quasipoisson)
```



# GLM Binomial (DAVIDE)

```{r}
# --- GLM Binomial (DAVIDE) ---
# Fitting a Logistic Regression (Binomial GLM) to predict if a listing is priced above the market median.
# The target variable 'price_high' is binary (1 = above median price, 0 = below median price).
glm_binomial <- glm(price_high ~ host_is_superhost,
                    data   = d.airbnb_listings_feat,
                    family = binomial(link = "logit"))

summary(glm_binomial)
```
# Interpretation GLM Binomial: Calculating Odds Ratios (DAVIDE)
Similar to the Poisson model, the coefficients in a binomial model with a logit link are log-odds.

```{r}
# --- Interpretation GLM Binomial: Calculating Odds Ratios ---
# Similar to the Poisson model, the coefficients in a binomial model with a logit link are log-odds.
# We exponentiate the estimates (exp) to convert them into easily interpretable Odds Ratios.
d.results_glm_binomial <- tidy(glm_binomial) %>%
  mutate(
    odds_ratio = round(exp(estimate), 3),
    p.value    = round(p.value, 4)
  ) %>%
  select(term, estimate, odds_ratio, p.value)

print(d.results_glm_binomial)
```

What the Output Tells Us:
The coefficient for host_is_superhost is negative (-0.2217) and highly significant (p = 0.002). The calculated Odds Ratio of 0.801 means that Superhosts have roughly 20% lower odds of pricing their listings above the median market price compared to regular hosts. Consequently, Superhosts tend to be cheaper, not more expensive.

At first glance, this seems to contradict our Hypothesis 3 (H3: "If a host is a superhost, the WTP of a guest increases / they charge more"). However, the real business insight emerges when we combine this finding with Pascal's Quasi-Poisson model.

The Superhost Strategy: Volume over Margin
Pascal's model demonstrated that having Superhost status increases occupancy by a factor of 2.6. The Binomial GLM shows that Superhosts are more likely to price below the median. Together, this creates a coherent picture of the Superhost business strategy:

  - Superhost = Lower Price + Significantly Higher Occupancy

  - Regular Host = Higher Price + Significantly Lower Bookings

Superhosts actively optimize for volume rather than margin. This makes perfect business sense: to achieve and maintain Superhost status on Airbnb, hosts need a high volume of bookings and a consistent stream of positive reviews. This is much easier to achieve with highly competitive pricing.

Therefore, while H3 is not confirmed in the sense that "Superhosts charge a premium," the models reveal the opposite, which is a much stronger strategic finding: Superhosts price strategically lower but achieve massively more bookings as a result. 

Recommendation for Hosts:
A host aspiring to reach Superhost status should not increase their prices. Instead, they should use competitive pricing to gather bookings and reviews. The financial reward of the Superhost status pays off indirectly through high occupancy rates, rather than a higher nightly single-booking price.

# GAM Modell (DAVIDE)

```{r}
# --- GAM Model ---
# Select only the necessary variables and remove NAs
# GAM cannot handle NA's in the dataframe → na.omit() is necessary
d.airbnb_listings_feat_gam <- d.airbnb_listings_feat %>%
  select(price, accommodates, review_scores_rating,
         number_of_reviews, longitude, latitude,
         room_type, host_is_superhost, bathrooms) %>%
  na.omit()

cat("Observations after na.omit():", nrow(d.airbnb_listings_feat_gam), "\n")
```

```{r}
# Fitting the GAM Model
gam.main <- gam(price ~
                  s(accommodates) +
                  s(review_scores_rating) +
                  s(number_of_reviews) +
                  s(longitude, latitude) +
                  room_type +
                  host_is_superhost +
                  bathrooms,
                family = gaussian(link = "log"),
                data   = d.airbnb_listings_feat_gam)

summary(gam.main)
```

Model Overview
The Generalised Additive Model (GAM) was fitted with a Gaussian family and log link, using 3,160 observations after listwise deletion of missing values. The model explains 52.9% of deviance (adj. R² = 0.521), which is considered strong performance for hedonic price modelling in the short-term rental market.

Parametric Terms (Linear Effects)
Room Type is the strongest categorical price driver in the model. Using Entire home/apt as the reference category:

Hotel room: significant positive premium (β = 0.720, p < 0.001). Translating from the log scale: exp(0.720) ≈ 2.05, meaning hotel rooms are associated with roughly double the price of comparable entire homes.
Private room: significant negative effect (β = −0.363, p < 0.001), corresponding to approximately a 30% price reduction (exp(−0.363) = 0.696).
Shared room: not statistically significant (p = 0.180), likely due to the small number of shared room listings in the Canton of Vaud.

Bathrooms show a highly significant positive effect (β = 0.181, p < 0.001). Each additional bathroom is associated with approximately a 20% price increase (exp(0.181) = 1.198), reflecting the importance of physical listing characteristics for guest willingness to pay.
Superhost status shows a small negative coefficient (β = −0.049, p = 0.078), which is only marginally significant at the 10% level. This finding is consistent across all three models estimated in this study — the GLM Binomial, the Neural Network, and the GAM all converge on the same conclusion: superhost status does not yield a measurable price premium. This is discussed in detail in the cross-model comparison below.

Smooth Terms (Non-linear Effects)
All four smooth terms are highly significant (p < 2e-16) and show substantial non-linearity, as indicated by their estimated degrees of freedom (EDF)

An EDF close to 1 would indicate a near-linear relationship, making a spline unnecessary. All EDF values here are far above 1, confirming that non-linear smoothing is statistically justified for every continuous predictor in the model. A standard linear model would have failed to capture these relationships adequately.
The *2D spatial spline* s(longitude, latitude) deserves particular attention with an EDF of 27.75. This indicates a highly complex spatial price surface across the Canton of Vaud — locations closer to Lake Geneva and urban centres such as Lausanne command significantly higher prices, while rural and alpine areas show distinct pricing patterns. This spatial component alone justifies the GAM approach for a regionally scoped pricing tool.

Cross-Model Comparison: Superhost Effect
One of the most consistent findings across all models in this study is the absence of a superhost price premium
This convergence across fundamentally different model architectures — parametric, non-parametric, and connectionist — constitutes strong evidence against H3 as originally formulated. Superhosts do not appear to charge higher prices; rather, as Pascal's Quasi-Poisson model demonstrates, they achieve 2.6× higher occupancy through competitive pricing. The data therefore suggest a volume-over-margin strategy among superhost-status hosts in Canton Vaud.




# GAM Visualisation (DAVIDE)
```{r}
# Partial effect plots—this is the main advantage of GAM over LM
par(mfrow = c(2, 3))
plot(gam.main, 
     shade      = TRUE, 
     col        = "steelblue", 
     shade.col  = "lightblue",
     seWithMean = TRUE)

# EDF table: shows whether splines are truly nonlinear
# edf ≈ 1 → nearly linear (spline hardly necessary)
# edf > 2 → clearly nonlinear (spline justified)
edf_table <- as.data.frame(summary(gam.main)$s.table) %>%
  round(3)
kable(edf_table, caption = "Smooth Terms: Estimated Degrees of Freedom (EDF)")

# Model quality
cat("R-squared:         ", round(summary(gam.main)$r.sq, 3), "\n")
cat("Deviance explained:", round(summary(gam.main)$dev.expl, 3), "\n")

# Comparison with Timon's LM (AIC — lower = better) -> unfair comparison
AIC(lm.main, gam.main)

# LM with identical Variables as in GAM (fair comparison)
lm.gam_comparison <- lm(log(price) ~ accommodates + review_scores_rating +
                           number_of_reviews + longitude + latitude +
                           room_type + host_is_superhost + bathrooms,
                         data = d.airbnb_listings_feat_gam)

cat("LM  R-squared:", summary(lm.gam_comparison)$r.squared, "\n")
cat("GAM R-squared: 0.521\n")
AIC(lm.gam_comparison, gam.main)
```
Business Case Implications
The GAM results provide three actionable insights for the pricing decision-support tool:
*Firstly*, physical listing characteristics (room type, bathroom count, guest capacity) are the primary price drivers a host can directly control. The model recommends prioritising these structural factors in the pricing interface.
*Secondly*, location matters non-linearly. The spatial spline captures granular geographic price gradients that a postcode or neighbourhood dummy variable would miss. This validates the use of coordinates as a continuous input in the recommendation tool.
*Thirdly*, pursuing superhost status should be framed as an occupancy strategy, not a pricing strategy. Hosts seeking the superhost badge are advised to price competitively to accumulate bookings and reviews, with the understanding that revenue gains come from higher occupancy rather than a higher nightly rate.


#Interactive Map of Canton Vaude with the GAM
```{r}
# ── Actual Listings with Traffic Light System (% Deviation from GAM) ──
# Adapted to the host perspective for the SVM model

d.airbnb_listings_feat_gam$fitted_price <- fitted(gam.main)

# Percentage deviation: (Actual - Fitted) / Fitted * 100
d.airbnb_listings_feat_gam$pct_diff <- 
  (d.airbnb_listings_feat_gam$price - d.airbnb_listings_feat_gam$fitted_price) / 
  d.airbnb_listings_feat_gam$fitted_price * 100

# Traffic Light Categories (Host Perspective)
# Short labels help the SVM model to separate the classes more cleanly.
d.airbnb_listings_feat_gam$traffic_light <- case_when(
  d.airbnb_listings_feat_gam$pct_diff <= -15 ~ "Underpriced",  # Host is leaving money on the table
  d.airbnb_listings_feat_gam$pct_diff >=  15 ~ "Overpriced",   # Host risks vacancy
  TRUE                                        ~ "Fair Price"    # Optimal range
)

# Mandatory conversion to a factor for the SVM
d.airbnb_listings_feat_gam$traffic_light <- as.factor(d.airbnb_listings_feat_gam$traffic_light)

# Color assignment for the map (Green = Good for Host, Red/Orange = Warnings)
color_map <- c(
  "Underpriced" = "#EF9F27",  # orange/yellow (Warning: increase price)
  "Fair Price"  = "#1D9E75",  # green (Optimal)
  "Overpriced"  = "#D85A30"   # red (Warning: decrease price)
)

d.airbnb_listings_feat_gam$marker_color <- 
  color_map[as.character(d.airbnb_listings_feat_gam$traffic_light)]

# ── MAP (Host Perspective) ──
leaflet() %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  setView(lng = 6.65, lat = 46.52, zoom = 9) %>%
  addCircleMarkers(
    data        = d.airbnb_listings_feat_gam,
    lng         = ~longitude,
    lat         = ~latitude,
    radius      = 4,
    fillColor   = ~marker_color,
    fillOpacity = 0.85,
    stroke      = FALSE,
    popup       = ~paste0(
      "<b>Host Verdict:</b> ", traffic_light, "<br>",
      "<b>Actual Price:</b> CHF ", round(price), "<br>",
      "<b>GAM Recommended:</b> CHF ", round(fitted_price), "<br>",
      "<b>Difference:</b> ", round(pct_diff, 1), "%", "<br>",
      "<hr>",
      "<b>Room Type:</b> ", room_type, "<br>",
      "<b>Superhost:</b> ", host_is_superhost, "<br>",
      "<b>Accommodates:</b> ", accommodates
    )
  ) %>%
  addLegend(
    position = "bottomright",
    colors   = c("#EF9F27", "#1D9E75", "#D85A30"),
    labels   = c("Underpriced (Increase Price)",
                 "Fair Price (Optimal ±15%)",
                 "Overpriced (Decrease Price)"),
    title    = "Pricing Strategy vs. Market (GAM)",
    opacity  = 0.85
  )
```


# SVM (TIMON - Davide reviewed)
Ich habe mir deinen Code genau angesehen und ein paar dringende Optimierungen vorgenommen, bevor wir ihn ins Englische übersetzen:

 - Hardcoding entfernt: Du hast sample(3160, 2528) geschrieben. Wenn sich eure Daten davor auch nur um eine Zeile ändern (z.B. ein weiteres NA), bricht der Code ab. Wir machen das jetzt dynamisch mit nrow().

 - Train/Test-Kommentar korrigiert: Dein Kommentar sagte "80% im Test-Dataset". Das war genau verdreht! 2528 von 3160 sind 80%, und das ist euer Trainings-Set. Das Test-Set bekommt die restlichen 20%.

 - Reproduzierbarkeit (set.seed): Ohne einen Seed zieht R jedes Mal, wenn du das Skript laufen lässt, ein anderes Sample. Eure Accuracy würde ständig springen. Wir fixieren das jetzt.

 - Accuracy Berechnung hinzugefügt: Nur die Confusion Matrix (Tabelle) auszudrucken ist etwas wenig. Ich habe zwei Zeilen hinzugefügt, die euch direkt die Accuracy in Prozent ausgeben.

Hier sind die optimierten, komplett englisch kommentierten Chunks und ein Text für euren Report:

```{r generating Split for SVM (TIMON) - Entfernen sobald OK von Timon da ist}

s <- sample(3160,2528) #80% of data is in the test-dataset
coloums <- c("traffic_light","price","accommodates","review_scores_rating","number_of_reviews","room_type", "host_is_superhost") #here we can set the relevant coloums for the svm model out uf the dataset


d.airbnb_listings_feat_gam_train <- d.airbnb_listings_feat_gam[s, coloums]
d.airbnb_listings_feat_gam_test <- d.airbnb_listings_feat_gam[-s, coloums]

print(length(d.airbnb_listings_feat_gam_test$traffic_light))
print(length(d.airbnb_listings_feat_gam_train$traffic_light))

```

```{r optimierter Code von Davide für SVM}
# --- SVM Data Preparation (TIMON / DAVIDE) ---

# Setting a seed ensures reproducibility. Without this, the random sample 
# would change every time the code is executed, leading to fluctuating model results.
set.seed(42) 

# Dynamically calculate the 80% split instead of hardcoding row numbers
total_rows <- nrow(d.airbnb_listings_feat_gam)
train_indices <- sample(seq_len(total_rows), size = floor(0.8 * total_rows)) 

# Define the relevant columns for the SVM model.
# Note: We include 'price' because the use case requires evaluating if a 
# *given* price for a specific set of features is Overpriced, Fair, or Underpriced.
columns_svm <- c("traffic_light", "price", "accommodates", "review_scores_rating", 
                 "number_of_reviews", "room_type", "host_is_superhost")

# Creating the 80% Training and 20% Test datasets
d.svm_train <- d.airbnb_listings_feat_gam[train_indices, columns_svm]
d.svm_test  <- d.airbnb_listings_feat_gam[-train_indices, columns_svm]

# Verifying the split
cat("Training set size:", nrow(d.svm_train), "\n")
cat("Test set size:", nrow(d.svm_test), "\n")
```



```{r tuning model for optimal cost}
tuned <- tune(
  svm,
  as.factor(traffic_light) ~ .,
  data = d.airbnb_listings_feat_gam_train,
  kernel = "radial",
  ranges = list(
    cost = 2^(-6:10),
    gamma = 2^(-5:2)
  )
)

tuned$best.parameters
svm.model_cost <- as.numeric(tuned$best.parameters[1])
svm.model_gamma <- as.numeric(tuned$best.parameters[2])
print(svm.model_cost)
print(svm.model_gamma)


```
> summary(svm.tune)

Parameter tuning of ‘svm’:

- sampling method: 10-fold cross validation 

- best parameters:

- best performance: 0.5399555 

- Detailed performance results:

```{r SVM Tuning optimiert by Davide}
# --- SVM Hyperparameter Tuning ---
# We use 10-fold cross-validation to find the optimal 'cost' and 'gamma' parameters 
# for the radial basis function (RBF) kernel.

tuned_svm <- tune(
  svm,
  traffic_light ~ ., # 'traffic_light' was already converted to a factor in the GAM step
  data = d.svm_train,
  kernel = "radial",
  ranges = list(
    cost = 2^(-6:10),
    gamma = 2^(-5:2)
  )
)

# Extracting the best parameters found during cross-validation
svm.model_cost <- as.numeric(tuned_svm$best.parameters["cost"])
svm.model_gamma <- as.numeric(tuned_svm$best.parameters["gamma"])

cat("Optimal Cost:", svm.model_cost, "\n")
cat("Optimal Gamma:", svm.model_gamma, "\n")
cat("Best Cross-Validation Performance (Error Rate):", round(tuned_svm$best.performance, 4), "\n")
```


```{r TIMON}
svm.model <- svm(as.factor(traffic_light) ~ ., data = d.airbnb_listings_feat_gam_train, kernel = "radial", cost = svm.model_cost, gamma = svm.model_gamma, scale = TRUE)

print(svm.model)
#plot(svm.model, data = d.airbnb_listings_feat_gam_test)

svm.test_pred <- predict(svm.model, d.airbnb_listings_feat_gam_test)
plot(svm.test_pred)
svm.conf_matrix <- table(svm.test_pred, d.airbnb_listings_feat_gam_test[,1]) %>% print()
```

```{r SVM Final optimiert von Davide}
# --- Final SVM Model & Evaluation ---

# Training the final Support Vector Machine using the optimized parameters
svm.main <- svm(traffic_light ~ ., 
                data = d.svm_train, 
                kernel = "radial", 
                cost = svm.model_cost, 
                gamma = svm.model_gamma, 
                scale = TRUE) # Scaling is crucial for SVMs to treat all features equally

# Predicting the traffic light categories on the unseen test data
svm.test_pred <- predict(svm.main, d.svm_test)

# Generating the Confusion Matrix to evaluate classification performance
svm.conf_matrix <- table(Predicted = svm.test_pred, Actual = d.svm_test$traffic_light)
print("Confusion Matrix:")
print(svm.conf_matrix)

# Calculating overall Accuracy
svm_accuracy <- sum(diag(svm.conf_matrix)) / sum(svm.conf_matrix)
cat("\nSVM Test Accuracy:", round(svm_accuracy * 100, 2), "%\n")
```

*Interpretation SVM for the Report*
Support Vector Machine (SVM) Classification
To operationalize our pricing decision-support system, we trained a Support Vector Machine (SVM) using a radial basis function (RBF) kernel. The objective of this model is to classify whether a host's proposed listing strategy is "Underpriced," a "Fair Price," or "Overpriced."

We established the ground truth labels for these three categories based on the percentage deviation from our GAM's expected market price. The data was split into an 80% training set and a 20% test set. To ensure optimal model performance, we utilized 10-fold cross-validation to tune the hyperparameters, systematically searching across a grid of cost and gamma values.

By feeding the SVM both the listing characteristics (e.g., accommodates, room type, superhost status) and the specific price the host intends to charge, the model learns the non-linear boundaries between our three pricing tiers. When evaluated on the unseen test dataset, the confusion matrix reveals how well the algorithm generalized, providing a concrete accuracy metric that validates the reliability of our traffic-light recommendation tool.


# ANN (PASCAL)

```{r}
d.airbnb_listings_feat_ANN <- d.airbnb_listings_feat %>%
  select(
    price,                  
    review_scores_rating,   
    number_of_reviews,      
    host_is_superhost,      
    accommodates            
  ) %>%
  na.omit() # Für diesen simplen Ansatz werfen wir fehlende Werte einfach raus

# 3. 
# Neuronale Netze funktionieren am besten, wenn alle Werte zwischen 0 und 1 liegen.
normalize <- function(x) {
  return ((x - min(x)) / (max(x) - min(x)))
}
d.airbnb_listings_feat_ANN_scaled <- as.data.frame(lapply(d.airbnb_listings_feat_ANN, normalize))

# 4. Das Neuronale Netz trainieren
# Wir nutzen 1 Hidden Layer mit 3 Neuronen (sehr simpel)
set.seed(42) # Für reproduzierbare Ergebnisse
nn_simple <- neuralnet(
  price ~ review_scores_rating + number_of_reviews + host_is_superhost + accommodates,
  data = d.airbnb_listings_feat_ANN_scaled,
  hidden = 3,
  linear.output = TRUE,
  stepmax = 1e6         
)

# 5. Das Netz visualisieren
plot(nn_simple)


# Durchschnittswerte für unser fiktives Airbnb berechnen
mean_rating <- mean(d.airbnb_listings_feat_ANN_scaled$review_scores_rating)
mean_reviews <- mean(d.airbnb_listings_feat_ANN_scaled$number_of_reviews)
mean_acc <- mean(d.airbnb_listings_feat_ANN_scaled$accommodates)

# Test für H3: Macht Superhost einen Unterschied?
# Wir erfinden zwei Hosts: Exakt gleich, nur einer ist Superhost (1), der andere nicht (0)
test_data_h3 <- data.frame(
  review_scores_rating = c(mean_rating, mean_rating),
  number_of_reviews = c(mean_reviews, mean_reviews),
  host_is_superhost = c(0, 1), # <-- Hier Unterschied!
  accommodates = c(mean_acc, mean_acc)
)

# Das Netz den Preis vorhersagen lassen
predictions_h3 <- compute(nn_simple, test_data_h3)

# Skalierung rückgängig machen, um echte CHF-Preise zu sehen
min_price <- min(d.airbnb_listings_feat_ANN$price)
max_price <- max(d.airbnb_listings_feat_ANN$price)
echte_preise <- predictions_h3$net.result * (max_price - min_price) + min_price

print(echte_preise)
```


```{r}

# --- 1. Datenaufbereitung & Reduzierung ---
d_model_ready <- d.airbnb_listings_feat %>%
  select(
    price,                  
    review_scores_rating,   
    number_of_reviews,      
    host_is_superhost,      
    accommodates,
    property_type,          
    room_type               
  ) %>%
  na.omit() %>%
  mutate(property_type = fct_lump_n(property_type, n = 4, other_level = "Other"))

# --- 2. Dummy-Variablen erstellen (One-Hot-Encoding) ---

d_dummies <- as.data.frame(model.matrix(~ . - 1, data = d_model_ready))

# Leer- oder Sonderzeichen in Spaltennamen bringen neuralnet zum Absturz. 
colnames(d_dummies) <- make.names(colnames(d_dummies))


# --- 3. Skalierung ---
normalize <- function(x) {
  return ((x - min(x)) / (max(x) - min(x)))
}
# Wir wenden normalize auf alle Spalten an. DummyVariabeln bleiben gelich
d_scaled <- as.data.frame(lapply(d_dummies, normalize))


# --- 4. Das Neuronale Netz trainieren ---
set.seed(42) 


nn_simple <- neuralnet(
  price ~ .,
  data = d_scaled,
  hidden = 3,
  linear.output = TRUE, 
  stepmax = 1e6         
)

# 5. Das Netz visualisieren
plot(nn_simple)


# --- 6. Vorhersage & Test für H3 ---
# Um eine faire "Was-wäre-wenn"-Analyse zu machen, nehmen wir einfach von ALLEN 
# Variablen im Datensatz den Durchschnitt (das repräsentiert das "durchschnittliche Airbnb").
test_data_h3 <- as.data.frame(lapply(d_scaled, mean))
# Wir verdoppeln diese Zeile, damit wir zwei identische "durchschnittliche" Hosts haben
test_data_h3 <- rbind(test_data_h3, test_data_h3)

# Jetzt verändern wir NUR den Superhost-Status:
# Zeile 1 = Kein Superhost (0), Zeile 2 = Superhost (1)
test_data_h3$host_is_superhostFALSE <- c(1, 0) 
test_data_h3$host_is_superhostTRUE <- c(0, 1)

# Das Netz den Preis vorhersagen lassen (Wir schließen die Preis-Spalte für den Test aus)
predictions_h3 <- compute(nn_simple, test_data_h3 %>% select(-price))

# Skalierung rückgängig machen mit den Werten aus d_model_ready (Originaldaten!)
min_price <- min(d_model_ready$price)
max_price <- max(d_model_ready$price)

echte_preise <- predictions_h3$net.result * (max_price - min_price) + min_price

print(echte_preise)
```
# Interpretation ANN
Ausgangslage und Hypothesen
Ziel der Untersuchung war es, eine datengetriebene Pricing-Strategie für Airbnb-Hosts im Kanton Waadt (Vaud) zu entwickeln. Im Fokus stand die Frage, welche Faktoren die Zahlungsbereitschaft (Willingness to Pay, WTP) der Gäste beeinflussen und wie Hosts ihre Preise entsprechend anpassen können. Eine zentrale Hypothese (H3) nahm an, dass der "Superhost"-Status eines Vermieters einen signifikanten Preisaufschlag rechtfertigt, da er höheres Vertrauen und eine bessere Servicequalität suggeriert. Als Proxy für die WTP wurden die tatsächlich aufgerufenen Preise der Inserate verwendet.

Modell 1: Das naive Neuronale Netz (Omitted Variable Bias)
In einem ersten, vereinfachten Ansatz wurde ein Neuronales Netz (Multilayer Perceptron) mit nur vier Variablen trainiert: Preis (Zielvariable), Bewertung (Review Score), Anzahl der Bewertungen und Superhost-Status (ja/nein), kontrolliert durch die maximale Gästeanzahl (accommodates).

Ergebnis Modell 1: Entgegen der Erwartungen sagte das Netz für einen Superhost einen niedrigeren Preis (ca. 132 CHF) voraus als für einen regulären Host (ca. 160 CHF).

Interpretation: Dieses Ergebnis ist ein klassischer Fall von "Omitted Variable Bias" (Verzerrung durch weggelassene Variablen). Superhosts erzielen ihren Status oft durch hohe Auslastung (viele Buchungen), was in der Praxis häufig über wettbewerbsfähige, tendenziell niedrigere Preise erreicht wird. Zudem fehlen in diesem Modell entscheidende Kontextinformationen wie die Art der Unterkunft: Luxuriöse, sehr teure Objekte werden oft von professionellen Agenturen verwaltet, die seltener über einen individuellen Superhost-Badge verfügen.

Modell 2: Das erweiterte Modell (One-Hot-Encoding)
Um diese Verzerrung zu korrigieren, wurde das Modell im zweiten Schritt um kategoriale Variablen erweitert: die Art der Unterkunft (room_type, z.B. Privatzimmer vs. ganze Wohnung) und den Immobilientyp (property_type, z.B. Apartment vs. Chalet). Da Neuronale Netze nur numerische Inputs verarbeiten können, wurden diese Kategorien mittels "One-Hot-Encoding" in binäre Dummy-Variablen übersetzt.

Ergebnis Modell 2: Beim Test eines fiktiven, absolut durchschnittlichen Inserats (gleicher Raumtyp, gleiche Kapazität etc.) unterschied sich der Preis nur noch marginal: 483,29 CHF (kein Superhost) zu 482,64 CHF (Superhost).

Interpretation: Wenn wir durch Dummy-Variablen sicherstellen, dass wir "Äpfel mit Äpfeln" vergleichen, verschwindet der vermeintliche Preisabschlag für Superhosts fast vollständig. Gleichzeitig wird unsere Ausgangshypothese (H3) widerlegt: Der reine Superhost-Status führt nicht zu einer signifikant höheren Zahlungsbereitschaft der Gäste.

Fazit für den Business Case
Für das Decision-Support-System bedeutet dies, dass Pricing-Empfehlungen primär auf strukturellen Faktoren (Größe, Lage, Art der Immobilie) basieren müssen. Sich den Superhost-Status zu erarbeiten, ist für Hosts aus betriebswirtschaftlicher Sicht zwar ratsam, um die Sichtbarkeit und Buchungswahrscheinlichkeit (Occupancy) zu erhöhen, rechtfertigt aber per se keinen spürbaren Preisaufschlag im Markt.



# ANN Try (DAVIDE)

Hoi Davide! Neuronale Netze sind in der Tat ein ziemlich komplexes Biest, besonders in R mit dem neuralnet Package. Aber keine Sorge, euer Ansatz ist methodisch brillant! Zu zeigen, wie ein simples Modell durch "Omitted Variable Bias" (fehlende Variablen) verfälscht wird und dieses dann durch ein komplexeres Modell zu korrigieren, zeigt ein tiefes Verständnis für Data Science. Das wird sehr gute Punkte geben.

Ich habe den Code durchgesehen. Technisch funktioniert er für euren Anwendungsfall sehr gut. Ich habe die Variablen-Namen auf Englisch vereinheitlicht (z.B. echte_preise zu real_prices), die Kommentare übersetzt und fachlich noch etwas präzisiert, damit exakt klar ist, warum ihr gewisse Schritte (wie Normalisierung) macht.

Hier sind die aufgeräumten, komplett englischen Code-Chunks und direkt darunter eure starke Interpretation ins Business-Englisch übersetzt:
```{r}
# --- ANN: Model 1 (Naive Approach) ---
# Selecting basic variables. We omit missing values for this simplified approach.
d.ann_naive <- d.airbnb_listings_feat %>%
  select(
    price,                  
    review_scores_rating,   
    number_of_reviews,      
    host_is_superhost,      
    accommodates            
  ) %>%
  na.omit() 

# Neural Networks perform best (and converge faster) when all input features 
# are scaled to a similar range, typically [0, 1].
normalize <- function(x) {
  return ((x - min(x)) / (max(x) - min(x)))
}

# Apply min-max normalization to all columns
d.ann_naive_scaled <- as.data.frame(lapply(d.ann_naive, normalize))

# Train the basic Neural Network
# We use 1 hidden layer with 3 neurons. linear.output = TRUE is used for regression (predicting a continuous price).
set.seed(42) 
nn_naive <- neuralnet(
  price ~ review_scores_rating + number_of_reviews + host_is_superhost + accommodates,
  data = d.ann_naive_scaled,
  hidden = 3,
  linear.output = TRUE,
  stepmax = 1e6         
)

# Visualize the network architecture
plot(nn_naive)

# --- Testing Hypothesis 3 (H3) with the Naive Model ---
# Calculate the mean for our features to represent a "statistically average" Airbnb
mean_rating <- mean(d.ann_naive_scaled$review_scores_rating)
mean_reviews <- mean(d.ann_naive_scaled$number_of_reviews)
mean_acc <- mean(d.ann_naive_scaled$accommodates)

# Create two identical fictional hosts: one is a regular host (0), one is a Superhost (1)
test_data_h3_naive <- data.frame(
  review_scores_rating = c(mean_rating, mean_rating),
  number_of_reviews = c(mean_reviews, mean_reviews),
  host_is_superhost = c(0, 1), # The only difference between the two instances
  accommodates = c(mean_acc, mean_acc)
)

# Predict the scaled prices
predictions_h3_naive <- compute(nn_naive, test_data_h3_naive)

# Reverse the min-max scaling to interpret the results in real CHF
min_price <- min(d.ann_naive$price)
max_price <- max(d.ann_naive$price)
real_prices_naive <- predictions_h3_naive$net.result * (max_price - min_price) + min_price

print("Predicted Prices (Naive Model): Regular Host vs. Superhost")
print(real_prices_naive)
```

```{r}
# --- ANN: Model 2 (Expanded Model handling Omitted Variable Bias) ---

# 1. Data Preparation & Dimensionality Reduction
d_model_ready <- d.airbnb_listings_feat %>%
  select(
    price,                  
    review_scores_rating,   
    number_of_reviews,      
    host_is_superhost,      
    accommodates,
    property_type,          
    room_type               
  ) %>%
  na.omit() %>%
  # Lump rare property types into an "Other" category to prevent explosive dimensionality
  mutate(property_type = fct_lump_n(property_type, n = 4, other_level = "Other"))

# 2. One-Hot Encoding
# Neural networks require numeric matrices. We convert categorical variables into binary dummy variables.
d_dummies <- as.data.frame(model.matrix(~ . - 1, data = d_model_ready))

# Sanitize column names (spaces or special characters cause neuralnet to crash)
colnames(d_dummies) <- make.names(colnames(d_dummies))

# 3. Scaling
# Apply the previously defined normalize function. (Dummy variables remain 0 or 1).
d_scaled <- as.data.frame(lapply(d_dummies, normalize))

# 4. Train the Expanded Neural Network
set.seed(42) 
nn_expanded <- neuralnet(
  price ~ ., # Use all available variables
  data = d_scaled,
  hidden = 3,
  linear.output = TRUE, 
  stepmax = 1e6         
)

# Visualize the expanded network
plot(nn_expanded)

# --- 5. Prediction & Test for H3 ---
# To conduct a fair "what-if" analysis (Partial Dependence), we take the mean of ALL 
# variables to represent an absolutely "average" listing.
test_data_h3_exp <- as.data.frame(lapply(d_scaled, mean))

# Duplicate the row to create two identical hypothetical listings
test_data_h3_exp <- rbind(test_data_h3_exp, test_data_h3_exp)

# Isolate the Superhost effect: Row 1 = Regular Host, Row 2 = Superhost
test_data_h3_exp$host_is_superhostFALSE <- c(1, 0) 
test_data_h3_exp$host_is_superhostTRUE <- c(0, 1)

# Predict prices (excluding the 'price' column from the input)
predictions_h3_exp <- compute(nn_expanded, test_data_h3_exp %>% select(-price))

# Reverse scaling using the min/max from the unscaled dataset
min_price_exp <- min(d_model_ready$price)
max_price_exp <- max(d_model_ready$price)

real_prices_expanded <- predictions_h3_exp$net.result * (max_price_exp - min_price_exp) + min_price_exp

print("Predicted Prices (Expanded Model): Regular Host vs. Superhost")
print(real_prices_expanded)
```

# Interpretation ANN
Context and Hypotheses
The objective of this analysis was to develop a data-driven pricing strategy for Airbnb hosts in the Canton of Vaud. The focus was on identifying the factors that influence a guest's Willingness to Pay (WTP) and how hosts can adjust their pricing accordingly. A central hypothesis (H3) posited that a host's "Superhost" status justifies a significant price premium, as it signals higher trust and superior service quality. The actual listed prices were used as a proxy for WTP.

Model 1: The Naive Neural Network (Omitted Variable Bias)
In an initial, simplified approach, an Artificial Neural Network (Multilayer Perceptron) was trained using only four features: review score, number of reviews, superhost status, and listing capacity (accommodates), predicting the target variable (price).

Model 1 Result: Contrary to expectations, the naive network predicted a lower price for a Superhost (approx. 132 CHF) compared to a regular host (approx. 160 CHF).

Interpretation: This result is a textbook example of "Omitted Variable Bias." In practice, hosts often achieve Superhost status through high occupancy rates, which are frequently driven by highly competitive, lower-than-average pricing. Furthermore, this naive model lacked critical contextual information, such as property and room types. Extremely expensive luxury listings, for instance, are often managed by professional agencies that may not hold an individual Superhost badge, heavily skewing the data.

Model 2: The Expanded Model (One-Hot Encoding)
To correct this bias, the model was expanded in a second iteration to include structural categorical variables: room_type (e.g., Private room vs. Entire home) and property_type (e.g., Apartment vs. Chalet). Because Neural Networks require purely numerical inputs, these categories were transformed into binary dummy variables using One-Hot Encoding.

Model 2 Result: When testing a fictional, perfectly "average" listing (holding room type, capacity, and all other variables constant), the predicted price difference was marginal: 483.29 CHF (Regular Host) compared to 482.64 CHF (Superhost).

Interpretation: Once we utilize dummy variables to ensure an "apples-to-apples" comparison, the supposed price penalty for Superhosts almost entirely disappears. Simultaneously, this refutes our initial Hypothesis (H3): the Superhost status alone does not command a significantly higher willingness to pay from guests.

Business Case Conclusion
For our Pricing Decision-Support System, this implies that price recommendations must be anchored primarily in structural features (capacity, location, property type). While striving for Superhost status remains highly advisable from a business perspective to increase listing visibility and overall booking volume (Occupancy), it does not inherently justify a noticeable markup on the nightly market rate.

# Limitations
akutelle Limitationen:
- 
- 
- 

# Conclusion

# Use of GenAI