---
title: "Main Group Work_ML"
author: "Pascal Ingold, Timon Reichmuth und Davide Gualdani"
date: "2026-03-22"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

Next Steps: 

Pascal: 
- Data Cleaning von 80 predictors auf <20 runterbrechen
- Feature Engineering ausbauen
- Business Case liefern

Timon: 
- Forschungsfragen ausarbeiten
- Outlier Handling definieren.

Davide: 
- EDA Ausbauen

TBD: 
- ML Methoden verteilen auf uns 3

# Business Case
We propose a data-driven pricing decision-support system for Airbnb hosts in the
Canton of Vaud, providing market-aligned price recommendations based on listing
characteristics. The tool targets both new and existing hosts, using inputs such
as location, room type, guest capacity, availability, ratings, and superhost status.
Machine learning models, including LM, GLM, GAM, SVM, and Neural
Networks, will be applied to identify key price drivers and generate accurate
predictions. The objective is to balance revenue and occupancy, helping hosts
understand their market value.
A key feature is interpretability, enabling users to see why a certain price is
recommended. From a business perspective, the system aims to increase booking
rates, reduce price dispersion, and improve host satisfaction.
The model is tailored to the Canton of Vaud to capture regional economic effects.

First step: Loading Packages

```{r}
library(tidyverse)
library(janitor)
library(mgcv)
library(e1071)
library(nnet)
library(caret)
library(pROC)
library(Metrics)
library(performance)
library(broom)
library(stringr)
library(ggplot2)
library(patchwork)
```



Second step: Reading all available Datasets
```{r}
path_listings <- "./Data_Vaud_all/listings-2.csv"
d.raw_data_listings <- read.csv(path_listings, header = TRUE)
```

Cleaning the Airbnb-Listings Data. 

```{r}
d.airbnb_listings <- d.raw_data_listings %>%
  clean_names()

d.listings_clean <- d.airbnb_listings %>%
  mutate(
    price = str_remove_all(price, "\\$|,"),
    price = as.numeric(price),
    host_is_superhost = ifelse(host_is_superhost %in% c("t", "TRUE", "1"), 1, 0),
    room_type = as.factor(room_type),
    property_type = as.factor(property_type)
  ) %>%
  filter(!is.na(price), price > 0)

d.listings_clean <- d.listings_clean %>%
  select(-listing_url, -scrape_id, -last_scraped, -source, -name, -description,
         -neighborhood_overview, -picture_url, -host_url, -host_name, -host_location,
         -host_about, -host_thumbnail_url, -host_picture_url, -host_neighbourhood,
         -neighbourhood, -neighbourhood_cleansed, -bathrooms_text, -amenities, 
         -minimum_minimum_nights, -maximum_minimum_nights, -minimum_maximum_nights,
         -maximum_maximum_nights, -minimum_nights_avg_ntm, -maximum_nights_avg_ntm,
         -calendar_updated, -has_availability, -availability_30, -availability_60,
         -availability_90, -calendar_last_scraped, -number_of_reviews_ltm, -number_of_reviews_l30d,
         -availability_eoy, -number_of_reviews_ly, -review_scores_accuracy, -review_scores_cleanliness,
         -review_scores_checkin, -review_scores_communication, -review_scores_location,
         -license, -instant_bookable, -calculated_host_listings_count_entire_homes, -calculated_host_listings_count_private_rooms,
         -calculated_host_listings_count_shared_rooms, -reviews_per_month, -bedrooms, -review_scores_value)
d.listings_clean <- d.listings_clean %>%
  mutate(host_since_year = str_extract(host_since, "^\\d{4}"))
```


First Feature Engineering for the airbnb_listings data

```{r}
d.airbnb_listings_feat <- d.listings_clean %>%
  mutate(
    log_price = log(price),
    price_per_guest = price / pmax(accommodates, 1),
    high_demand = ifelse(availability_365 < 100, 1, 0)
  )
```

Exploratory Data Analysis

```{r}
hist(d.airbnb_listings_feat$price)
hist(d.airbnb_listings_feat$log_price)
ggplot(data= d.airbnb_listings_feat, aes(x = room_type, y = price)) +
  geom_boxplot() +
  scale_y_log10() +
  labs(
    title = "Preisverteilung nach Room Type (log-Skala)",
    x = "Room Type",
    y = "Log_Preis"
  ) +
  theme_minimal()

ggplot(data = d.airbnb_listings_feat, aes(x = accommodates, y = log_price)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "loess", se = FALSE) +
  labs(
    title = "Preis vs. Kapazität (log-Skala)",
    x = "Anzahl Gäste",
    y = "Log_Preis"
  ) +
  theme_minimal()
  
ggplot(data = d.airbnb_listings_feat, aes(x = log_price, y = number_of_reviews)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "loess", se = FALSE) +
  labs(
    title = "Anzahl Reviews vs. Preis",
    x = "Log_Preis",
    y = "Anzahl Reviews"
  ) +
  theme_minimal()

ggplot(data = d.airbnb_listings_feat, aes(x = longitude, y = latitude)) +
  geom_point(aes(color = log_price), alpha = 0.6) +
  scale_color_viridis_c() +
  labs(
    title = "Airbnb Listings (Farbe = Log-Preis)",
    x = "Longitude",
    y = "Latitude",
    color = "Log_Preis"
  ) +
  theme_minimal()

ggplot(data = d.airbnb_listings_feat, aes(x = longitude, y = latitude)) +
  geom_point(aes(color = log_price, size = number_of_reviews), alpha = 0.5) +
  scale_color_viridis_c() +
  labs(
    title = "Listings: Preis und Nachfrage",
    x = "Longitude",
    y = "Latitude",
    color = "Log_Preis",
    size = "Reviews"
  ) +
  theme_minimal()
```

Davide's EDA
```{r}
# --- 1. Distribution of the Target Variable (Price) ---
# Upgrading base R hist() to ggplot2 for a more professional look.
# Showing both raw and log side-by-side helps visually justify the transformation to the client.
p_raw <- ggplot(d.airbnb_listings_feat, aes(x = price)) +
  geom_histogram(bins = 50, fill = "steelblue", color = "white") +
  labs(title = "Raw Price Distribution", x = "Price", y = "Count") +
  theme_minimal()

p_log <- ggplot(d.airbnb_listings_feat, aes(x = log_price)) +
  geom_histogram(bins = 50, fill = "darkgreen", color = "white") +
  labs(title = "Log-transformed Price Distribution", x = "Log(Price)", y = "Count") +
  theme_minimal()
p_raw + p_log # Displays them side-by-side

# --- 2. Categorical Predictor: Room Type vs. Price ---
# Your original code was good, just cleaned up the aesthetics.
ggplot(data = d.airbnb_listings_feat, aes(x = room_type, y = log_price, fill = room_type)) +
  geom_boxplot(alpha = 0.7) +
  labs(
    title = "Price Distribution by Room Type",
    x = "Room Type",
    y = "Log(Price)"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

# --- 3. Continuous Predictor: Capacity vs. Price ---
ggplot(data = d.airbnb_listings_feat, aes(x = accommodates, y = log_price)) +
  geom_point(alpha = 0.3, color = "darkgray") +
  geom_smooth(method = "loess", color = "blue", se = FALSE) +
  labs(
    title = "Price vs. Capacity",
    subtitle = "Smoother applied to identify potential non-linear trends",
    x = "Number of Guests (Accommodates)",
    y = "Log(Price)"
  ) +
  theme_minimal()

# --- 4. Continuous Predictor: Reviews vs. Price ---
# Correction: The response variable (log_price) should generally be on the y-axis 
# when plotting against a predictor[cite: 255].
ggplot(data = d.airbnb_listings_feat, aes(x = number_of_reviews, y = log_price)) +
  geom_point(alpha = 0.3, color = "darkgray") +
  geom_smooth(method = "loess", color = "red", se = FALSE) +
  labs(
    title = "Does the Number of Reviews impact Price?",
    x = "Total Number of Reviews",
    y = "Log(Price)"
  ) +
  theme_minimal()

# --- 5. Binary Predictor: Superhost Status ---
# Adding this because it's a key part of your business case and fulfills the binary variable requirement.
ggplot(data = d.airbnb_listings_feat, aes(x = as.factor(host_is_superhost), y = log_price, fill = as.factor(host_is_superhost))) +
  geom_boxplot(alpha = 0.7) +
  scale_x_discrete(labels = c("0" = "Regular Host", "1" = "Superhost")) +
  labs(
    title = "Price Difference: Superhosts vs. Regular Hosts",
    x = "Host Status",
    y = "Log(Price)"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

# --- 6. Spatial Data: Mapping the Listings ---
# Cleaned up the map to focus purely on location and price.
ggplot(data = d.airbnb_listings_feat, aes(x = longitude, y = latitude)) +
  geom_point(aes(color = log_price), alpha = 0.6, size = 1.5) +
  scale_color_viridis_c(option = "plasma") +
  labs(
    title = "Airbnb Listings in Vaud by Price",
    x = "Longitude",
    y = "Latitude",
    color = "Log(Price)"
  ) +
  theme_minimal()
```
Davide's overall Data Analysis
1. Overall Market Overview (Prices & Capacity)
This is the absolute minimum, maximum, median, and mean for the most important numeric variables.
```{r}
# Basic summary statistics for key numerical variables
summary_stats <- d.airbnb_listings_feat %>%
  select(price, log_price, accommodates, number_of_reviews, availability_365) %>%
  summary()

print(summary_stats)
```
Client Interpretation: "The average Airbnb in Vaud costs X CHF per night, but the median is Y CHF, indicating that a few extremely expensive luxury listings might be skewing the average upward. Prices range from a minimum of A CHF to a maximum of B CHF."

2. Analysis by Room Type
Since we want to know how many different types of rooms there are, we group the data and calculate the counts, percentages, and price ranges for each type.
```{r}
# Detailed breakdown by Room Type
room_type_analysis <- d.airbnb_listings_feat %>%
  group_by(room_type) %>%
  summarise(
    Total_Listings = n(),
    Percentage_of_Market = round((n() / nrow(d.airbnb_listings_feat)) * 100, 1),
    Min_Price = min(price, na.rm = TRUE),
    Median_Price = median(price, na.rm = TRUE),
    Average_Price = round(mean(price, na.rm = TRUE), 1),
    Max_Price = max(price, na.rm = TRUE)
  ) %>%
  arrange(desc(Total_Listings)) # Sorts by the most common room type

print(room_type_analysis)
```
Client Interpretation: "Entire homes/apartments dominate the Vaud market, making up 75% of all listings. As expected, the Hotel room command the highest median price (264.5 CHF), compared to shared rooms (47 CHF)."

3. Superhost vs. Regular Host Breakdown
This business case specifically mentions superhost status. So we took a look to see how many there are and if they actually charge more or even less.
```{r}
# Superhost statistics
superhost_analysis <- d.airbnb_listings_feat %>%
  group_by(host_is_superhost) %>%
  summarise(
    Total_Hosts = n(),
    Percentage = round((n() / nrow(d.airbnb_listings_feat)) * 100, 1),
    Median_Price = median(price, na.rm = TRUE),
    Avg_Reviews = round(mean(number_of_reviews, na.rm = TRUE), 1)
  ) %>%
  mutate(host_is_superhost = ifelse(host_is_superhost == 1, "Yes", "No"))

print(superhost_analysis)
```


