---
title: "Main Group Work_ML"
author: "Pascal Ingold, Timon Reichmuth und Davide Gualdani"
date: "2026-03-22"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

Next Steps: 

Pascal: 
- Data Cleaning von 80 predictors auf <20 runterbrechen
- Feature Engineering ausbauen
- Business Case liefern

Timon: 
- Forschungsfragen ausarbeiten
- Outlier Handling definieren.

Davide: 
- EDA Ausbauen

TBD: 
- ML Methoden verteilen auf uns 3

# Business Case
We propose a data-driven pricing decision-support system for Airbnb hosts in the
Canton of Vaud, providing market-aligned price recommendations based on listing
characteristics. The tool targets both new and existing hosts, using inputs such
as location, room type, guest capacity, availability, ratings, and superhost status.
Advanced machine learning models—including LM, GLM, GAM, SVM, and Neural
Networks—will be applied to identify key price drivers and generate accurate
predictions. The objective is to balance revenue and occupancy, helping hosts
understand their market value.
A key feature is interpretability, enabling users to see why a certain price is
recommended. From a business perspective, the system aims to increase booking
rates, reduce price dispersion, and improve host satisfaction.
The model is tailored to the Canton of Vaud to capture regional economic effects
and can be extended to dynamic pricing in future iterations.

First step: Loading Packages

```{r}
library(tidyverse)
library(janitor)
library(mgcv)
library(e1071)
library(nnet)
library(caret)
library(pROC)
library(Metrics)
library(performance)
library(broom)
library(stringr)
library(ggplot2)
```



Second step: Reading all available Datasets
```{r}
path_listings <- "./Data_Vaud_all/listings-2.csv"
d.raw_data_listings <- read.csv(path_listings, header = TRUE)
```

Cleaning the Airbnb-Listings Data. 

```{r}
d.airbnb_listings <- d.raw_data_listings %>%
  clean_names()

d.listings_clean <- d.airbnb_listings %>%
  mutate(
    price = str_remove_all(price, "\\$|,"),
    price = as.numeric(price),
    host_is_superhost = ifelse(host_is_superhost %in% c("t", "TRUE", "1"), 1, 0),
    room_type = as.factor(room_type),
    property_type = as.factor(property_type)
  ) %>%
  filter(!is.na(price), price > 0)

d.listings_clean <- d.listings_clean %>%
  select(-listing_url, -scrape_id, -last_scraped, -source, -name, -description,
         -neighborhood_overview, -picture_url, -host_url, -host_name, -host_location,
         -host_about, -host_thumbnail_url, -host_picture_url, -host_neighbourhood,
         -neighbourhood, -neighbourhood_cleansed, -bathrooms_text, -amenities, 
         -minimum_minimum_nights, -maximum_minimum_nights, -minimum_maximum_nights,
         -maximum_maximum_nights, -minimum_nights_avg_ntm, -maximum_nights_avg_ntm,
         -calendar_updated, -has_availability, -availability_30, -availability_60,
         -availability_90, -calendar_last_scraped, -number_of_reviews_ltm, -number_of_reviews_l30d,
         -availability_eoy, -number_of_reviews_ly, -review_scores_accuracy, -review_scores_cleanliness,
         -review_scores_checkin, -review_scores_communication, -review_scores_location,
         -license, -instant_bookable, -calculated_host_listings_count_entire_homes, -calculated_host_listings_count_private_rooms,
         -calculated_host_listings_count_shared_rooms, -reviews_per_month, -bedrooms, -review_scores_value)
d.listings_clean <- d.listings_clean %>%
  mutate(host_since_year = str_extract(host_since, "^\\d{4}"))
```


First Feature Engineering for the airbnb_listings data

```{r}
d.airbnb_listings_feat <- d.listings_clean %>%
  mutate(
    log_price = log(price),
    price_per_guest = price / pmax(accommodates, 1),
    high_demand = ifelse(availability_365 < 100, 1, 0)
  )
```

Exploratory Data Analysis

```{r}
hist(d.airbnb_listings_feat$price)
hist(d.airbnb_listings_feat$log_price)
ggplot(data= d.airbnb_listings_feat, aes(x = room_type, y = price)) +
  geom_boxplot() +
  scale_y_log10() +
  labs(
    title = "Preisverteilung nach Room Type (log-Skala)",
    x = "Room Type",
    y = "Log_Preis"
  ) +
  theme_minimal()

ggplot(data = d.airbnb_listings_feat, aes(x = accommodates, y = log_price)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "loess", se = FALSE) +
  labs(
    title = "Preis vs. Kapazität (log-Skala)",
    x = "Anzahl Gäste",
    y = "Log_Preis"
  ) +
  theme_minimal()
  
ggplot(data = d.airbnb_listings_feat, aes(x = log_price, y = number_of_reviews)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "loess", se = FALSE) +
  labs(
    title = "Anzahl Reviews vs. Preis",
    x = "Log_Preis",
    y = "Anzahl Reviews"
  ) +
  theme_minimal()

ggplot(data = d.airbnb_listings_feat, aes(x = longitude, y = latitude)) +
  geom_point(aes(color = log_price), alpha = 0.6) +
  scale_color_viridis_c() +
  labs(
    title = "Airbnb Listings (Farbe = Log-Preis)",
    x = "Longitude",
    y = "Latitude",
    color = "Log_Preis"
  ) +
  theme_minimal()

ggplot(data = d.airbnb_listings_feat, aes(x = longitude, y = latitude)) +
  geom_point(aes(color = log_price, size = number_of_reviews), alpha = 0.5) +
  scale_color_viridis_c() +
  labs(
    title = "Listings: Preis und Nachfrage",
    x = "Longitude",
    y = "Latitude",
    color = "Log_Preis",
    size = "Reviews"
  ) +
  theme_minimal()
```

