---
title: "Pricing Decision-Support for Airbnb Hosts in Vaud"
subtitle: "Machine Learning 1 - Group Project"
author: "Pascal Ingold · Timon Reichmuth · Davide Gualdani"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    df_print: paged
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
    theme: cosmo
    fontsize: 10pt
    geometry: margin=2cm
    linestretch: 1.0
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, comment = NA, warning = FALSE, message=FALSE)

```

Tasks ab 17.05:
-   Lesson Learned Kapitel (Jeder selbst: unterkapitel pro person)(Pascal done)
-   Formatting conventions übernehmen (jeder für sich)(Pascal done)
-   Kapitelstruktur überarbeiten - Davide, Pascal(done)
-   Kontrollieren ob alle Modelle einen "Leader" haben (NAMEN aus Kapiteltitel löschen). PASCAL-->Done
-   Variabel-"call" kontrollieren. (ALLE)

-   beautifying HTML-Doc:
      - 2.0 Output --> drin lassen wenn ja verschönern. (TIMON)
      - 3.4 Grafiken grösser machen (TIMON) *erledigt*
      - 3.5 Grafiken grösser machen (TIMON) *erledigt*
      - 3.8 Grafiken grösser machen / Titelproblem lösen. (TIMON) *erledigt*
      - 3.10 Grafik kleiner machen (TIMON)
      - 4 & 4.1 Grafiken kleiner machen (TIMON)
      - 4.2 NA-Check irgendwie noch was dazu schreiben (TIMON)
      - 5.1 & 5.2 Outputs kabeln und verschönern (TIMON)
      - 5.4 Korrelationsmatrix grösser machen. (TIMON)
      - 5.8 Grafik grösser machen. (TIMON)
      - 5.9 OUtput verschönern. (TIMON)
      - Braucht es bei 6 für die eine Seite wirklich drei Kapitel oder kann man das in eines zusammenfassen? (TIMON)
      - 7.2 Rmodel einblenden (kabel) allenfalls direkt mit VIF kombinineren. (TIMON)
      - 7.3 Grafiken Kleiner machen. 
      - 8.4 Grafik kleiner machen
      - 9.3 Modelouput schön machen, bzw den OUtput oben entfernen. 
      - 10.3 Train und Test irgedwo sauber verpacken. nicht einfach so einblenden. 
      - 10.3output verschönern 
      - 10.3 machen diese 4 Visualisierungen wirklich Sinn? 
      - 10.3 einzelne kleine R-OUtputs in eine Tabelle und zusammen darstellen. 
      - 10.4.1 die einzelnen OUtputs in eine Tablle bringen und zusammen darstellen
      - 11.2 COnfusion matrix schöner machen (TIMON) *erledigt*
      - 12.3 Output schöner darstellen (kabel)
      - 12 keine einzige Darstellung beim ANN?
      
      (jeder seine Modelle, TIMON macht alles vor den Modellen. )



ToDos backlog: - R-chunks sauber personen zuweisen

-   Data Quality kontrollieren & übersetzen.\
-   EDA/DQ/UTD aussortieren falls platzbedarf
-   tuning wieder reaktivieren --\> done
-   README finalisieren
-   color fomatting kontrollieren. 
-   Ai-Kontrolle




OFFENE FRAGEN:
- 3.6 brauchen wir den zweiten Plot?
- 3.5 wie können wir die überlappungen lösen?
- 3.10 brauchen wir das?







- generell Grafiken auf eine Grösse fixieren --> Darstellung ist auf 100%, Darstellung kann mehr als 1 Plot haben. 

Infos zu bedenken: ABGABEDATUM: 12 Juni

Kaitelstruktur: 
Theorie (Theorieteil, Businesscase und Link zu Hypothese)
Model
Optionales
Interpretation (Bezug auf Businescase)
Limitation (Abstriche am Businescase)



Farbschema:

Hauptfarbe: deepskyblue\
schattigerung: lightblue Sekundärfarbe: maroon1 --\> wenn dunkler
benötigt bis "maroon4" Drittfarbe: gold Akzentfarbe: firebrick1

generell können auch dunklere Varianten der Farben verwendet werden.
dies ist normalerweise dadurch möglich, dass man die zahl hinden bis 4
erhöht. Dies gilt für alle Farben ausser deepskyblue

**Note zu RQ und Hypos**

Es sollte Möglich sein, die jeweiligen Modell in einer oder beiden
Hypothesen zu verordnen. Sprich: ich untersuche mit meinem Modell H1 und
finde dass mehr reviews den Preis steigen lassen blablabla. Zudem kann
es in gewissen Modellen sinn machen, beide Hypothesen anzuschauen und zu
beantworten. es kann aber auch sein, dass man zb H1 anschaut und mit
Faktoren wie "accommodates" kontrolliert. dann habe ich zwar
location-sided-Indicators, jedoch habe ich keinen Fokus auf die H2.\*\*

Eine dritte Option ist, dass ich nicht eine der beiden Hypothesen
untersuche, sondern die zweite Forshcungsfrage und versuche vorhersagen
zu treffen. dies macht beispielsweise beim SVM sinn. da müsst ihr selbst
enscheiden, welchen Fokus ihr setzen wollt

Coding Conventions. 
• data sets are named with the prefix “d.” (e.g.,
d.cats). 
• linear model objects are named with the prefix “lm.” (e.g.,
lm.cats) 
• plots are named with the prefix "p." 

• within the text R-objects are written using code formatting (e.g., lm.cats) 

• variable names are written using italic (e.g., Body.weight) 
• these conventions
are arbitrary, but hopefully useful to the reader



# Introduction

## Business Case

We propose a data-driven pricing decision-support system for Airbnb
hosts in the Canton of Vaud, providing market-aligned price
recommendations based on listing characteristics. The tool targets both
new and existing hosts, using inputs such as location, room type, guest
capacity, availability, ratings, and superhost status. Six machine learning models, LM, GLM Poisson, GLM Binomial, GAM, SVM and a Neural Network, are applied to identify key price drivers and generate accurate predictions.
The objective is to balance revenue and occupancy, helping hosts
understand their market value. A key feature is interpretability,
enabling users to see why a certain price is recommended. From a
business perspective, the system aims to increase booking rates, reduce
price dispersion, and improve host satisfaction. The model is tailored
to the Canton of Vaud to capture regional economic effects.

## Research Questions

Several questions can be derived from the business case. These questions
are intended to help us examine pricing dynamics and facilitate the
development of a pricing strategy for Airbnb.

-   RQ1 - First, **we need to determine which factors influence guests’
    willingness to pay**. Furthermore, the nature of this influence is,
    of course, central to our analysis.
-   RQ2 - Once these factors are identified, a second question
    immediately arises: **to what extent can these factors be used to
    predict whether the prices of a rental property are justified?**

Willingness to pay is actually the central variable here. However, it is
not easy to measure, which is why a different approach must be taken.
The approach we have chosen stems from the idea that the market
generally functions well and that landlords intuitively already align
with willingness to pay. As a result, the existing price for a rental
property can also be assumed to be the price that tenants are willing to
pay. Any deviations are offset in our subsequent calculations by the
large number of observations. Thus, on average, the actual price will
correspond to the price that tenants are willing to pay.

## Hypothesis

-   H1: When host-centered indicators improve, tenants' willingness to
    pay increases.
-   H2: When location-centered indicators improve, tenants’ willingness
    to pay increases.

Host-centered indicators are those that relate to the host’s
performance, such as ratings or Superhost status. In contrast,
location-centered indicators are those that relate to factors such as
the size of the rental property.

# Data Preparation


In a first step we need to load the relevant packages. following that we load all the datasets that are relevant to us. 


```{r Libraries, results = "hide", echo=TRUE, warning=FALSE}
library(tidyverse)
library(janitor)
library(mgcv)
library(e1071)
library(nnet)
library(pROC)
library(Metrics)
library(performance)
library(broom)
library(kableExtra)
library(stringr)
library(ggplot2)
library(patchwork)
library(knitr)
library(performance)
library(naniar)
library(visdat)
library(neuralnet)
library(leaflet)
library(leaflet.extras)
library(patchwork)

theme_set(theme_minimal())

path_listings <- "./Data_Vaud_all/listings-2.csv"
d.raw_data_listings <- read.csv(path_listings, header = TRUE)
```

Cleaning the Airbnb-Listings Data. We keep only the variables needed for
our project. We reduce the amount of variables to exact 20, according to
the requirements by our lecturers.

```{r cleaning Dataset, echo=TRUE}
d.airbnb_listings <- d.raw_data_listings %>%
  clean_names()

d.listings_clean <- d.airbnb_listings %>%
  mutate(
    price = str_remove_all(price, "\\$|,"),
    price = as.numeric(price),
    host_is_superhost = ifelse(host_is_superhost %in% c("t", "TRUE", "1"), 1, 0),
    room_type = as.factor(room_type),
    property_type = as.factor(property_type),
    host_response_rate = as.numeric(sub("%", "", host_response_rate)),
    host_acceptance_rate = as.numeric(sub("%", "", host_acceptance_rate)),
    host_is_superhost = ifelse(host_is_superhost==1, TRUE, FALSE)
  ) %>%
  filter(!is.na(price), price > 0)




d.listings_clean <- d.listings_clean %>%
  select(-id, -host_id, -host_listings_count, -host_total_listings_count, -listing_url, -scrape_id, -last_scraped,
         -source, -name, -description, -maximum_nights, -first_review, -last_review, -calculated_host_listings_count, -host_since, -availability_365, -estimated_revenue_l365d, 
         -neighborhood_overview, -picture_url, -host_url, -host_name, -host_location,
         -host_about, -host_thumbnail_url, -host_picture_url, -host_neighbourhood,
         -neighbourhood, -neighbourhood_cleansed, -bathrooms_text, -amenities, 
         -minimum_minimum_nights, -maximum_minimum_nights, -minimum_maximum_nights,
         -maximum_maximum_nights, -minimum_nights_avg_ntm, -maximum_nights_avg_ntm,
         -calendar_updated, -has_availability, -availability_30, -availability_60,
         -availability_90, -calendar_last_scraped, -number_of_reviews_ltm, -number_of_reviews_l30d,
         -availability_eoy, -number_of_reviews_ly, -review_scores_accuracy, -review_scores_cleanliness,
         -review_scores_checkin, -review_scores_communication, -review_scores_location,
         -license, -instant_bookable, -calculated_host_listings_count_entire_homes, -calculated_host_listings_count_private_rooms,
         -calculated_host_listings_count_shared_rooms, -reviews_per_month, -bedrooms, -review_scores_value)
```

According to the requirements for this paper, the dataset is limited to
a maximum of 20 variables. To meet this specification, we have removed
columns from the dataset that will not be used in our analysis.
Consequently, `d.listings_clean` is our final, cleaned dataset, which we
will use to run all subsequent models.

Subsequently, we calculated several additional features to include them
as permanent columns in the dataset, ensuring they do not have to be
recomputed for each analysis. These include: the log-transformed
price—as price data is typically heavily right-skewed—the price per
guest (total price divided by the number of guests accommodated), a
boolean indicator for high demand, and a binary variable identifying
whether the price exceeds the median.

```{r initial FE (Pascal), echo=TRUE}
d.airbnb_listings_feat <- d.listings_clean %>%
  mutate(
    log_price = log(price),
    price_per_guest = price / pmax(accommodates, 1),
    high_demand = ifelse(estimated_occupancy_l365d > 200, 1, 0),
    price_high = as.integer(price > median(price, na.rm =TRUE))
  )
```

# Understanding the Data

To inform our modelling choices, we briefly walk through how the data is
structured e.g. with distributions, variable types, and edge cases.
Airbnb already knows this data well, so we keep the commentary short and
focus on what actually matters for the models that follow.

During our exploratory data analysis, we identified one 'Private room'
listing with an unrealistically high price of 2'147'483'647 CHF. We
excluded this single observation from our dataset as it is an extreme
outlier that does not reflect the true market conditions and would
negatively skew our predictive models.

```{r}
# This keeps everything except the row that is a Private room and costs more than 5000
d.airbnb_listings_feat <- d.airbnb_listings_feat %>%
  filter(!(room_type == "Private room" & price > 100000))
```

in addition to the one outlier we wanted to remove one NA specifically

```{r filter NA in ProfilePic}
d.airbnb_listings_feat <- d.airbnb_listings_feat %>%
  filter(!host_has_profile_pic == "")
```

## Prices

```{r, fig.show='hold', out.width="100%", echo=FALSE}
p.price <- ggplot(d.airbnb_listings_feat, aes(x = price)) +
  geom_histogram(bins = 50, fill = "deepskyblue", color = "white", alpha = 0.7) +
  labs(title = "Distribution of Raw Price", x = "Price", y = "Count")

p.logprice <- ggplot(d.airbnb_listings_feat, aes(x = log_price)) +
  geom_histogram(bins = 50, fill = "deepskyblue", color = "white", alpha = 0.7) +
  labs(title = "Distribution of Log Price", x = "Log Price", y = "Count")

p.price | p.logprice
```

Prices are heavily right-skewed. After log-transformation the
distribution is much closer to normal. That is the form we use for the
LM, GAM and most diagnostics later.

## Capacity & Rating

```{r, fig.show='hold', out.width='100%'}
p.acc <- ggplot(d.airbnb_listings_feat, aes(x = accommodates)) +
  geom_histogram(binwidth = 1, fill = "deepskyblue", color = "white") +
  scale_x_continuous(breaks = seq(1, max(d.airbnb_listings_feat$accommodates, na.rm = TRUE), by = 1)) +
  labs(title = "Listing Capacity", x = "Accommodates (Guests)", y = "Count")

p.rev <- ggplot(d.airbnb_listings_feat, aes(x = review_scores_rating)) +
  geom_histogram(bins = 30, fill = "deepskyblue", color = "white", alpha = 0.7) +
  labs(title = "Review Scores Rating Distribution", x = "Rating", y = "Count")

p.acc + p.rev
```

Most listings are built for 2 or 4 guests. Review scores cluster heavily
between 4 and 5. Both hosts and guests rate each other, which probably
produces a leniency bias on top of any genuine quality signal.

## Room Types & Superhost Status

```{r, fig.show='hold', out.width='100%'}
p.room <- ggplot(d.airbnb_listings_feat, aes(x = room_type, fill = room_type)) +
  geom_bar(show.legend = FALSE, fill = "deepskyblue") +
  labs(title = "Proportion of Room Types", x = "Room Type", y = "Count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

p.host <- ggplot(d.airbnb_listings_feat %>% filter(!is.na(host_is_superhost)), aes(x = as.factor(host_is_superhost), fill = as.factor(host_is_superhost))) +
  geom_bar(show.legend = FALSE) +
  scale_fill_manual(values = c("deepskyblue", "deepskyblue")) +
  labs(title = "Superhost Status", x = "Is Superhost?", y = "Count")

p.room + p.host
```

`Entire homes/apartments` and `Private rooms` dominate the inventory;
`Hotel rooms` and `Shared rooms` are marginal. Roughly 30% of hosts are
Superhosts. This is a meaningful minority, but far from the norm.

## Log Prices and rooms / neighbourhoods

```{r, fig.show='hold', out.width='100%'}
p.logprice_room <- ggplot(d.airbnb_listings_feat, aes(x = room_type, y = log_price, fill = room_type)) +
  geom_boxplot(alpha = 0.7, show.legend = FALSE, fill = "deepskyblue") +
  labs(title = "Log Price by Room Type", x = "Room Type", y = "Log Price") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

d.top_neighborhoods <- d.airbnb_listings_feat %>% 
  count(neighbourhood_group_cleansed, sort = TRUE) %>% 
  top_n(10, n) %>% 
  pull(neighbourhood_group_cleansed)

p.logprice_neighbourhood <- ggplot(d.airbnb_listings_feat %>% filter(neighbourhood_group_cleansed %in% d.top_neighborhoods), 
       aes(x = reorder(neighbourhood_group_cleansed, log_price, FUN = median), y = log_price)) +
  geom_boxplot(fill = "deepskyblue", alpha = 0.7) +
  coord_flip() +
  labs(title = "Log Price by Top 10 Neighbourhood Groups", 
       x = "Neighbourhood Group", 
       y = "Log Price")

p.logprice_room | p.logprice_neighbourhood
```

Among room types, hotel rooms post the highest median with a tight distribution. Likely a sign of more professional pricing. Entire homes and private rooms show the same shape but a long upper tail (luxury villas, chalets). Shared rooms are the cheapest option as expected.
Across neighbourhoods, Riviera-Pays-d'Enhaut and Aigle sit at the top, Gros-de-Vaud at the bottom. Almost all top neighbourhoods show high-end outliers, confirming that luxury listings are spread across the canton, though the baseline price (the "floor") varies significantly by location.

## Host Variables

```{r host_features, fig.show='hold', out.width='100%'}
p.resp_time <- ggplot(d.airbnb_listings_feat %>% filter(host_response_time != "N/A" & !is.na(host_response_time)), aes(x = host_response_time)) +
  geom_bar(fill = "deepskyblue") +
  labs(title = "Response Time", x = "Response Time", y = "Count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

p.resp_rate <- ggplot(d.airbnb_listings_feat, aes(x = host_response_rate)) +
  geom_histogram(bins = 20, fill = "deepskyblue", color = "white") +
  labs(title = "Response Rate", x = "Rate (0 to 1)", y = "Count")

p.acc_rate <- ggplot(d.airbnb_listings_feat, aes(x = host_acceptance_rate)) +
  geom_histogram(bins = 20, fill = "deepskyblue", color = "white") +
  labs(title = "Acceptance Rate", x = "Rate", y = "Count")

p.pic <- ggplot(d.airbnb_listings_feat %>% filter(!is.na(host_has_profile_pic)), aes(x = host_has_profile_pic)) +
  geom_bar(fill = "deepskyblue") +
  labs(title = "Profile Pic", x = "t = True, f = False", y = "Count")

p.verified <- ggplot(d.airbnb_listings_feat %>% filter(!is.na(host_identity_verified)), aes(x = host_identity_verified)) +
  geom_bar(fill = "deepskyblue") +
  labs(title = "Identity Verified", x = "t = True, f = False", y = "Count")

p.resp_time | p.resp_rate | p.acc_rate | p.pic | p.verified

```

Hosts in Vaud are highly responsive and reliable: most reply within an hour, response and acceptance rates pile up at 100%, and nearly everyone has a profile picture and verified identity. The downside for modelling: these variables show very little variation, so they probably won't carry much weight as predictors.

## Properties & Neighborhoods

```{r, fig.show='hold', out.width='90%'}

p.prop <- d.airbnb_listings_feat %>%
  count(property_type) %>%
  top_n(10, n) %>%
  ggplot(aes(x = reorder(property_type, n), y = n)) +
  geom_bar(stat = "identity", fill = "deepskyblue") +
  coord_flip() +
  labs(title = "Top 10 Property Types (Count)", x = "Property Type", y = "Count")

p.neigh <- d.airbnb_listings_feat %>%
  count(neighbourhood_group_cleansed) %>%
  top_n(10, n) %>%
  ggplot(aes(x = reorder(neighbourhood_group_cleansed, n), y = n)) +
  geom_bar(stat = "identity", fill = "deepskyblue") +
  coord_flip() +
  labs(title = "Top 10 Neighbourhood Groups (Count)", x = "Neighbourhood", y = "Count")

p.prop | p.neigh
```

Entire rental units dominate the property mix, followed by private rooms in rental units. Specialty types like chalets and villas are present but rare. Geographically, Aigle and Lausanne are the hubs — Aigle alone has over 1,000 listings. In contrast, regions like Gros-de-Vaud show significantly less activity in the top 10 ranking.

## Furnishing

```{r, fig.show='hold', out.width='100%'}
p.baths <- ggplot(d.airbnb_listings_feat, aes(x = bathrooms)) +
  geom_histogram(binwidth = 0.5, fill = "deepskyblue", color = "white") +
  scale_x_continuous(breaks = seq(0, max(d.airbnb_listings_feat$bathrooms, na.rm = TRUE), by = 1)) +
  labs(title = "Number of Bathrooms", x = "Bathrooms", y = "Count")

p.beds <- ggplot(d.airbnb_listings_feat, aes(x = beds)) +
  geom_histogram(binwidth = 1, fill = "deepskyblue", color = "white") +
  scale_x_continuous(breaks = seq(0, max(d.airbnb_listings_feat$beds, na.rm = TRUE), by = 2)) +
  labs(title = "Number of Beds", x = "Beds", y = "Count")

p.baths | p.beds


```

Bathroom and bed counts cluster at the low end as expected. The bed distribution has a few extreme outliers (40+ beds), which could be large hostels or dormitories. We checked a handful manually but didn't verify each one, so they remain in the data as potential noise.

## Utilisation

```{r, fig.show='hold', out.width='100%'}
# Minimum Nights (Log scale due to extreme outliers like 365+ nights)
p.nights <- ggplot(d.airbnb_listings_feat, aes(x = minimum_nights)) +
  geom_histogram(bins = 30, fill = "deepskyblue", color = "white") +
  scale_x_log10() +
  labs(title = "Minimum Nights", x = "Minimum Nights (Log)", y = "Count")

p.num_rev <- ggplot(d.airbnb_listings_feat, aes(x = number_of_reviews)) +
  geom_histogram(bins = 30, fill = "deepskyblue", color = "white") +
  labs(title = "Total Number of Reviews", x = "Reviews", y = "Count")

p.occ <- ggplot(d.airbnb_listings_feat, aes(x = estimated_occupancy_l365d)) +
  geom_histogram(bins = 30, fill = "deepskyblue", color = "white") +
  labs(title = "Estimated Occupancy", x = "Occupied Days", y = "Count")

p.nights | p.num_rev | p.occ
```

The histograms illustrate the distributions of booking requirements,
guest engagement, and listing utilization.

Minimum nights (log scale) shows a clear preference for short stays with a 1 to 3 nights dominance. A handful of listings enforce 365-day minimums, which is why we plot on a log scale.
Review counts are L-shaped: many listings have very few reviews, a small group has hundreds. Estimated occupancy follows the same pattern, with a small spike at the very top end — those are likely commercial operators running close to full capacity year-round.

## Feature Engineered Data

```{r, fig.show='hold', out.width='100%'}

# High Demand (Binary)
p.demand <- ggplot(d.airbnb_listings_feat, aes(x = as.factor(high_demand))) +
  geom_bar(fill = "deepskyblue") +
  labs(title = "High Demand Classification", x = "0 = No, 1 = Yes", y = "Count")

# Price High (Binary)
p.price_high <- ggplot(d.airbnb_listings_feat, aes(x = as.factor(price_high))) +
  geom_bar(fill = "deepskyblue") +
  labs(title = "High Price Classification", x = "0 = No, 1 = Yes", y = "Count")

p.demand | p.price_high
```

High demand is heavily imbalanced: very few listings cross our threshold. Accuracy would be a misleading metric for this target, so for classification on it we'd lean on F1 and precision-recall instead.
High price is balanced by design — we split at the median, so both classes have equal weight. The GLM Binomial uses this target.

## Verifications

```{r, fig.show='hold'}
p.verifications <- d.airbnb_listings_feat %>%
  count(host_verifications) %>%
  top_n(5, n) %>%
  ggplot(aes(x = reorder(host_verifications, n), y = n)) +
  geom_bar(stat = "identity", fill = "deepskyblue") +
  coord_flip() +
  labs(title = "Top 5 Host Verification Combinations", x = "Verification Methods", y = "Count")

p.verifications
```

Most hosts verify themselves with email plus phone — that's the platform baseline. A smaller group adds a work email, which often signals professional or commercial hosts. Profiles with one or zero verifications are very rare.

# Data Quality

We checked the dataset for missing values using the `naniar`
and `visdat` packages. Using `vis_dat()` and `gg_miss_var()`,
we identified that the missing data is not uniformly distributed but
concentrated in specific features. 
The plots below show where NAs sit and how they overlap.
The biggest blocks of missing data are in the rate columns (response, acceptance) and the review score variables. For the rates, we considered treating NAs as zero but decided against it — that would penalise hosts who simply have no reported value rather than a low one.
For review scores, the data is most likely Missing Not At Random: a listing without reviews is usually a new or unbooked listing rather than a poorly-rated one. Imputing zeros would mislabel these as bad listings, and imputing the mean would artificially shrink variance. We therefore drop these observations from the models that need a rating, rather than imputing.


```{r}
gg_miss_var(d.airbnb_listings_feat)
```

## Missing Value Patterns

```{r}
# Calculate percentage of missing values per column
missing_data <- d.airbnb_listings_feat %>%
  summarise(across(everything(), ~ mean(is.na(.)) * 100)) %>%
  pivot_longer(everything(), names_to = "Variable", values_to = "Missing_Percentage") %>%
  filter(Missing_Percentage > 0) %>%
  arrange(desc(Missing_Percentage))

# Plot missing values
p.missing_values <- ggplot(missing_data, aes(x = reorder(Variable, Missing_Percentage), y = Missing_Percentage)) +
  geom_bar(stat = "identity", fill = "deepskyblue") +
  coord_flip() +
  labs(title = "Percentage of Missing Values per Variable",
       x = "Variables",
       y = "% Missing")
p.missing_values
```

The plot below summarises the share of missing values per variable. Anything above ~10% gets explicit handling further down the pipeline.

## NA Check on Key Variables

A quick count of NAs on the variables we use most often in the models below. lm() and svm() silently drop rows with missing values, which can quietly shrink the training set — this check makes sure we know exactly how much.

```{r}
# Count the number of NAs for each key column in your dataset
na_summary <- d.airbnb_listings_feat %>%
  select(price, log_price, accommodates, number_of_reviews, host_is_superhost) %>%
  summarise(across(everything(), ~ sum(is.na(.)))) %>%
  t() %>% # Transpose for better readability
  as.data.frame() %>%
  rename(Missing_Values = V1)

kable(na_summary, caption = "Count of Missing Values per Variable")
```

# Exploratory Data Analysis

## Overall Market Overview (Prices & Capacity)

This is the absolute minimum, maximum, median, and mean for the most
important numeric variables.

```{r}
# Basic summary statistics for key numerical variables
summary_stats <- d.airbnb_listings_feat %>%
  select(price, log_price, accommodates, number_of_reviews) %>%
  summary()

print(summary_stats)
```
The mean price sits well above the median, which means a small number of luxury listings is pulling the average upward. The price range stretches over several orders of magnitude. For any market-level summary we use the median to avoid this skew.

## Analysis by Room Type

Since we want to know how many different types of rooms there are, we
group the data and calculate the counts, percentages, and price ranges
for each type.

```{r}
# Detailed breakdown by Room Type
room_type_analysis <- d.airbnb_listings_feat %>%
  group_by(room_type) %>%
  summarise(
    Total_Listings = n(),
    Percentage_of_Market = round((n() / nrow(d.airbnb_listings_feat)) * 100, 1),
    Min_Price = min(price, na.rm = TRUE),
    Median_Price = median(price, na.rm = TRUE),
    Average_Price = round(mean(price, na.rm = TRUE), 1),
    Max_Price = max(price, na.rm = TRUE)
  ) %>%
  arrange(desc(Total_Listings)) # Sorts by the most common room type

print(room_type_analysis)
```

Entire homes and apartments make up roughly three quarters of the market. Hotel rooms have the highest median at 264.5 CHF, shared rooms the lowest at 47 CHF. The same pattern we already saw in the log-price boxplots above.

## Detailed property_type Analysis

We already analyzed room_type, but Airbnb often has dozens of
property_types (e.g., "Entire chalet", "Private room in castle"...).
Summarizing the Top 10 property types in the Canton of Vaud provides
us with a great market insight.

```{r}
# Top 10 Property Types by frequency and their median price
top_properties <- d.airbnb_listings_feat %>%
  group_by(property_type) %>%
  summarise(
    Total_Listings = n(),
    Median_Price = median(price, na.rm = TRUE)
  ) %>%
  arrange(desc(Total_Listings)) %>%
  slice_head(n = 10) # Keep only the top 10

kable(top_properties, caption = "Top 10 Property Types in the Canton of Vaud")
```

Standard rental units dominate, but specialty properties like chalets, villas, homes command a noticeable premium. The LM and GAM both include `room_type` for exactly this reason.

## Correlation Matrix (Crucial before Machine Learning)

Before we train our LM, GAM, or SVM, we need to know how our numerical
variables relate to each other. This helps us identify multicollinearity
(when two predictors provide the same information) and strong price
drivers.

```{r}
# Select relevant numerical variables (drop NAs just for the correlation calculation)
cor_data <- d.airbnb_listings_feat %>%
  select(log_price, accommodates, number_of_reviews) %>%
  drop_na()

# Calculate the correlation matrix
cor_matrix <- cor(cor_data)

# Render a nice table for your RMarkdown document
kable(round(cor_matrix, 2), caption = "Correlation Matrix of Key Numerical Drivers")
```

```{r correlation_heatmap, out.width="100%", fig.show='hold'}
# Select all numerical variables, drop binaries we already analysed separately
num_vars <- d.airbnb_listings_feat %>%
  select(where(is.numeric)) %>%
  select(-price_high, -high_demand)

# Correlation matrix with pairwise complete observations to handle NAs
cor_matrix_full <- cor(num_vars, use = "pairwise.complete.obs")

# Reshape for ggplot
cor_long <- as.data.frame(cor_matrix_full) %>%
  rownames_to_column("Var1") %>%
  pivot_longer(-Var1, names_to = "Var2", values_to = "Correlation")

ggplot(cor_long, aes(x = Var1, y = Var2, fill = Correlation)) +
  geom_tile(color = "white") +
  geom_text(aes(label = round(Correlation, 2)), size = 2.8) +
  scale_fill_gradient2(low = "maroon4", mid = "deepskyblue", high = "gold",
                       midpoint = 0, limits = c(-1, 1)) +
  labs(title = "Correlation Heatmap of Numerical Variables",
       x = NULL, y = NULL) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

The heatmap confirms what we already suspected: `accommodates`, `beds` and `bathrooms` *correlate strongly with each other*. A strong sign of multicollinearity to watch for in the LM (see LM Chapter). `price` and `log_price` track capacity as expected; review variables form their own cluster. `number_of_reviews` and `estimated_occupancy_l365d` move together, because listings that get booked more naturally collect more reviews.

## Log Price vs. Accommodates

The most important Relationship in the dataset from what we see in the
correlation matrix above.

```{r price_vs_accommodates, out.width="100%", fig.show='hold'}
ggplot(d.airbnb_listings_feat, aes(x = accommodates, y = log_price)) +
  geom_jitter(alpha = 0.3, width = 0.2, color = "deepskyblue") +
  geom_smooth(method = "lm", color = "firebrick1", se = TRUE) +
  scale_x_continuous(breaks = seq(1, max(d.airbnb_listings_feat$accommodates, na.rm = TRUE), 2)) +
  labs(title = "Log Price vs. Accommodates",
       x = "Number of Guests",
       y = "Log Price")
```

*Capacity* is the strongest single predictor in the dataset, and the relationship looks roughly linear on the log scale, which is exactly why we work with `log(price)` throughout the LM and GAM.

## Log Price by Review Score Group

We display `review_scores_rating` as a histogram, but never in relation to
the price.

```{r price_by_rating, out.width="100%", fig.show='hold'}
d.airbnb_listings_feat %>%
  filter(!is.na(review_scores_rating)) %>%
  mutate(rating_group = cut(review_scores_rating,
                            breaks = c(0, 3, 4, 4.5, 4.8, 5),
                            labels = c("≤3", "3–4", "4–4.5", "4.5–4.8", "4.8–5"),
                            include.lowest = TRUE)) %>%
  ggplot(aes(x = rating_group, y = log_price, fill = rating_group)) +
  geom_boxplot(alpha = 0.7, show.legend = FALSE, fill = "deepskyblue") +
  labs(title = "Log Price by Review Score Group",
       x = "Review Score Rating",
       y = "Log Price")
```

Medians across rating groups are surprisingly close. Higher-rated listings do not automatically charge more. A first hint at the Superhost pattern we'll come back to: in this market, reputation correlates with occupancy strategies more than with price premiums.

## Spatial Data: Mapping the Listings by Log(Price)

```{r, out.width="100%", fig.show='hold'}
# Cleaned up the map to focus purely on location and price.
ggplot(data = d.airbnb_listings_feat, aes(x = longitude, y = latitude)) +
  geom_point(aes(color = log_price), alpha = 0.6, size = 1.5) +
  scale_color_continuous(palette = c("deepskyblue", "firebrick1", "gold")) +
  labs(
    title = "Airbnb Listings in Vaud by Log(Price)",
    x = "Longitude",
    y = "Latitude",
    color = "Log(Price)"
  ) +
  theme_minimal()
```

Hot spots cluster along Lake Geneva and around Lausanne, with rural and alpine areas trailing off. This is exactly the surface the GAM picks up later with its 2D spline — and the reason we use raw coordinates rather than a postcode dummy.

## Price per Guest Distribution

We calculated price per guest as a feature then visualized them.

```{r price_per_guest, out.width="100%", fig.show='hold'}
p_ppg <- ggplot(d.airbnb_listings_feat, aes(x = price_per_guest)) +
  geom_histogram(bins = 50, fill = "deepskyblue", color = "white", alpha = 0.7) +
  scale_x_log10() +
  labs(title = "Price per Guest (Log Scale)",
       x = "Price per Guest (CHF)", y = "Count")

p_ppg_room <- ggplot(d.airbnb_listings_feat, aes(x = room_type, y = price_per_guest, fill = room_type)) +
  geom_boxplot(alpha = 0.7, show.legend = FALSE, fill = "deepskyblue") +
  scale_y_log10() +
  labs(title = "Price per Guest by Room Type",
       x = "Room Type", y = "Price per Guest (Log)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

p_ppg + p_ppg_room
```

Most listings sit between 30 and 100 CHF per guest. Hotel rooms charge the highest per-head price; shared rooms are clearly the cheapest. Entire homes look expensive overall but actually offer the best value once you split the cost across guests.

## Superhost vs. Regular Host Breakdown

This business case specifically mentions superhost status. So we took a
look to see how many there are and if they actually charge more or even
less.

```{r}
# Superhost statistics
superhost_analysis <- d.airbnb_listings_feat %>%
  group_by(host_is_superhost) %>%
  summarise(
    Total_Hosts = n(),
    Percentage = round((n() / nrow(d.airbnb_listings_feat)) * 100, 1),
    Median_Price = median(price, na.rm = TRUE),
    Avg_Reviews = round(mean(number_of_reviews, na.rm = TRUE), 1)
  ) %>%
  mutate(host_is_superhost = ifelse(host_is_superhost == 1, "Yes", "No"))

print(superhost_analysis)
```

At first glance the boxplot shows almost no difference in median price between Superhosts and regular hosts. This is our first sign that the Superhost badge does not translate into a price premium — a pattern that the GLM Binomial, the GAM and the Neural Network will all confirm independently in the next chapters.

```{r Binary Predictor: Superhost Status, out.width="100%", fig.show='hold'}
# Adding this because it's a key part of the business case and fulfills the binary variable requirement.
ggplot(data = d.airbnb_listings_feat, aes(x = as.factor(host_is_superhost), y = log_price, fill = as.factor(host_is_superhost))) +
  geom_boxplot(alpha = 0.7, fill = "deepskyblue") +
  scale_x_discrete(labels = c("0" = "Regular Host", "1" = "Superhost")) +
  labs(
    title = "Price Difference: Superhosts vs. Regular Hosts",
    x = "Host Status",
    y = "Log(Price)"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

# Modelling Approach

We build six models, each looking at price or its proxy, occupancy from a different angle. The point isn't to pick one winner. It's to triangulate. When models with very different mechanics converge on the same answer, we can trust it more than any single result.

## Models and their roles

| Model | Question | Hypothesis | Output |
|---|---|---|---|
| Linear Model | Which variables shift price linearly? | H1 | Coefficients |
| GLM Poisson | Which variables drive *occupancy*? | H1 + H2 | Rate ratios |
| GLM Binomial | Who ends up in the upper price half? | H1 | Odds ratios |
| GAM | What does the non-linear price surface look like? | H2 (with H1 as control) | Smooth curves + fair price per listing |
| SVM | Is a given listing under-, fair-, or overpriced? | RQ2 (prediction) | Classification verdict |
| ANN | Does any host effect survive flexible controls? | H1 | Predicted prices for hypothetical hosts |

## How the models connect

The **LM** is our baseline — anything more complex needs to clearly beat it. The **GLM Poisson** runs the same question on a different target (occupancy instead of price), which lets us spot situations where a variable affects one but not the other. The **GLM Binomial** reduces the price question to a yes/no decision (above or below median), which makes the host effect cleaner to read.

The **GAM** is the strongest single model in this study. It captures the non-linear effects we already saw in the EDA and adds a 2D spatial term that no other model has. The **SVM** then operationalises the GAM into a working tool — given a listing and a chosen price, it returns Underpriced, Fair, or Overpriced.

The **ANN** doesn't add new information so much as it checks our assumptions. If a flexible neural network barely beats the GAM, the GAM is already capturing most of the available signal — and we can trust the interpretable model over the black box.

## Train/Test discipline

The GAM, SVM and ANN are evaluated on held-out test data. The split is made **once** — on the cleaned dataset, before any model fits — and reused throughout. This matters especially for the SVM: its training labels come from the GAM's fair-price band, so if we split after labelling, GAM information would leak into the SVM's test set. We split first, fit the GAM only on the training set, then derive labels.


# Linear Model

## Theory

With this initial model, we aim to examine the first hypothesis if host-centered indicators shift price?

For this, we use the *review_scores_rating*. We also include the factor
*host_is_superhost* to distinguish between the two groups. We measure
willingness to pay here using the effective price of the rental
property. Additionally, we introduce two control variables. First, the
size of the rental property, as this obviously influences the rental
price, and second, the type of room. This also significantly influences
the price.

A linear model allows us to analyze multiple variables and their effect
on a dependent variable. It is important that the dependent variable be
normally distributed. Since this is not the case here, as is often the
case with price variables, this variable is logarithmized. We assumed that there is a risk of high collinearity in our data. Therefore we checked for it, to make sure, that the results are robust. 

## Model

```{r lm lead Timon}

# We use log(price) to account for the right-skewed distribution of Airbnb prices.
lm.main <- lm(log(price) ~ 
                review_scores_rating + 
                as.factor(host_is_superhost) +
                host_has_profile_pic + 
                accommodates + 
                as.factor(room_type), 
              data = d.airbnb_listings_feat)

tidy(lm.main) %>%
  kable(format = "latex", digits = 2, booktabs = TRUE, caption = "Model Output") %>%
  kable_styling(font_size = 8, latex_options = "scale_down")



# Checking for multicollinearity using Variance Inflation Factor (VIF).
# Values range from 1 to infinity. A value of exactly 1 indicates no collinearity.
# Values > 5-10 suggest problematic multicollinearity among the predictor variables.
multicollinearity(lm.main)
```

## Visualisation

```{r, out.width="100%", fig.show='hold'}

# plot log-price and scores + superhost
p.lm_uncontrolled <- ggplot(data = d.airbnb_listings_feat, mapping = aes(review_scores_rating, log(price), color = host_is_superhost), ylim = 8) +
  geom_point(alpha = 0.4, color = "deepskyblue") +
  geom_smooth(method = "lm") +
  scale_color_manual(values = c("FALSE" = "gold2", "TRUE" = "firebrick1")) +
    labs(title = "Uncontrolled LM",
       x = "Review Scores Rating",
       y = "Log(Price)")+
  theme(legend.position = "none" )+
  ylim(3,8)

# creating grid for hypothetical data holding 'room_type' and 'accommodates' constant 
d.lm_predict <- expand.grid(
  review_scores_rating = seq(3, 5, length.out = 100),
  host_is_superhost = c("TRUE", "FALSE"),
  room_type = "Entire home/apt",   # Fixed reference category
  host_has_profile_pic = c("t", "f"),
  accommodates = mean(d.airbnb_listings_feat$accommodates, na.rm = TRUE) # Fixed at mean value
)

# Generating predictions
d.lm_predict$pred <- predict(lm.main, newdata = d.lm_predict)

# Plotting the actual data points alongside our model's predicted regression lines
p.lm_controlled <-ggplot(d.airbnb_listings_feat, 
       aes(review_scores_rating, log(price), color = host_is_superhost, )) +
  geom_point(alpha = 0.3, color = "deepskyblue") +
  geom_line(data = d.lm_predict, aes(y = pred)) +
  scale_color_manual(values = c("FALSE" = "gold2", "TRUE" = "firebrick1")) +
  labs(title = "Controlled LM",
       x = "Review Scores Rating",
       y = "Log(Price)") +
    ylim(3,8)
  
(p.lm_uncontrolled | p.lm_controlled) +
  plot_layout(widths = c(10,10))
```

## Interpretation

The model shows that both the hosts’ rating and their Superhost status
have a significant influence. The rating has a positive effect of
*`r round(lm.main$coefficients[2], digits = 3)`*. In contrast, Superhost
status has a negative effect of
*`r round(lm.main$coefficients[3], digits = 3)`*. The assumption that
tenants are willing to pay more when renting from a Superhost therefore
does not necessarily seem to hold true. One possible explanation for
this phenomenon is that Superhosts may have a better understanding of
the market because they have been active for a longer time, and
therefore set lower prices in order to achieve a higher occupancy rate.
This is relevant for Airbnb, as these results show that reviews on the
host side are a key factor in price setting. It is also interesting that
Superhosts might develop their own understanding for how prices can
be set. The exact reasons underlying this difference cannot be examined
within the scope of this paper.
Through this analysis we can also answer the first hypothesis. Host centered Variables, in this case the *review_scores_rating* and *host_is_superhost*, do change the price of a venue.

## Limitations
A linear model often comes with certain limitations, as the degree of complexity is limited. We can only observe linear effects. We can assume that capacity is not a linear effect, since the price of a location can vary depending on whether it accommodates 3 or 4 people, whereas the effect of one fewer person is less significant if the location accommodates 15 or 16 people. A GAM model could reveal such effects. 
Another limitation is the absence of the “location” variable. While we could include latitude and longitude, these variables alone do not have a significant impact on price. For example, we would need the distance to the nearest city center to measure this accurately. 

# GLM Poisson

## Theory

With the GLM model with Poisson-family / quasi-poisson family is used to
analyse count data. Count data is strictly positive and counted (not
continuous). Using this model we want to see which variables lead to the
highest occupancy, meaning most of the time the airbnb is rented out. As
the data is probably overdispersed, meaning the data has not the same
variance and much more cheap listings than expensive ones, we need to
use a quasi-poisson model that can handle this kind of data much better.
First we fit a normal poisson-family, then check for overdispersion, and
if necessary we use a quasi-poisson-family.

## Link to Hypothesis
This is the only model in our set that targets occupancy rather than price. Both hypotheses are in scope — host-centered indicators (H1) and structural indicators (H2) can both drive how often a listing gets booked. The answer turns out to be the missing piece of the Superhost puzzle the other models will surface.

## Model

```{r, GLM Poisson Lead Pascal}

glm.poisson <- glm(estimated_occupancy_l365d ~ price + host_is_superhost + 
                     review_scores_rating + accommodates + room_type, 
                   data = d.airbnb_listings_feat, 
                   family = "poisson")
# Dispersion check
dispersion_ratio <- glm.poisson$deviance / glm.poisson$df.residual

glm.quasipoisson <- glm(estimated_occupancy_l365d ~ price + host_is_superhost + 
                          review_scores_rating + accommodates + room_type, 
                        data = d.airbnb_listings_feat, 
                        family = "quasipoisson")

tidy.poisson <- tidy(glm.poisson) %>% mutate(Model = "Poisson")
tidy.quasi <- tidy(glm.quasipoisson) %>% mutate(Model = "Quasi-Poisson")

combined_models <- bind_rows(tidy.poisson, tidy.quasi) %>%
  select(Model, term, estimate, std.error, statistic, p.value)

combined_models %>%
  kable( 
        digits = 3, 
        booktabs = TRUE, 
        caption = paste0("Comparing Models (Dispersion Ratio: ", round(dispersion_ratio, 2), ")")) %>%
  kable_styling(font_size = 8, latex_options = c("scale_down", "hold_position")) %>%
  collapse_rows(columns = 1, valign = "top")

```

## Crossvalidation

```{r cross-validation Datasets 10-fold (TIMON)}
set.seed(1)

cv.mse.poisson <- c()
cv.mse.quasipoisson <- c()

for(i in 1:200){
  #dataprep
  n <- nrow(d.airbnb_listings_feat)
  cv.ind_permuted <- sample(1:n, replace = FALSE)
  d.airbnb_listings_feat_permuted <- d.airbnb_listings_feat[cv.ind_permuted,]
  
  #create folds
  cv.folds <- cut(1:n, breaks = 3, labels = FALSE)
  cv.ind_test <- which(cv.folds == 1)
  
  d.cv_test <- d.airbnb_listings_feat_permuted[cv.ind_test, ]
  d.cv_train <- d.airbnb_listings_feat_permuted[-cv.ind_test, ]
  
  #train model poisson
  cv.glm_poisson_train <- glm(formula = formula(glm.poisson), 
                          data = d.cv_train)
  cv.glm_poisson_predict <- predict(cv.glm_poisson_train, newdata = d.cv_test)
  
  cv.mse.poisson[i] <- mean((d.cv_test$estimated_occupancy_l365d - cv.glm_poisson_predict)^2, 
                           na.rm = TRUE)
  
  
  #train model quasipoisson
  cv.glm_quasipoisson_train <- glm(formula = formula(glm.quasipoisson), 
                          data = d.cv_train)
  cv.glm_quasipoisson_predict <- predict(cv.glm_quasipoisson_train, newdata = d.cv_test)
  
  cv.mse.quasipoisson[i] <- mean((d.cv_test$estimated_occupancy_l365d - cv.glm_quasipoisson_predict)^2, 
                           na.rm = TRUE)
  
}

boxplot(cv.mse.poisson, cv.mse.quasipoisson, names = c("Poisson-Model","Quasipoisson-Model"), col = "deepskyblue")

```

## Interpretations

### Interpretation GLM-Poisson

As expected, the GLM with a Poisson distribution produced
overdispersion, since the variability in the Airbnb data was to be
expected. The model calculated the standard errors too small, which is
why all variables were highly significant in the first output. The
dispersion was then calculated and found to be greater than 1,
specifically `r round(dispersion_ratio, 2)`. A quasi-Poisson model was used as a solution, which
adjusts the standard errors upward.

### Interpretation GLM-Quasipoisson

The GLM with quasi-Poisson distribution has corrected the
overdispersion. The standard errors were adjusted upwards
(significantly) and now provide a realistic picture of which factors
truly have an influence on occupancy. Interpretations: - Price -\>
highly significant negative. As expected: The higher the price, the
lower the occupancy - Superhost: highly significant positive: Attention,
logged rates, therefore calculate the effect with the exponential
function exp(0.9577) = 2.6. Being a Superhost increases occupancy by 2.6
times compared to a normal host. - Reviews: Positive and significant.
Better reviews lead to more bookings. - Capacity: Positive and
significant. The more people that fit, the more bookings. - Room types:
In the intercept is "Entire home/apt". Private Room is booked
significantly worse than the entire accommodation. Shared Room has a
high p-value. Statistically, it cannot be said whether the occupancy of
a shared room differs from the entire apartment.

### Interpretation Cross-validation

To ensure that the higher number of variables does not cause
overfitting, we use cross-validation. In this process, we repeatedly
split the data into new subsets, which we use to retrain and re-evaluate
the model. We then check to see if there are any significant
fluctuations. If the model were overfitting, we would repeatedly observe
random fluctuations. In our case, however, we only get a single outlier,
which suggests that we have a stable model without overfitting.


## Limitations

A primary limitation of this model is the potential for reverse
causality, as higher occupancy naturally generates more reviews, making
it easier to maintain high ratings and Superhost status. Additionally,
the deletion of 714 observations due to missingness could introduce
survivorship bias if these removed data points represent poorly managed
or unsuccessful listings. Furthermore, our feature set omits critical
demand drivers like precise neighborhood desirability, seasonality, and
listing photo quality, meaning unobserved factors likely account for
significant residual variance. A potential next step could be exploring a Negative
Binomial model as an alternative method to handle overdispersion while
retaining standard likelihood metrics for easier comparison.

# GLM Binomial

## Theory

When the outcome we want to model is binary — yes/no, above/below — a standard linear regression no longer fits. A Binomial GLM with a logit link is the natural choice: it bounds predictions between 0 and 1 and returns coefficients we can read as log-odds. Exponentiating gives odds ratios, which are easier to communicate to a non-technical audience.
Our binary target `price_high` equals 1 if a listing's price is above the median and 0 otherwise. The median split gives us two equally sized classes, so the model is not biased toward a majority outcome.

## Link to Hypothesis

This model speaks mainly to H1: do host-centered indicators (Superhost
status, ratings) push up the price? The control variables
(*accommodates*, *room_type*) are not the focus but make sure we don't
confuse host effects with listing characteristics.

## Model

Feature Engeneering

```{r glm_binomial_extended, Lead Davide}
# Predict the probability that a listing is priced above the market median
# We control for the most obvious price drivers so the host coefficients can be read cleanly
glm.binomial <- glm(price_high ~ host_is_superhost + 
                                 accommodates + 
                                 room_type + 
                                 review_scores_rating,
                    data   = d.airbnb_listings_feat,
                    family = binomial(link = "logit"))

summary(glm.binomial)
```

```{r glm_binomial_results, Lead Davide}
# Convert log-odds into odds ratios with 95% confidence intervals for easier reading
d.results_glm_binomial <- tidy(glm.binomial, conf.int = TRUE) %>%
  mutate(
    odds_ratio = round(exp(estimate), 3),
    or_lower   = round(exp(conf.low), 3),
    or_upper   = round(exp(conf.high), 3),
    p.value    = round(p.value, 4)
  ) %>%
  select(term, estimate, odds_ratio, or_lower, or_upper, p.value)

kable(d.results_glm_binomial, 
      caption = "GLM Binomial: Odds Ratios with 95% Confidence Intervals")
```

## Visualisation

A coefficient table is hard to read for hosts. The plot below shows the
predicted probability of charging an above-median price across listing
sizes, split by Superhost status. Everything else is held at its average
so the two lines differ only in the variable we care about.

```{r glm_binomial_plot, Lead Davide}
# Build a prediction grid: vary accommodates and superhost, hold the rest constant
d.glm_predict <- expand.grid(
  accommodates         = seq(1, 10, by = 1),
  host_is_superhost    = c(TRUE, FALSE),
  room_type            = "Entire home/apt",
  review_scores_rating = mean(d.airbnb_listings_feat$review_scores_rating, na.rm = TRUE)
)

# Predict on the link scale to also get standard errors, then transform back
glm.pred <- predict(glm.binomial, newdata = d.glm_predict, type = "link", se.fit = TRUE)
d.glm_predict$prob       <- plogis(glm.pred$fit)
d.glm_predict$prob_lower <- plogis(glm.pred$fit - 1.96 * glm.pred$se.fit)
d.glm_predict$prob_upper <- plogis(glm.pred$fit + 1.96 * glm.pred$se.fit)

p.glm_binomial <- ggplot(d.glm_predict, aes(x = accommodates, y = prob, 
                                            color = host_is_superhost, fill = host_is_superhost)) +
  geom_ribbon(aes(ymin = prob_lower, ymax = prob_upper), alpha = 0.2, color = NA) +
  geom_line(size = 1.2) +
  scale_x_continuous(breaks = 1:10) +
  scale_color_manual(values = c("gold", "firebrick1"),
                     labels = c("Regular Host", "Superhost")) +
  scale_fill_manual(values = c("gold", "firebrick1"),
                    labels = c("Regular Host", "Superhost")) +
  labs(title    = "Probability of Charging an Above-Median Price",
       subtitle = "By guest capacity and Superhost status (entire home, average rating)",
       x        = "Accommodates (Guests)",
       y        = "Predicted Probability",
       color    = NULL, fill = NULL) +
  theme(legend.position = "top")

p.glm_binomial
```

## Interpretation GLM Binomial

After controlling for size, room type and rating, Superhosts have **31%
lower odds** of charging an above-median price (OR = 0.69, CI [0.57,
0.84], p = 0.0002).

The control variables behave as expected: each additional guest nearly
doubles the odds of being in the upper price half (OR = 1.83), private
and shared rooms are far less likely to land above the median, and an
one-point increase in rating raises the odds by about 62%. The
hotel-room coefficient is significant but its wide confidence interval
(1.26 to 40.6) reflects how few hotel listings are in the dataset.

This is the opposite of what H1 would predict. Superhost status does not
raise willingness to pay, it correlates with charging less. Combined
with Pascal's Quasi-Poisson result (Superhosts get 2.6× more bookings),
the picture is consistent: Superhosts compete on price and earn through
volume, not through a price premium. The practical takeaway for hosts
aiming for the badge is to price competitively and let occupancy do the
work.

## Limitations

A few things to keep in mind:

-   Median split is rough. A listing one franc above the median counts
    the same as one 500 francs above. The LM and GAM handle this much
    better.
-   No location. We left location out to keep the focus on host
    variables, but in Vaud, location is a major price driver. Some of
    what we attribute to room type or capacity is probably picking up
    location indirectly.
-   Snapshot only. No seasonality, no weekday effects, no events.
-   Correlation, not causation. Superhosts charge less on average but we
    can't say the badge causes it. It's just as likely that hosts who
    already price competitively end up qualifying for the badge.
-   The cutoff moves with the market. The median is just our sample's
    median. If the market shifts, so does it.

# GAM

## Theory

**Why GAM for Airbnb Pricing?** A standard linear model assumes that
every predictor moves the price at a constant rate. For Airbnb data
that's almost never true. A Generalized Additive Model (GAM) replaces
the fixed coefficients with smooth functions that the model learns from
the data itself. Each predictor gets its own curve instead of a single
slope, which lets the model bend where the data actually bends.

The result sits between a linear regression and a black-box model: more
flexible than the LM, more interpretable than a Random Forest or Neural
Network. Each smooth term can still be plotted on its own and read like
a coefficient.

Three things in our data make a GAM the right fit:

1.  Price effects are non-linear. Going from 2 to 4 guests roughly
    doubles the price; going from 8 to 10 barely moves it. Review scores
    show a similar pattern. A drop from 4.0 to 3.5 hurts much more than
    the difference between 4.8 and 5.0. A linear model averages these
    curves and gets both ends wrong.
2.  Location is non-linear by nature. Vaud combines lakeshore towns,
    urban centres, alpine resorts, and rural areas, each with their own
    price logic. A 2D smoother over longitude and latitude captures this
    surface without us having to engineer postcode dummies or
    distance-to-lake variables. This is the main reason we picked GAM
    over LM.
3.  Interpretability stays. Hosts should be able to see why a price is
    recommended, not just get a number from a black box. The partial
    effect plots make every smooth term readable on its own.

## Link to Hypothesis

The GAM speaks mainly to **H2** through the spatial spline *s(longitude,
latitude)*, while controlling for **H1** via *host_is_superhost* and
*review_scores_rating*. It is our main tool for checking whether
location effects on willingness to pay are non-linear and clustered,
something a linear model cannot answer.

## Model

```{r gam_data_split, Lead Davide}
# Data preparation: keep only variables used in the GAM, drop rows with NAs
d.airbnb_listings_feat_gam <- d.airbnb_listings_feat %>%
  select(price, accommodates, review_scores_rating,
         number_of_reviews, longitude, latitude,
         room_type, host_is_superhost, bathrooms) %>%
  na.omit()

# Train/test split BEFORE fitting the model.
# This matters for the SVM later, which uses GAM labels — splitting first prevents leakage.
set.seed(42)
n <- nrow(d.airbnb_listings_feat_gam)
train.idx <- sample(seq_len(n), size = floor(0.8 * n))

d.train <- d.airbnb_listings_feat_gam[ train.idx, ]
d.test  <- d.airbnb_listings_feat_gam[-train.idx, ]

cat("Train:", nrow(d.train), " | Test:", nrow(d.test), "\n")
```

We model *log(price)* rather than the raw price. The price distribution
is heavily right-skewed (see EDA-Part), and the log transform stabilises
variance and brings the residuals close to normal. Coefficients on the
log scale can be read multiplicatively: a coefficient of β corresponds
to roughly a $(exp(β) − 1) × 100%$ price change. We use *method =
"REML"* instead of the default. REML is less prone to undersmoothing and
gives more reliable confidence intervals

```{r gam_fit, Lead Davide}
gam.main <- gam(log(price) ~
                  s(accommodates) +
                  s(review_scores_rating) +
                  s(log1p(number_of_reviews)) +
                  s(longitude, latitude) +
                  room_type +
                  host_is_superhost +
                  bathrooms,
                family = gaussian(),
                method = "REML",
                data   = d.train)

summary(gam.main)
```

## Visualisation

The partial effect plots are the main reason to choose a GAM in the
first place. Each smooth term is shown on its own, with the shaded area
representing the 95% confidence band.

```{r gam_partial_effects, Lead Davide}
plot(gam.main, pages = 1, shade = TRUE,
     col = "firebrick1", shade.col = "lightblue", seWithMean = TRUE)

d.table <- as.data.frame(summary(gam.main)$s.table) %>% round(3)
kable(d.table, caption = "Smooth Terms: Estimated Degrees of Freedom (EDF)")

cat("R-squared:         ", round(summary(gam.main)$r.sq,  3), "\n")
cat("Deviance explained:", round(summary(gam.main)$dev.expl, 3), "\n")
```

To make the comparison fair, we fit a linear model on the exact same
data, target, and predictors. The only difference is the absence of
smooth terms.

```{r gam_lm_benchmark, Lead Davide}
# Fair LM benchmark (same target, same training data)
lm.gam_comparison <- lm(log(price) ~ accommodates + review_scores_rating +
                          log1p(number_of_reviews) + longitude + latitude +
                          room_type + host_is_superhost + bathrooms,
                        data = d.train)

cat("LM  R²:", round(summary(lm.gam_comparison)$r.squared, 3), "\n")
cat("GAM R²:", round(summary(gam.main)$r.sq, 3), "\n")
AIC(lm.gam_comparison, gam.main)
```

### Traffic Light Map

The GAM predicts an expected price for every listing. Comparing actual
to expected gives hosts a quick verdict — underpriced, fair, or
overpriced. We set the threshold at ±15%, anchored in the residual
standard deviation of the GAM on the log scale (exp(σ) − 1 ≈ 15%).
Listings outside this band deviate by more than one standard deviation
from the model's expected price. The threshold is data-driven, but it
remains a choice. A stricter or looser band would shift the proportions
accordingly. This is one of the limitations we discuss below.

```{r traffic_light_labels}
# Predictions on log scale
d.train$fitted_logprice <- as.numeric(predict(gam.main, newdata = d.train))
d.test$fitted_logprice  <- as.numeric(predict(gam.main, newdata = d.test))

# Smearing correction (Duan 1983) for back-transforming log predictions
sigma2   <- summary(gam.main)$scale
smearing <- exp(sigma2 / 2)
d.train$fitted_price <- exp(d.train$fitted_logprice) * smearing
d.test$fitted_price  <- exp(d.test$fitted_logprice)  * smearing

# Apply traffic-light rule consistently to train and test
make.traffic_light <- function(price, fitted, threshold = 15) {
  pct <- (as.numeric(price) - as.numeric(fitted)) / as.numeric(fitted) * 100
  case_when(
    pct <= -threshold ~ "Underpriced",
    pct >=  threshold ~ "Overpriced",
    TRUE              ~ "Fair Price"
  ) %>% as.factor()
}

d.train$pct_diff      <- (d.train$price - d.train$fitted_price) / d.train$fitted_price * 100
d.test$pct_diff       <- (d.test$price  - d.test$fitted_price)  / d.test$fitted_price  * 100
d.train$traffic_light <- make.traffic_light(d.train$price, d.train$fitted_price)
d.test$traffic_light  <- make.traffic_light(d.test$price,  d.test$fitted_price)

# Sanity check: distribution should look similar across train and test
cat("\nTrain:\n"); print(prop.table(table(d.train$traffic_light)))
cat("\nTest:\n");  print(prop.table(table(d.test$traffic_light)))
```

### Interactive Map of Vaud

The map below plots every listing in the dataset, coloured by its
traffic-light verdict. Each point shows the actual price, the
GAM-recommended price, and the percentage deviation when clicked.

```{r gam_leaflet_map}
d.map <- bind_rows(d.train, d.test)
color_map <- c("Underpriced" = "#EF9F27",
               "Fair Price"  = "#1D9E75",
               "Overpriced"  = "#D85A30")
d.map$marker_color <- color_map[as.character(d.map$traffic_light)]

leaflet() %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  setView(lng = 6.65, lat = 46.52, zoom = 9) %>%
  addCircleMarkers(
    data = d.map,
    lng = ~longitude, lat = ~latitude,
    radius = 4, fillColor = ~marker_color,
    fillOpacity = 0.85, stroke = FALSE,
    popup = ~paste0(
      "<b>Host Verdict:</b> ", traffic_light, "<br>",
      "<b>Actual Price:</b> CHF ", round(price), "<br>",
      "<b>GAM Recommended:</b> CHF ", round(fitted_price), "<br>",
      "<b>Difference:</b> ", round(pct_diff, 1), "%"
    )
  ) %>%
  addLegend(
    position = "bottomright",
    colors   = color_map,
    labels   = c("Underpriced (Increase)", "Fair Price (±15%)", "Overpriced (Decrease)"),
    title    = "Pricing Strategy vs. Market (GAM)",
    opacity  = 0.85
  )

# Note: d.train and d.test are passed on to the SVM block.
```

## Interpretation GAM

The GAM explains 66.4% of the deviance (adjusted R² = 0.658) on 2,527
training listings. That's a clear improvement over the LM benchmark (R²
= 0.607) on the same data, and the AIC drops from 2,599 to 2,279. A
substantial gain that confirms the smooth terms earn their additional
degrees of freedom.

Parametric terms Room type is the strongest categorical driver. Compared
to entire homes, hotel rooms charge roughly 1.9× more, private rooms are
about 30% cheaper, and shared rooms drop another step to around 55%
below entire homes. The hotel-room result rests on a small number of
listings, so we shouldn't over-read it. Bathrooms show a clean linear
effect. Each additional bathroom adds roughly 25% to the price. Besides
room type and capacity, this is the strongest structural driver in the
model. Superhost status is the most telling result. After controlling
for everything else, the coefficient is essentially zero and not
significant. The badge has no measurable effect on the nightly price.
This lines up with the GLM Binomial and the Neural Network. Three very
different models converge on the same answer.

Smooth terms All four smooth terms are highly significant (p \< 0.001)
and every EDF is well above 1, which means a linear approximation would
have been wrong for each of them.

-   *s(accommodates)* (EDF = 4.58): the curve rises steeply from 1 to
    about 6 guests and then flattens. Doubling capacity from 8 to 16
    adds far less than doubling from 2 to 4, which is exactly the
    diminishing-returns pattern we expected.
-   *s(review_scores_rating)* (EDF = 3.00): the effect is real but
    modest. The curve trends slightly upward across the range, with a
    wide confidence band — ratings matter, but they're not a major price
    lever.
-   *s(log1p(number_of_reviews))* (EDF = 2.63): a slight downward trend.
    Listings with more reviews tend to be priced lower — consistent with
    the volume-over-margin pattern we keep seeing.
-   *s(longitude, latitude)* (EDF = 24.93): by far the most flexible
    term in the model. The 2D surface shows clear high-price clusters
    along the Lake Geneva shore and around Lausanne, with sharper drops
    in rural and alpine areas. This single term is the strongest
    empirical support for H2 — location is a major, non-linear driver of
    willingness to pay.

Traffic-light distribution The traffic-light split is consistent across
train and test (Fair ≈ 33%, Underpriced ≈ 41%, Overpriced ≈ 25%), so the
model generalises well. The slight overweight of "Underpriced" listings
is interesting: it suggests that a notable share of hosts are leaving
money on the table compared to what the model considers fair given their
listing's features and location.

Take-aways for the pricing tool: 1. Structural features come first. Room
type, bathrooms, and capacity together account for most of the
explainable price variation that hosts can directly influence. The
pricing interface should highlight these. 2. Location effects are real
but non-linear. Coordinates as a continuous input clearly outperform
postcode dummies — the spatial spline alone justifies the GAM approach
for a regional tool. 3. The Superhost badge is an occupancy strategy,
not a pricing one. Hosts aiming for the badge should price competitively
and let bookings build up.

## Limitations

The GAM is the strongest model in this study, but a few things are worth
keeping in mind: Snapshot only. The model is trained on a single point
in time. No seasonality, no weekday effects, no Montreux Jazz Festival,
no ski season. Recommendations represent an average market price, not a
dynamic one.

Listed price isn't willingness to pay. We use the listed price as a
proxy, but it's what hosts charge, not what guests would actually pay.
Listings that sit unbooked at their current price are still in the
training data, which probably tilts the model towards slightly
optimistic price levels.

±15% is a chosen threshold. It's anchored in the residual SD, which
makes it defensible, but it's still a choice. Using prediction intervals
from the GAM directly, or empirical residual quantiles, would be a more
rigorous alternative.

Missing predictors. Amenity counts (pool, parking, Wi-Fi quality), photo
quality, view, distance to public transport, noise. None of this is in
the dataset. Some of the unexplained variance lives there rather than in
the model's flexibility.

Spatial spline describes, doesn't explain. It fits the price surface
very well but doesn't tell us why certain locations are expensive. For a
pricing tool that's enough; for causal claims (does proximity to the
lake cause higher prices, or just correlate with better properties?) it
isn't.

# SVM

## Theory
A Support Vector Machine separates classes by drawing the widest possible margin between them. We use a radial kernel. This allows us to define the hyperplane more precisely. As a result, with SVM we do not obtain a coefficient that can be interpreted, but rather a model that enables us to predict which group an observation will fall into. 
In the following model, we apply exactly that. We look at the traffic light groups defined in the GAM and try to predict, using the available variables, which group a location will fall into. By splitting the data into a training and test dataset, we enable ourselves to test the model, with the drawback that we lose a portion of the training data. 

With the SVM, we are not testing any of the hypotheses, but rather the second research question: to what extent can we predict whether the prices of a rental property are justified? The GAM generates a fair market price for each rental property. The SVM, on the other hand, examines whether the current rental prices are justified. 

## Model

```{r SVM, Lead Timon}

# Setting a seed ensures reproducibility. 
set.seed(42) 

# Define the relevant columns for the SVM model.
# We include 'price' because the use case requires evaluating if a 
# *given* price for a specific set of features is Overpriced, Fair, or Underpriced.
svm.coloums <- c("traffic_light", "price", "accommodates", "review_scores_rating", 
                 "number_of_reviews", "room_type", "host_is_superhost")

# Creating the 80% Training and 20% Test datasets
d.svm_train <- d.train[, svm.coloums]
d.svm_test  <- d.test[, svm.coloums]


```

```{r tuning model for optimal cost}
# We use 10-fold cross-validation to find the optimal 'cost' and 'gamma' parameters 
# for the radial basis function (RBF) kernel.

# tuned_svm <- tune(
#   svm,
#   traffic_light ~ ., # 'traffic_light' was already converted to a factor in the GAM step
#   data = d.svm_train,
#   kernel = "radial",
#   ranges = list(
#     cost = 2^(-6:10),
#     gamma = 2^(-5:2)
#   )
# )

# Extracting the best parameters found during cross-validation
#svm_model_cost <- as.numeric(tuned_svm$best.parameters["cost"])
#svm_model_gamma <- as.numeric(tuned_svm$best.parameters["gamma"])

###einschub!!!!!!
svm_model_cost <- 256
svm_model_gamma <- 0.625












#displaying Parameters for visual confirmation
#cat("Optimal Cost:", svm_model_cost, "\n")
#cat("Optimal Gamma:", svm_model_gamma, "\n")
#cat("Best Cross-Validation Performance (Error Rate):", round(tuned_svm$best.performance, 4), "\n")

```

```{r TIMON}
# --- Final SVM Model & Evaluation ---

# Training the final Support Vector Machine using the optimized parameters
svm.main <- svm(traffic_light ~ ., 
                data = d.svm_train, 
                kernel = "radial", 
                cost = svm_model_cost, 
                gamma = svm_model_gamma, 
                scale = TRUE) # Scaling to treat all features equally

# Predicting the traffic light categories on test-data
svm.test_pred <- predict(svm.main, d.svm_test)

# Generating the Confusion Matrix to evaluate classification performance
svm.conf_matrix <- table(Predicted = svm.test_pred, Actual = d.svm_test$traffic_light)

# Calculating overall Accuracy
svm_accuracy <- sum(diag(svm.conf_matrix)) / sum(svm.conf_matrix)


kable(svm.conf_matrix, booktabs = TRUE, escape = FALSE, caption = "<span style='color:black; font-weight:bold;'>SVM Confusion Matrix</span>", align = "lccc") %>%
  kable_styling(bootstrap_options = c("bordered")) %>%
  column_spec(2, background = c("deepskyblue", "maroon1", "maroon1")) %>%
  column_spec(3, background = c("maroon1", "deepskyblue", "maroon1")) %>%
  column_spec(4, background = c("maroon1", "maroon1", "deepskyblue")) %>%
  footnote(general = paste0("SVM Test Accuracy: ", round(svm_accuracy * 100, 2), "% \n", "Training set size: ", nrow(d.svm_train), "\n Test set size: " , nrow(d.svm_test)), general_title = "Information:")
             
```


## Interpretion SVM

The tuned SVM with an RBF kernel reaches `r round(svm_accuracy * 100, 2)`% accuracy on the test set. This means the model clearly outperforms chance, which would be 33% for a three-class model. The main issues arise when classifying *Fair Price*. This makes sense, however, since this is the only class that is bounded both above and below and therefore does not exhibit the significant deviations that can be accounted for in the other two categories. For a company like Airbnb, the accuracy is likely too low to make predictions. However, the model shows that estimation within such classes is fundamentally possible. The second Research Question is therefore not easy to answer. We are able to predict wether a price for a location is justified, the accuracy is just not good enough for a company as Airbnb. 

## Limitations

There are a few clear limitations to the model that can be identified right away. The accuracy is too low for a company like Airbnb to provide pricing recommendations to hosts, especially since the category most severely affected is the *Fair Price* category. This leads, in particular, to hosts being incorrectly advised to adjust their prices. Once again, the main cause of the inaccuracy in this model is the location data. We suspect that at the very least, the distance to the nearest city center would have a significant impact on the accuracy of the estimates. 


# ANN

## Theory
An Artificial Neural Network learns flexible, non-linear relationships by chaining together layers of weighted nodes. Unlike the LM or GAM, it does not produce readable coefficients, it learns a mapping from inputs to output and treats the rest as internal weights. That makes it powerful but opaque. We use the ANN here as a flexibility check: if the much more interpretable GAM already captures most of the signal, the ANN should not gain much from being more flexible. If it does, we are leaving structure on the table.

## Link to Hypothesis
We use the ANN to revisit H1 from a different angle. The Naive ANN tests whether host-centered variables alone (*review score*, *review count*, *Superhost status*, *capacity*) can predict price. The Expanded ANN adds *room* and *property type* as controls. Comparing the two shows whether any apparent Superhost effect survives once we compare like, with like a stronger test than just plotting raw medians.

## Model

```{r, Basic ANN Lead Pascal}
# Selecting basic variables. We omit missing values for this simplified approach.
d.ann_naive <- d.airbnb_listings_feat %>%
  select(
    price,                  
    review_scores_rating,   
    number_of_reviews,      
    host_is_superhost,      
    accommodates            
  ) %>%
  na.omit() 

# Neural Networks perform best (and converge faster) when all input features are scaled to a similar range, typically [0, 1].

normalize <- function(x) {
  return ((x - min(x)) / (max(x) - min(x)))
}

# Apply min-max normalization to all columns
d.ann_naive_scaled <- as.data.frame(lapply(d.ann_naive, normalize))

# Train the basic Neural Network
# We use 1 hidden layer with 3 neurons. linear.output = TRUE is used for regression (predicting a continuous price).
set.seed(42) 
nn.naive <- neuralnet(
  price ~ review_scores_rating + number_of_reviews + host_is_superhost + accommodates,
  data = d.ann_naive_scaled,
  hidden = 3,
  linear.output = TRUE,
  stepmax = 1e6         
)

plot(nn.naive)

# Calculate the mean for our features to represent a "statistically average" Airbnb
mean_rating <- mean(d.ann_naive_scaled$review_scores_rating)
mean_reviews <- mean(d.ann_naive_scaled$number_of_reviews)
mean_acc <- mean(d.ann_naive_scaled$accommodates)

# Create two identical fictional hosts: one is a regular host (0), one is a Superhost (1)
test_data_h3_naive <- data.frame(
  review_scores_rating = c(mean_rating, mean_rating),
  number_of_reviews = c(mean_reviews, mean_reviews),
  host_is_superhost = c(0, 1), # The only difference between the two instances
  accommodates = c(mean_acc, mean_acc)
)

# Predict the scaled prices
predictions_h3_naive <- compute(nn.naive, test_data_h3_naive)

# Reverse the min-max scaling to interpret the results in real CHF
min_price <- min(d.ann_naive$price)
max_price <- max(d.ann_naive$price)
real_prices_naive <- predictions_h3_naive$net.result * (max_price - min_price) + min_price

print("Predicted Prices (Naive Model): Regular Host vs. Superhost")
print(real_prices_naive)
```

```{r, Expanded ANN, Lead Pascal}

# Data Preparation & Dimensionality Reduction
d_model_ready <- d.airbnb_listings_feat %>%
  select(
    price,                  
    review_scores_rating,   
    number_of_reviews,      
    host_is_superhost,      
    accommodates,
    property_type,          
    room_type               
  ) %>%
  na.omit() %>%
  # put rare property types into an "Other" category to keep the variable tidy
  mutate(property_type = fct_lump_n(property_type, n = 4, other_level = "Other"))

# One-Hot Encoding: We convert categorical variables into binary dummy variables.
d_dummies <- as.data.frame(model.matrix(~ . - 1, data = d_model_ready))

# Sanitize column names to prevent crashes
colnames(d_dummies) <- make.names(colnames(d_dummies))

d_scaled <- as.data.frame(lapply(d_dummies, normalize))

set.seed(42) 
nn.expanded <- neuralnet(
  price ~ ., 
  data = d_scaled,
  hidden = 3,
  linear.output = TRUE, 
  stepmax = 1e6         
)

plot(nn.expanded)

# To conduct a fair "what-if" analysis (Partial Dependence), we take the mean of ALL variables to represent an absolutely "average" listing.
test_data_h3_exp <- as.data.frame(lapply(d_scaled, mean))

# Duplicate the row to create two identical hypothetical listings
test_data_h3_exp <- rbind(test_data_h3_exp, test_data_h3_exp)

# Isolate the Superhost effect: Row 1 = Regular Host, Row 2 = Superhost
test_data_h3_exp$host_is_superhostFALSE <- c(1, 0) 
test_data_h3_exp$host_is_superhostTRUE <- c(0, 1)

# Predict prices (excluding the 'price' column from the input)
predictions_h3_exp <- compute(nn.expanded, test_data_h3_exp %>% select(-price))

# Reverse scaling using the min/max from the unscaled dataset
min_price_exp <- min(d_model_ready$price)
max_price_exp <- max(d_model_ready$price)

real_prices_expanded <- predictions_h3_exp$net.result * (max_price_exp - min_price_exp) + min_price_exp

print("Predicted Prices (Expanded Model): Regular Host vs. Superhost")
print(real_prices_expanded)
```

## Interpretation ANN

Our initial Neural Network predicted price using only review metrics,
listing capacity, and Superhost status. Surprisingly, it priced
Superhosts lower (132 CHF) than regular hosts (160 CHF). This was a
clear case of omitted variable bias: Superhosts often use competitive
pricing to maintain the high occupancy required for the badge, while
expensive luxury rentals are usually managed by agencies that don't hold
Superhost status.

We corrected this in a second model by adding room and property types,
using One-Hot Encoding to convert these categories into numerical
inputs. Once we controlled for these structural factors to compare
similar listings, the price gap practically vanished (482.64 CHF for
Superhosts vs. 483.29 CHF for regular hosts). This refutes our initial
hypothesis: when comparing apples to apples, Superhost status alone does
not significantly increase a guest's willingness to pay. Superhosts often achieve their high status through
competitive pricing strategies designed to maximize occupancy rather
than nightly margins. Ultimately, a guest's willingness to pay is
dictated almost entirely by physical listing attributes like
accommodates, room_type, and property_type. Therefore, we advise hosts
to anchor their pricing strictly to their property's structural
features, leveraging their Superhost badge purely as a marketing tool to
capture market share and drive booking volume.

## Limitations

A primary limitation of the `nn_expanded model` is the inherent "black
box" nature of Artificial Neural Networks, which drastically reduces the
interpretability of individual price drivers compared to transparent
linear models. Because neural networks require mathematical
normalization, calculating the exact monetary impact of isolated
variables like review_scores_rating is complex and requires continuous
reverse-scaling. Furthermore, while we corrected for some omitted
variable bias by using One-Hot Encoding for property categories, the
model still lacks precise geospatial data, meaning the massive price
impact of specific, highly desirable neighborhoods remains uncaptured.
Testing our hypothesis also relied on creating a mathematically
"average" hypothetical listing, which may not perfectly reflect the
nuanced reality of distinct luxury or budget segments in the Canton of
Vaud. Finally, applying the na.omit() function removed listings with
incomplete profiles, which could introduce bias if those missing data
points systematically represented a specific market segment, such as
poorly managed properties.

# Synthesis, Conclusion & Recommendations

## What the Models Agree On

Three findings hold across very different model architectures:

**The Superhost badge does not raise the nightly price.** Four models with completely different mechanics — the linear regression, the logistic regression (GLM Binomial), the spline-based GAM, and the neural network — all return the same answer: after controlling for listing size, room type and other structural features, Superhost status has no measurable positive effect on price. Some models even show a small *negative* effect. This is the most robust single finding in this study.

**Structural features are the real price drivers.** Capacity (`accommodates`), room type, and the number of bathrooms account for most of the explainable variation in price across all models that include them. The GAM shows that an additional bathroom adds roughly 25% to the price; a private room sells for about 30% less than an entire home, all else equal. Reviews and host attributes matter, but their effect is small compared to the physical features a host can describe in a listing.

**Location matters, and matters non-linearly.** Lake Geneva and Lausanne sit at the top of the price surface; rural and alpine areas drop off sharply. The GAM's 2D spatial spline captures this with an effective degree of freedom around 25 — far too curved to approximate with linear coordinates or postcode dummies. Any tool that ignores location will systematically misprice listings outside the canton's average baseline.

## What the Different Lenses Add

The price models and the occupancy model tell complementary halves of the Superhost story:

| Lens | Finding |
|---|---|
| LM, GLM Binomial, GAM, ANN | Superhosts charge **no more** (and slightly less) than regular hosts |
| GLM Poisson | Superhosts get **2.6× more bookings** than regular hosts |

Read together, these results sketch out the Superhost strategy: price competitively, earn through volume. A host who chases the badge by raising prices is doing it backwards. The badge comes from booking volume and consistent review quality — both of which are easier to achieve at a competitive price point.

A second contrast worth highlighting is the Naive ANN versus the Expanded ANN. The naive version, trained only on host variables, predicts a *lower* price for Superhosts (132 vs 160 CHF). Once room and property type are added, the gap collapses to less than a franc. This isn't really a disagreement between models — it's a textbook case of omitted variable bias, and a useful warning against drawing conclusions from raw price comparisons.

## Recommendations for Hosts

For hosts setting or revising their nightly price:

- **Anchor on structure, not status.** What a guest actually pays for is the listing itself: room type, capacity, bathrooms, location. The Superhost badge helps with visibility and booking volume, not with the per-night rate.
- **Use the traffic-light tool as a sanity check, not a target.** The GAM-based recommendation gives a fair-market range based on what comparable listings achieve. A listing flagged "Underpriced" is leaving money on the table; one flagged "Overpriced" risks sitting empty.
- **If you're building toward Superhost status, compete on price.** The 2.6× occupancy boost comes from satisfied guests and a full calendar. Premium pricing makes both harder. Start with competitive pricing, accumulate reviews, then raise prices selectively once the badge is secure.
- **Review scores matter less than you think for price.** Across the pricing models the reward for higher ratings is modest. Reviews matter, but their role is to drive *bookings* — and through them, the badge — rather than to justify a higher nightly rate.

## Recommendations for Airbnb

For the platform itself, the same data points in three directions:

- **A pricing assistant should use the GAM, not a linear formula.** Linear or postcode-based tools will systematically underprice lakeshore listings and overprice rural ones. The 2D spatial spline is the most defensible way to capture local dynamics with the data Airbnb already collects.
- **Coach the "Underpriced" segment.** Roughly 41% of listings in the dataset sit below the GAM's fair-price band. These hosts are the easiest group to help: a recommendation alone could shift their pricing toward market value, lifting both host revenue and platform commission.
- **Add operationalised location features.** Raw coordinates work well for a smoother like the GAM but limit simpler tools. Distance to the nearest lake, city centre, ski lift, and public transport stop would let classifiers like the SVM catch up to the GAM's accuracy — and these features are trivial to compute centrally and serve to every model on the platform.

# Overall Limitations

Five limitations apply to this entire study, not just to individual models:

**Snapshot, no time component.** All models are trained on a single point in time. Seasonal pricing, weekday effects, and event-driven demand spikes (Montreux Jazz Festival, ski season, weekends) are nowhere in the data. The recommendations represent an *average* market price, not a dynamic one. A real production pricing tool would need a time dimension we don't have.

**Listed price as a proxy for willingness to pay.** We treat what hosts charge as a proxy for what guests would pay. This holds on average across a large sample, but it breaks down for listings that sit unbooked at their current price — those listings still shape the model and probably tilt it slightly toward optimistic price levels. A true willingness-to-pay study would need booking data, completed transactions, and ideally A/B tests of competing prices.

**Missing predictors.** Photo quality, amenity counts, listing description sentiment, view from the property, noise levels, distance to public transport — none of this is in the dataset. Some of the unexplained variance (around a third for the GAM) lives in these omitted variables rather than in any model's limitation.

**Reverse causality.** Several relationships in our data probably flow in both directions. More bookings produce more reviews, which make Superhost status easier to maintain, which probably increases visibility, which leads to more bookings. None of our models can untangle which effect causes which. For descriptive purposes that's fine, but the recommendations should be read as "what works on average," not "what works because of X."

**Vaud-specific.** The dataset covers one Swiss canton at one point in time. The price levels, location dynamics (lakeshore premium, alpine resorts), and host behaviours are specific to this market. The methodology transfers; the exact coefficients do not. Applying the same models to Geneva or Bern would require refitting on local data.

# Use of GenAI

Throughout this machine learning project, we leveraged various
Generative AI tools specifically ChatGPT, Gemini, and Claude to support
our analytical workflow. A primary use case was troubleshooting complex
issues that extended beyond our existing knowledge base. When
encountering technical bottlenecks, we used these models to break down
and explain the underlying concepts, enabling us to systematically find
the right solutions. Furthermore, we utilized GenAI as a critical review
mechanism for our modeling approach. We consulted the AI to verify
whether we had made the appropriate theoretical assumptions, such as
selecting the correct data distributions for our models. It also served
as a valuable sounding board to double-check our final analytical
interpretations, ensuring our conclusions were statistically sound.
Finally, these tools provided essential linguistic support. We actively
used GenAI to translate our documentation and findings into professional
English, ensuring a high-quality final report.

# Appendix
## Lessons learned
### Pascal
Working on this project was a huge reality check in the best way possible. In our classes, we are used to getting "perfect" datasets that are practically made for the specific algorithms we are learning. But working with our actual `d.airbnb_listings_feat` dataset showed me how much effort actually goes into data cleaning and prepping before you can even start modeling. I personally focused on building the GLM-Poisson and ANN models, and trying to get them to perform well on real, messy data required a lot more critical thinking and trial-and-error than our standard assignments.

Since we split up the workload, I didn't code every single model myself. Instead, my role for the LM, GLM-Binomial, GAM, and SVM models was to check and review the work of my team members. This turned out to be a super valuable learning experience on its own. Reading through their code allowed me to see exactly how other people translate what we learned in class into practice. Seeing their different approaches to problem-solving and troubleshooting bugs really helped expand my own coding mindset.

Finally, figuring out how to collaborate effectively on code was a massive win for us. We used GitHub with RStudio for the first time, and it honestly made a world of difference. We finally found a setup that actually works for coding together as a group, which saved us a lot of headaches and is definitely a workflow I’ll be taking into future projects.