---
title: "Main Group Work_ML"
author: "Pascal Ingold, Timon Reichmuth und Davide Gualdani"
date: "2026-03-22"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, comment = NA)

```

Next Steps: 
bis Montag 09.00 Timon & Davide hypo schärfen.
- Forschungsfragen ausarbeiten (präzision auf Host richten)

Pascal: 
- Feature Engineering ausbauen

Timon: 
- 

Davide:
- 

Modelle zu verteilen: 
– a Linear Model --> lm(price ~ rewview_rating + Superhost T/F) --> für unterscheidung und kontrolle. 
– a Generalised Linear Model with family set to Poisson --> glm(Price ~ reviews)
– a Generalised Linear Model with family set to Binomial --> glm(Price ~ Superhost T/F)
– a Generalised Additive Model -->  gam(price ~  s(accommodates) +   --> allenfalls aufbauend auf dem lm()
                                    s(review_scores_rating) + 
                                    s(availability_365) +   
                                    s(number_of_reviews) +   
                                    s(longitude, latitude) +   
                                    s(host_listings_count) +   
                                    room_type +   
                                    host_is_superhost +   
                                    bathrooms + 
                                    bedrooms,   
                                    family = gaussian(link="log"),   
                                    data = df_clean)

– a Neural Network --> 
– a Support Vector Machine --> Overpriced Underpriced Right Price



Infos zu bedenken: 
ABGABEDATUM: irwo im Juni 26

- Outlier Handling definieren.
--> falsche Datenpunkte sollten gelöscht werden. Extremwerte im Zweifel drin lassen. Ansonsten kann das als p-hacking verstanden werden. 


#Research Questions

What factors, which can be modified by Hosts on AirBnb, affect guest willingness to pay more for a certain room?

What are factors that influence the guest willingness to pay (WTP)

What are factors on which the guest can adjust their pricing model?

##Concepts: Willingness to pay (WTP)

WTP is the hypothetical value, how much a customer is willing to pay for a location or a room for a certain standard. This can rise, if for example the host provides the ability to wash clothes in the apartment. So for the same location, once with and once without a washing machine, the customers WTP might increase with the washer. In a more abstract example, if the host is a super host, the customer might trust the host more and his WTP is higher. 

##Hypothesis
H1: If the review-score of a host is better, the WTP of a guest increases. 
H2: If a host has more reviews, the WTP of a guest increases. 
H3: If a host is e superhost, the WTP of a guest increases. 
H4: 

WTP is measured in the Price displayed. We assume, that the mean of prices is similar to the WTP and residuals even out. Therefore we can work with the effective prices displayed on AirBnb. 



# Business Case
We propose a data-driven pricing decision-support system for Airbnb hosts in the
Canton of Vaud, providing market-aligned price recommendations based on listing
characteristics. The tool targets both new and existing hosts, using inputs such
as location, room type, guest capacity, availability, ratings, and superhost status.
Machine learning models, including LM, GLM, GAM, SVM, and Neural
Networks, will be applied to identify key price drivers and generate accurate
predictions. The objective is to balance revenue and occupancy, helping hosts
understand their market value.
A key feature is interpretability, enabling users to see why a certain price is
recommended. From a business perspective, the system aims to increase booking
rates, reduce price dispersion, and improve host satisfaction.
The model is tailored to the Canton of Vaud to capture regional economic effects.

First step: Loading Packages

```{r}
library(tidyverse)
library(janitor)
library(mgcv)
library(e1071)
library(nnet)
library(caret)
library(pROC)
library(Metrics)
library(performance)
library(broom)
library(stringr)
library(ggplot2)
library(patchwork)
library(knitr)
```



Second step: Reading all available Datasets
```{r}
path_listings <- "./Data_Vaud_all/listings-2.csv"
d.raw_data_listings <- read.csv(path_listings, header = TRUE)
```

Cleaning the Airbnb-Listings Data. 

```{r}
d.airbnb_listings <- d.raw_data_listings %>%
  clean_names()

d.listings_clean <- d.airbnb_listings %>%
  mutate(
    price = str_remove_all(price, "\\$|,"),
    price = as.numeric(price),
    host_is_superhost = ifelse(host_is_superhost %in% c("t", "TRUE", "1"), 1, 0),
    room_type = as.factor(room_type),
    property_type = as.factor(property_type)
  ) %>%
  filter(!is.na(price), price > 0)

d.listings_clean <- d.listings_clean %>%
  select(-id, -host_id, -host_listings_count, -host_total_listings_count, -listing_url, -scrape_id, -last_scraped,
         -source, -name, -description, -maximum_nights, -first_review, -last_review, -calculated_host_listings_count, -host_since, -availability_365, -estimated_revenue_l365d, 
         -neighborhood_overview, -picture_url, -host_url, -host_name, -host_location,
         -host_about, -host_thumbnail_url, -host_picture_url, -host_neighbourhood,
         -neighbourhood, -neighbourhood_cleansed, -bathrooms_text, -amenities, 
         -minimum_minimum_nights, -maximum_minimum_nights, -minimum_maximum_nights,
         -maximum_maximum_nights, -minimum_nights_avg_ntm, -maximum_nights_avg_ntm,
         -calendar_updated, -has_availability, -availability_30, -availability_60,
         -availability_90, -calendar_last_scraped, -number_of_reviews_ltm, -number_of_reviews_l30d,
         -availability_eoy, -number_of_reviews_ly, -review_scores_accuracy, -review_scores_cleanliness,
         -review_scores_checkin, -review_scores_communication, -review_scores_location,
         -license, -instant_bookable, -calculated_host_listings_count_entire_homes, -calculated_host_listings_count_private_rooms,
         -calculated_host_listings_count_shared_rooms, -reviews_per_month, -bedrooms, -review_scores_value)
```


First Feature Engineering for the airbnb_listings data

```{r}
d.airbnb_listings_feat <- d.listings_clean %>%
  mutate(
    log_price = log(price),
    price_per_guest = price / pmax(accommodates, 1),
    high_demand = ifelse(estimated_occupancy_l365d > 200, 1, 0)
  )
```

```{r}
str(d.airbnb_listings_feat)
```


Exploratory Data Analysis

```{r}
hist(d.airbnb_listings_feat$price)
hist(d.airbnb_listings_feat$log_price)
ggplot(data= d.airbnb_listings_feat, aes(x = room_type, y = price)) +
  geom_boxplot() +
  scale_y_log10() +
  labs(
    title = "Preisverteilung nach Room Type (log-Skala)",
    x = "Room Type",
    y = "Log_Preis"
  ) +
  theme_minimal()

ggplot(data = d.airbnb_listings_feat, aes(x = accommodates, y = log_price)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "loess", se = FALSE) +
  labs(
    title = "Preis vs. Kapazität (log-Skala)",
    x = "Anzahl Gäste",
    y = "Log_Preis"
  ) +
  theme_minimal()
  
ggplot(data = d.airbnb_listings_feat, aes(x = log_price, y = number_of_reviews)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "loess", se = FALSE) +
  labs(
    title = "Anzahl Reviews vs. Preis",
    x = "Log_Preis",
    y = "Anzahl Reviews"
  ) +
  theme_minimal()

ggplot(data = d.airbnb_listings_feat, aes(x = longitude, y = latitude)) +
  geom_point(aes(color = log_price), alpha = 0.6) +
  scale_color_viridis_c() +
  labs(
    title = "Airbnb Listings (Farbe = Log-Preis)",
    x = "Longitude",
    y = "Latitude",
    color = "Log_Preis"
  ) +
  theme_minimal()

ggplot(data = d.airbnb_listings_feat, aes(x = longitude, y = latitude)) +
  geom_point(aes(color = log_price, size = number_of_reviews), alpha = 0.5) +
  scale_color_viridis_c() +
  labs(
    title = "Listings: Preis und Nachfrage",
    x = "Longitude",
    y = "Latitude",
    color = "Log_Preis",
    size = "Reviews"
  ) +
  theme_minimal()
```

*Important to mention*
During our exploratory data analysis, we identified one 'Private room' listing with an unrealistically high price of 2'147'483'647 CHF. We excluded this single observation from our dataset as it is an extreme outlier that does not reflect the true market conditions and would negatively skew our predictive models.
```{r}
# This keeps everything EXCEPT the row that is a Private room AND costs more than 5000
d.airbnb_listings_feat <- d.airbnb_listings_feat %>%
  filter(!(room_type == "Private room" & price > 100000))
```


Davide's EDA
```{r}
# --- 1. Distribution of the Target Variable (Price) ---
# Upgrading base R hist() to ggplot2 for a more professional look.
# Showing both raw and log side-by-side helps visually justify the transformation to the client.
p_raw <- ggplot(d.airbnb_listings_feat, aes(x = price)) +
  geom_histogram(bins = 50, fill = "steelblue", color = "white") +
  labs(title = "Raw Price Distribution", x = "Price", y = "Count") +
  theme_minimal()

p_log <- ggplot(d.airbnb_listings_feat, aes(x = log_price)) +
  geom_histogram(bins = 50, fill = "darkgreen", color = "white") +
  labs(title = "Log-transformed Price Distribution", x = "Log(Price)", y = "Count") +
  theme_minimal()
p_raw + p_log # Displays them side-by-side

# --- 2. Categorical Predictor: Room Type vs. Price ---
# Your original code was good, just cleaned up the aesthetics.
ggplot(data = d.airbnb_listings_feat, aes(x = room_type, y = log_price, fill = room_type)) +
  geom_boxplot(alpha = 0.7) +
  labs(
    title = "Price Distribution by Room Type",
    x = "Room Type",
    y = "Log(Price)"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

# --- 3. Continuous Predictor: Capacity vs. Price ---
ggplot(data = d.airbnb_listings_feat, aes(x = accommodates, y = log_price)) +
  geom_point(alpha = 0.3, color = "darkgray") +
  geom_smooth(method = "loess", color = "blue", se = FALSE) +
  labs(
    title = "Price vs. Capacity",
    subtitle = "Smoother applied to identify potential non-linear trends",
    x = "Number of Guests (Accommodates)",
    y = "Log(Price)"
  ) +
  theme_minimal()

# --- 4. Continuous Predictor: Reviews vs. Price ---
# Correction: The response variable (log_price) should generally be on the y-axis 
# when plotting against a predictor[cite: 255].
ggplot(data = d.airbnb_listings_feat, aes(x = number_of_reviews, y = log_price)) +
  geom_point(alpha = 0.3, color = "darkgray") +
  geom_smooth(method = "loess", color = "red", se = FALSE) +
  labs(
    title = "Does the Number of Reviews impact Price?",
    x = "Total Number of Reviews",
    y = "Log(Price)"
  ) +
  theme_minimal()

# --- 5. Binary Predictor: Superhost Status ---
# Adding this because it's a key part of your business case and fulfills the binary variable requirement.
ggplot(data = d.airbnb_listings_feat, aes(x = as.factor(host_is_superhost), y = log_price, fill = as.factor(host_is_superhost))) +
  geom_boxplot(alpha = 0.7) +
  scale_x_discrete(labels = c("0" = "Regular Host", "1" = "Superhost")) +
  labs(
    title = "Price Difference: Superhosts vs. Regular Hosts",
    x = "Host Status",
    y = "Log(Price)"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

# --- 6. Spatial Data: Mapping the Listings ---
# Cleaned up the map to focus purely on location and price.
ggplot(data = d.airbnb_listings_feat, aes(x = longitude, y = latitude)) +
  geom_point(aes(color = log_price), alpha = 0.6, size = 1.5) +
  scale_color_viridis_c(option = "plasma") +
  labs(
    title = "Airbnb Listings in Vaud by Price",
    x = "Longitude",
    y = "Latitude",
    color = "Log(Price)"
  ) +
  theme_minimal()
```
Davide's overall Data Analysis
# 1. Overall Market Overview (Prices & Capacity)
This is the absolute minimum, maximum, median, and mean for the most important numeric variables.
```{r}
# Basic summary statistics for key numerical variables
summary_stats <- d.airbnb_listings_feat %>%
  select(price, log_price, accommodates, number_of_reviews, availability_365) %>%
  summary()

print(summary_stats)
```
Client Interpretation: "The average Airbnb in Vaud costs X CHF per night, but the median is Y CHF, indicating that a few extremely expensive luxury listings might be skewing the average upward. Prices range from a minimum of A CHF to a maximum of B CHF."

# 2. Analysis by Room Type
Since we want to know how many different types of rooms there are, we group the data and calculate the counts, percentages, and price ranges for each type.
```{r}
# Detailed breakdown by Room Type
room_type_analysis <- d.airbnb_listings_feat %>%
  group_by(room_type) %>%
  summarise(
    Total_Listings = n(),
    Percentage_of_Market = round((n() / nrow(d.airbnb_listings_feat)) * 100, 1),
    Min_Price = min(price, na.rm = TRUE),
    Median_Price = median(price, na.rm = TRUE),
    Average_Price = round(mean(price, na.rm = TRUE), 1),
    Max_Price = max(price, na.rm = TRUE)
  ) %>%
  arrange(desc(Total_Listings)) # Sorts by the most common room type

print(room_type_analysis)
```
Client Interpretation: "Entire homes/apartments dominate the Vaud market, making up 75% of all listings. As expected, the Hotel room command the highest median price (264.5 CHF), compared to shared rooms (47 CHF)."

# 3. Superhost vs. Regular Host Breakdown
This business case specifically mentions superhost status. So we took a look to see how many there are and if they actually charge more or even less.
```{r}
# Superhost statistics
superhost_analysis <- d.airbnb_listings_feat %>%
  group_by(host_is_superhost) %>%
  summarise(
    Total_Hosts = n(),
    Percentage = round((n() / nrow(d.airbnb_listings_feat)) * 100, 1),
    Median_Price = median(price, na.rm = TRUE),
    Avg_Reviews = round(mean(number_of_reviews, na.rm = TRUE), 1)
  ) %>%
  mutate(host_is_superhost = ifelse(host_is_superhost == 1, "Yes", "No"))

print(superhost_analysis)
```

# 1. Host Experience vs. Price (host_since_year)
Now let's test the hypothesis: can "older" (more experienced) hosts charge higher prices than newcomers?
```{r}
# First, ensure the year is treated as a numeric variable
d.airbnb_listings_feat <- d.airbnb_listings_feat %>%
  mutate(host_since_year = as.numeric(host_since_year))

# Plot: Price trend over the years of host membership
ggplot(data = d.airbnb_listings_feat %>% filter(!is.na(host_since_year)), 
       aes(x = host_since_year, y = log_price)) +
  geom_point(alpha = 0.2, color = "darkcyan") +
  # Adding a smoother is mandatory for exploratory plots according to the hints!
  geom_smooth(method = "gam", formula = y ~ s(x, bs = "cs"), color = "orange", se = TRUE) +
  labs(
    title = "Can established hosts command higher prices?",
    x = "Host Since (Year)",
    y = "Log(Price)"
  ) +
  theme_minimal()
```
Client Interpretation: "We observe that hosts who joined the platform earlier (e.g., around 2015) do (or do not) command higher average prices compared to recent joiners. For our pricing tool, this means..."


# 2. Correlation Matrix (Crucial before Machine Learning)
Before we train our LM, GAM, or SVM, we need to know how our numerical variables relate to each other. This helps us identify multicollinearity (when two predictors provide the same information) and strong price drivers.

```{r}
# Select relevant numerical variables (drop NAs just for the correlation calculation)
cor_data <- d.airbnb_listings_feat %>%
  select(log_price, accommodates, number_of_reviews, availability_365) %>%
  drop_na()

# Calculate the correlation matrix
cor_matrix <- cor(cor_data)

# Render a nice table for your RMarkdown document
kable(round(cor_matrix, 2), caption = "Correlation Matrix of Key Numerical Drivers")
```
Client Interpretation: "As expected, accommodates shows the strongest positive correlation with price. It is also interesting to note the relationship between availability and number_of_reviews..."

# 3. Detailed property_type Analysis
We already analyzed room_type, but Airbnb often has dozens of property_types (e.g., "Entire chalet", "Private room in castle", etc.). Summarizing the Top 10 property types in the Canton of Vaud provides maybe a great market insight.
```{r}
# Top 10 Property Types by frequency and their median price
top_properties <- d.airbnb_listings_feat %>%
  group_by(property_type) %>%
  summarise(
    Total_Listings = n(),
    Median_Price = median(price, na.rm = TRUE)
  ) %>%
  arrange(desc(Total_Listings)) %>%
  slice_head(n = 10) # Keep only the top 10

kable(top_properties, caption = "Top 10 Property Types in the Canton of Vaud")
```

Client Interpretation: "While standard apartments like "entire rental unit" are the most common, specialty properties like chalets, homes and villas command a significant premium, which our model will need to capture."

# 4. Missing Values (NA) Check
This step is less for the client and essential for your ML models. Most modeling functions (like lm() or svm()) will throw errors or silently drop entire rows if NAs are present. You need to know exactly where your data is missing!

```{r}
# Count the number of NAs for each key column in your dataset
na_summary <- d.airbnb_listings_feat %>%
  select(price, log_price, accommodates, number_of_reviews, host_is_superhost, availability_365) %>%
  summarise(across(everything(), ~ sum(is.na(.)))) %>%
  t() %>% # Transpose for better readability
  as.data.frame() %>%
  rename(Missing_Values = V1)

kable(na_summary, caption = "Count of Missing Values per Variable")
```





Bi gad Moti gsi
Davide: Linear Model
Before we jump straight into the lm() function, we need to do one crucial step for your business case: Train/Test Splitting. Since your goal is to build a predictive pricing tool, you need to train your model on one chunk of the data (e.g., 80%) and test its accuracy on data it has never seen before (the remaining 20%). This proves to the client that your tool actually works in the real world.Additionally, the evaluation guidelines explicitly warn against fitting a "very simple linear model" if you are going to compare it to complex models like Neural Networks or SVMs. To make it a "fair way" of comparing, you should include at least one interaction term in your LM.

```{r}
# --- 1. Split Data into Training and Testing Sets ---
# Using the caret package (which you already loaded) to create an 80/20 split
set.seed(42) # Set a seed so you get the exact same split every time you compile!
train_index <- createDataPartition(d.airbnb_listings_feat$log_price, p = 0.8, list = FALSE)

train_data <- d.airbnb_listings_feat[train_index, ]
test_data <- d.airbnb_listings_feat[-train_index, ]

# --- 2. Fit the Linear Model ---
# Note: We MUST use log_price because "amounts" must virtually always be log-transformed[cite: 192, 193].
# We add an interaction term (*) between accommodates and room_type for a fairer comparison with complex models[cite: 182].
lm_model <- lm(log_price ~ accommodates * room_type + 
                 host_is_superhost + 
                 number_of_reviews + 
                 availability_365, 
               data = train_data)

# --- 3. View the Results ---
summary(lm_model)
```

# Interpreting the Coefficients
Because your target variable is log_price, interpreting the coefficients requires a slight mental shift. A coefficient in a log-linear model represents an *approximate percentage change in the original price*.

## The Baseline (The Intercept)
*What the math says:* The intercept is 4.371.
*Client Translation:* Our baseline reference listing—an 'Entire home/apt' (the category missing from the list, meaning it's the default)—has an estimated base price of roughly 79 CHF per night (exp(4.371)), assuming all other factors are zero.

## Capacity (Accommodates)
*What the math says:* The coefficient is 0.1553 (Highly significant).
*Client Translation:* Capacity is a massive price driver. For a standard entire apartment, every additional guest the property can accommodate increases the nightly price by 16.8% ((exp(0.1553) - 1) * 100).

## Room Types & The Interaction Term
This is where your interaction term (accommodates * room_type) pays off, proving why you shouldn't use a simple, plain model!

*Private Rooms:* They are inherently cheaper. The base price drops by roughly 30.8% (exp(-0.3688) - 1) compared to an entire home. The interaction term for private rooms is not statistically significant (p = 0.52), meaning an extra guest in a private room adds about the same 16.8% premium as it does in an entire home.

*Shared Rooms:* This is the most interesting part. The base price drops, but look at the interaction term: -0.1585 (Highly significant). If you add the base guest coefficient (0.1553) to this shared room interaction (-0.1585), you get essentially zero.

*Client Translation:* "While adding guests allows hosts of 'Entire homes' and 'Private rooms' to charge about 17% more per guest, this rule does not apply to 'Shared rooms'. For shared rooms, adding more beds does not allow the host to increase the price at all.

## Surprising Business Insights (Other Drivers)
*Superhost Status (-0.0648):* "Counter-intuitively, Superhosts actually charge about 6.3% less than regular hosts. From a business perspective, this suggests that hosts might achieve Superhost status because they offer highly competitive, slightly lower prices to ensure high booking rates and good reviews."

*Number of Reviews (-0.0014):* "Listings with more reviews tend to be slightly cheaper (each review drops the price by about 0.14%). This likely points to 'budget' listings having a higher turnover of guests, and therefore accumulating more reviews."

*Availability (0.00029):* "Listings with higher availability are slightly more expensive. This makes sense: highly expensive places are booked less often, leaving more days available on their calendar."

# Evaluating Predictive Performance
To see how "accurate" our model is, we will calculate the Root Mean Squared Error (RMSE) on the test data. The guidelines state that if we cannot find another measure of fit, we can use RMSE.
```{r}
# --- 4. Make Predictions on the Test Data ---
# Predict the log_price for the unseen test data
lm_predictions <- predict(lm_model, newdata = test_data)

# FIX: Force R to use the Metrics package specifically
lm_rmse <- Metrics::rmse(test_data$log_price, lm_predictions)

print(paste("Linear Model RMSE on Test Data:", round(lm_rmse, 4)))

```

*Client Translation:* "Our baseline Linear Model achieved a Root Mean Squared Error (RMSE) of 0.4624 on unseen test data. Because we are predicting the logarithm of the price, this means our model's predictions are typically off by a factor of roughly exp(0.4624), which equals about 1.58. In practical terms, our base model's price estimates are generally within a ~58% margin of error. While this is a good starting point, it highlights the need for our more advanced machine learning models to capture the complex, non-linear pricing dynamics in the Vaud market."

