---
title: "Main Group Work_ML"
author: "Pascal Ingold, Timon Reichmuth und Davide Gualdani"
date: "2026-03-22"
output:
  html_document:
    df_print: paged
    code_folding: hide
    toc: true
    tod-depth: 3
    number_sections: true
    theme: cosmo
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, comment = NA, warning = FALSE)

```

Tasks ab 29.04
- Frühlingsputz (PASCAL)
- Limitationen und Business Case im Kapitel  ergänzen. jeweils für die eigenen
Modelle. (TIMON) (DAVIDE) (Pascal --> done)
- GLM (Binominal) gemäss Feedback Timon noch anpassen und mit Text ausschmücken sowie Visualisieren (DAVIDE) - DONE
- GAM text kürzen (DAVIDE) - DONE
- EDA verdichten und Platz sparen (Plots) - TBD welche wir drin lassen.
- Colors vereinheitlichen (TIMON)
- README File (PASCAL) --> done
- interpretation Crossvalidation schreiben (TIMON)
- Coding Conventions kontrollieren (ALLE) - Davide GAM / GLM done, 
- 





ToDos backlog:
- AI-Texte: Praktisch alle Modelle haben ganz klar AI-generierten Code. Auch Kommentare und Interpretationen sind noch sehr offensichtlich. Unbedingt überarbeiten!
- beautifying HTML-Doc
- Data Quality kontrollieren & übersetzen.  
- AI Chapter
- Lesson Learned Kapitel




Infos zu bedenken: ABGABEDATUM: 12 Juni


Coding Conventions. 
• data sets are named with the prefix “d.” (e.g.,
d.cats). 
• linear model objects are named with the prefix “lm.” (e.g.,
lm.cats) 
• plots are named with the prefix "p."
• within the text R-objects are written using code formatting
(e.g., lm.cats) 14 
• variable names are written using italic (e.g.,
Body.weight) 
• these conventions are arbitrary, but hopefully useful to
the reader

# Introduction

## Business Case

We propose a data-driven pricing decision-support system for Airbnb
hosts in the Canton of Vaud, providing market-aligned price
recommendations based on listing characteristics. The tool targets both
new and existing hosts, using inputs such as location, room type, guest
capacity, availability, ratings, and superhost status. Machine learning
models, including LM, GLM, GAM, SVM, and Neural Networks, will be
applied to identify key price drivers and generate accurate predictions.
The objective is to balance revenue and occupancy, helping hosts
understand their market value. A key feature is interpretability,
enabling users to see why a certain price is recommended. From a
business perspective, the system aims to increase booking rates, reduce
price dispersion, and improve host satisfaction. The model is tailored
to the Canton of Vaud to capture regional economic effects.

## Research Questions

Several questions can be derived from the business case. These questions are intended to help us examine pricing dynamics and facilitate the development of a pricing strategy for Airbnb. 
RQ1 - First, **we need to determine which factors influence guests’ willingness to pay**. Furthermore, the nature of this influence is, of course, central to our analysis. 
RQ2 - Once these factors are identified, a second question immediately arises: **to what extent can these factors be used to predict whether the prices of a rental property are justified?**

Willingness to pay is actually the central variable here. However, it is not easy to measure, which is why a different approach must be taken. The approach we have chosen stems from the idea that the market generally functions well and that landlords intuitively already align with willingness to pay. As a result, the existing price for a rental property can also be assumed to be the price that tenants are willing to pay. Any deviations are offset in our subsequent calculations by the large number of observations. Thus, on average, the actual price will correspond to the price that tenants are willing to pay. 


### Hypothesis

H1: When host-centered indicators improve, tenants' willingness to pay increases. 
H2: When location-centered indicators improve, tenants’ willingness to pay increases. 

Host-centered indicators are those that relate to the host’s performance, such as ratings or Superhost status. In contrast, location-centered indicators are those that relate to factors such as the size of the rental property. 


**Note**

**Es sollte Möglich sein, die jeweiligen Modell in einer oder beiden Hypothesen zu verordnen. Sprich: ich untersuche mit meinem Modell H1 und finde dass mehr reviews den Preis steigen lassen blablabla. Zudem kann es in gewissen Modellen sinn machen, beide Hypothesen anzuschauen und zu beantworten. es kann aber auch sein, dass man zb H1 anschaut und mit Faktoren wie "accommodates" kontrolliert. dann habe ich zwar location-sided-Indicators, jedoch habe ich keinen Fokus auf die H2.**
**Eine dritte Option ist, dass ich nicht eine der beiden Hypothesen untersuche, sondern die zweite Forshcungsfrage und versuche vorhersagen zu treffen. dies macht beispielsweise beim SVM sinn. da müsst ihr selbst enscheiden, welchen Fokus ihr setzen wollt**


# Data Preparation

## Editing of Dataset

First step: Loading Packages

```{r Libraries, results = "hide", echo=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(mgcv)
library(e1071)
library(nnet)
library(pROC)
library(Metrics)
library(performance)
library(broom)
library(stringr)
library(ggplot2)
library(patchwork)
library(knitr)
library(performance)
library(naniar)
library(visdat)
library(neuralnet)
library(leaflet)
library(leaflet.extras)
library(patchwork)


theme_set(theme_minimal())
```

Second step: Reading all available Datasets

```{r reading Dataset, echo=FALSE}
path_listings <- "./Data_Vaud_all/listings-2.csv"
d.raw_data_listings <- read.csv(path_listings, header = TRUE)
```

Cleaning the Airbnb-Listings Data. We keep only the variables needed for
our project. We reduce the amount of variables to exact 20, according to
the requirements by our lecturers.

```{r cleaning Dataset, echo=FALSE}
d.airbnb_listings <- d.raw_data_listings %>%
  clean_names()

d.listings_clean <- d.airbnb_listings %>%
  mutate(
    price = str_remove_all(price, "\\$|,"),
    price = as.numeric(price),
    host_is_superhost = ifelse(host_is_superhost %in% c("t", "TRUE", "1"), 1, 0),
    room_type = as.factor(room_type),
    property_type = as.factor(property_type),
    host_response_rate = as.numeric(sub("%", "", host_response_rate)),
    host_acceptance_rate = as.numeric(sub("%", "", host_acceptance_rate)),
    host_is_superhost = ifelse(host_is_superhost==1, TRUE, FALSE)
  ) %>%
  filter(!is.na(price), price > 0)




d.listings_clean <- d.listings_clean %>%
  select(-id, -host_id, -host_listings_count, -host_total_listings_count, -listing_url, -scrape_id, -last_scraped,
         -source, -name, -description, -maximum_nights, -first_review, -last_review, -calculated_host_listings_count, -host_since, -availability_365, -estimated_revenue_l365d, 
         -neighborhood_overview, -picture_url, -host_url, -host_name, -host_location,
         -host_about, -host_thumbnail_url, -host_picture_url, -host_neighbourhood,
         -neighbourhood, -neighbourhood_cleansed, -bathrooms_text, -amenities, 
         -minimum_minimum_nights, -maximum_minimum_nights, -minimum_maximum_nights,
         -maximum_maximum_nights, -minimum_nights_avg_ntm, -maximum_nights_avg_ntm,
         -calendar_updated, -has_availability, -availability_30, -availability_60,
         -availability_90, -calendar_last_scraped, -number_of_reviews_ltm, -number_of_reviews_l30d,
         -availability_eoy, -number_of_reviews_ly, -review_scores_accuracy, -review_scores_cleanliness,
         -review_scores_checkin, -review_scores_communication, -review_scores_location,
         -license, -instant_bookable, -calculated_host_listings_count_entire_homes, -calculated_host_listings_count_private_rooms,
         -calculated_host_listings_count_shared_rooms, -reviews_per_month, -bedrooms, -review_scores_value)
```

According to the requirements for this paper, the dataset is limited to a maximum of 20 variables. To meet this specification, we have removed columns from the dataset that will not be used in our analysis. Consequently, `d.listings_clean` is our final, cleaned dataset, which we will use to run all subsequent models.

```{r final Call cleaned Dataset, echo=TRUE}
str(d.listings_clean)
rm(d.airbnb_listings, d.raw_data_listings)

```

Subsequently, we calculated several additional features to include them as permanent columns in the dataset, ensuring they do not have to be recomputed for each analysis. These include: the log-transformed price—as price data is typically heavily right-skewed—the price per guest (total price divided by the number of guests accommodated), a boolean indicator for high demand, and a binary variable identifying whether the price exceeds the median.

```{r initial FE (Pascal), echo=FALSE}
d.airbnb_listings_feat <- d.listings_clean %>%
  mutate(
    log_price = log(price),
    price_per_guest = price / pmax(accommodates, 1),
    high_demand = ifelse(estimated_occupancy_l365d > 200, 1, 0),
    price_high = as.integer(price > median(price, na.rm =TRUE))
  )
```




# Understanding the Data

Für unseren Business Case wollen wir nochmals aufzeigen, wie Daten strukturiert sind, Verteilungen und Variabelarten, damit wir wissen, welche Modelle wir wie anwenden können. Unser Auftraggeber AirBnB kennt diese Daten natürlich bereits sehr gut, deshalb werden hier im Projektdokument nur die wichtigsten Plots zeigen, alle Visualisierungen finden sich im Anhang. 

*Important to mention* During our exploratory data analysis, we
identified one 'Private room' listing with an unrealistically high price
of 2'147'483'647 CHF. We excluded this single observation from our
dataset as it is an extreme outlier that does not reflect the true
market conditions and would negatively skew our predictive models.

```{r}
# This keeps everything except the row that is a Private room and costs more than 5000
d.airbnb_listings_feat <- d.airbnb_listings_feat %>%
  filter(!(room_type == "Private room" & price > 100000))
```

in addition to the one outlier we wanted to remove one NA specifically
```{r filter NA in ProfilePic}
d.airbnb_listings_feat <- d.airbnb_listings_feat %>%
  filter(!host_has_profile_pic == "")
```
 
 
## Prices
 
```{r, fig.show='hold', out.width="49%", echo=FALSE}
p.price <- ggplot(d.airbnb_listings_feat, aes(x = price)) +
  geom_histogram(bins = 50, fill = "deepskyblue", color = "white", alpha = 0.7) +
  labs(title = "Distribution of Raw Price", x = "Price", y = "Count")

p.logprice <- ggplot(d.airbnb_listings_feat, aes(x = log_price)) +
  geom_histogram(bins = 50, fill = "deepskyblue", color = "white", alpha = 0.7) +
  labs(title = "Distribution of Log Price", x = "Log Price", y = "Count")

p.price + p.logprice
```
 
The histograms above clearly illustrate that the price distribution is heavily right-skewed. To satisfy the assumptions of our models and improve performance, we applied a log transformation. Post-transformation, the data exhibits a distribution that more closely resembles a Gaussian (normal) distribution.

## Capacity & Rating

```{r, fig.show='hold', out.width='49%'}
p.acc <- ggplot(d.airbnb_listings_feat, aes(x = accommodates)) +
  geom_histogram(binwidth = 1, fill = "deepskyblue", color = "white") +
  scale_x_continuous(breaks = seq(1, max(d.airbnb_listings_feat$accommodates, na.rm = TRUE), by = 1)) +
  labs(title = "Listing Capacity", x = "Accommodates (Guests)", y = "Count")

p.rev <- ggplot(d.airbnb_listings_feat, aes(x = review_scores_rating)) +
  geom_histogram(bins = 30, fill = "deepskyblue", color = "white", alpha = 0.7) +
  labs(title = "Review Scores Rating Distribution", x = "Rating", y = "Count")

p.acc + p.rev
```
The listing capacity plot shows that most accommodations are designed for 2 or 4 guests. Regarding the review scores, a distinct trend is visible, with the vast majority of ratings falling between 4 and 5. Since both hosts and guests rate each other, we suspect this distribution may be influenced by a reciprocal rating bias (or leniency bias)

## Room Types & Superhost Status
 
```{r, fig.show='hold', out.width='49%'}
p.room <- ggplot(d.airbnb_listings_feat, aes(x = room_type, fill = room_type)) +
  geom_bar(show.legend = FALSE) +
  labs(title = "Proportion of Room Types", x = "Room Type", y = "Count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

p.host <- ggplot(d.airbnb_listings_feat %>% filter(!is.na(host_is_superhost)), aes(x = as.factor(host_is_superhost), fill = as.factor(host_is_superhost))) +
  geom_bar(show.legend = FALSE) +
  scale_fill_manual(values = c("deepskyblue", "deepskyblue")) +
  labs(title = "Superhost Status", x = "Is Superhost?", y = "Count")

p.room + p.host
```
The plots illustrate the distribution of room types and host statuses within the dataset. It is evident that the inventory is predominantly composed of `Entire homes/apartments` and `Private rooms`, while `Hotel rooms` and `Shared rooms` constitute a negligible minority. Furthermore, the data reveals that Superhost status is relatively uncommon; a significant majority of hosts do not hold this designation, with only approximately 30% of the sample categorized as Superhosts.

## Log Prices and rooms / neighbourhoods
```{r, fig.show='hold', out.width='33%'}
p.logprice_room <- ggplot(d.airbnb_listings_feat, aes(x = room_type, y = log_price, fill = room_type)) +
  geom_boxplot(alpha = 0.7, show.legend = FALSE) +
  labs(title = "Log Price by Room Type", x = "Room Type", y = "Log Price") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

d.top_neighborhoods <- d.airbnb_listings_feat %>% 
  count(neighbourhood_group_cleansed, sort = TRUE) %>% 
  top_n(10, n) %>% 
  pull(neighbourhood_group_cleansed)

p.logprice_neighbourhood <- ggplot(d.airbnb_listings_feat %>% filter(neighbourhood_group_cleansed %in% d.top_neighborhoods), 
       aes(x = reorder(neighbourhood_group_cleansed, log_price, FUN = median), y = log_price)) +
  geom_boxplot(fill = "deepskyblue", alpha = 0.7) +
  coord_flip() +
  labs(title = "Log Price by Top 10 Neighbourhood Groups", 
       x = "Neighbourhood Group", 
       y = "Log Price")

p.logprice_room | p.logprice_neighbourhood
```

The boxplot *"Log Price by Room Type"* reveals significant differences in pricing structures across categories:

- Entire homes/apartments and Private rooms: Both categories exhibit a large number of high-end outliers. In the context of the Swiss market, these likely represent luxury listings such as villas or mountain chalets. While private rooms show a similar outlier pattern, their median price remains significantly lower than that of entire homes.

- Hotel rooms: Interestingly, hotel rooms show the highest median price. The distribution is much tighter with almost no outliers, suggesting a more standardized and professional pricing strategy compared to the highly variable peer-to-peer market.

- Shared rooms: As expected, this category represents the most budget-friendly option with the lowest median and overall price range.

The second plot, *"Log Price by Top 10 Neighbourhood Group"*, highlights the impact of location on pricing:

There is a clear geographic disparity: regions like Riviera-Pays-d'Enhaut and Aigle command the highest median prices, whereas Gros-de-Vaud is the most affordable area.

Almost all top neighbourhoods show high-end outliers, confirming that luxury listings are spread across the canton, though the baseline price (the "floor") varies significantly by location.

## Host Variables

```{r host_features, fig.show='hold', out.width='16%'}
p.resp_time <- ggplot(d.airbnb_listings_feat %>% filter(host_response_time != "N/A" & !is.na(host_response_time)), aes(x = host_response_time)) +
  geom_bar(fill = "deepskyblue") +
  labs(title = "Host Response Time", x = "Response Time", y = "Count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

p.resp_rate <- ggplot(d.airbnb_listings_feat, aes(x = host_response_rate)) +
  geom_histogram(bins = 20, fill = "deepskyblue", color = "white") +
  labs(title = "Host Response Rate", x = "Rate (0 to 1)", y = "Count")

p.acc_rate <- ggplot(d.airbnb_listings_feat, aes(x = host_acceptance_rate)) +
  geom_histogram(bins = 20, fill = "deepskyblue", color = "white") +
  labs(title = "Host Acceptance Rate", x = "Rate (0 to 1)", y = "Count")

p.pic <- ggplot(d.airbnb_listings_feat %>% filter(!is.na(host_has_profile_pic)), aes(x = host_has_profile_pic)) +
  geom_bar(fill = "deepskyblue") +
  labs(title = "Has Profile Pic", x = "t = True, f = False", y = "Count")

p.verified <- ggplot(d.airbnb_listings_feat %>% filter(!is.na(host_identity_verified)), aes(x = host_identity_verified)) +
  geom_bar(fill = "deepskyblue") +
  labs(title = "Identity Verified", x = "t = True, f = False", y = "Count")

p.resp_time | p.resp_rate | p.acc_rate | p.pic | p.verified

```
The visualizations regarding host behavior and profile attributes reveal a high degree of professionalism and a strong intent to maximize booking opportunities:

Responsiveness: A vast majority of hosts respond "within an hour," with the Host Response Rate peaking sharply at 100%. This indicates that hosts are highly motivated to engage with potential guests and secure bookings.

Reliability: Similarly, the Host Acceptance Rate is predominantly concentrated at 100%, suggesting that once a request is made, hosts are very likely to confirm the stay.

Trust and Verification: To establish credibility in a peer-to-peer marketplace, nearly all hosts utilize profile pictures and have completed identity verification. These "trust signals" are essential for reducing perceived risk for bookers and presenting the host as a reliable partner.

## Properties & Neighborhoods

```{r, fig.show='hold', out.width='49%'}

p.prop <- d.airbnb_listings_feat %>%
  count(property_type) %>%
  top_n(10, n) %>%
  ggplot(aes(x = reorder(property_type, n), y = n)) +
  geom_bar(stat = "identity", fill = "deepskyblue") +
  coord_flip() +
  labs(title = "Top 10 Property Types (Count)", x = "Property Type", y = "Count")

p.neigh <- d.airbnb_listings_feat %>%
  count(neighbourhood_group_cleansed) %>%
  top_n(10, n) %>%
  ggplot(aes(x = reorder(neighbourhood_group_cleansed, n), y = n)) +
  geom_bar(stat = "identity", fill = "deepskyblue") +
  coord_flip() +
  labs(title = "Top 10 Neighbourhood Groups (Count)", x = "Neighbourhood", y = "Count")

p.prop | p.neigh
```
"To further understand the market composition, we examined the most frequent property types and their geographic distribution.

- Property Types: The plot 'Top 10 Property Types (Count)' shows a massive dominance of 'Entire rental units', which account for nearly 2,000 listings. This is followed by 'Private rooms in rental units' and 'Entire homes'. Specialized accommodations like chalets or villas are present but represent a much smaller share of the market.

- Geographic Distribution: The chart 'Top 10 Neighbourhood Groups (Count)' highlights a strong regional concentration. Aigle and Lausanne are the most prominent hubs, with Aigle leading the count at over 1,000 listings. In contrast, regions like Gros-de-Vaud show significantly less activity in the top 10 ranking.

## Furnishing

```{r, fig.show='hold', out.width='49%'}
p.baths <- ggplot(d.airbnb_listings_feat, aes(x = bathrooms)) +
  geom_histogram(binwidth = 0.5, fill = "deepskyblue", color = "white") +
  scale_x_continuous(breaks = seq(0, max(d.airbnb_listings_feat$bathrooms, na.rm = TRUE), by = 1)) +
  labs(title = "Number of Bathrooms", x = "Bathrooms", y = "Count")

p.beds <- ggplot(d.airbnb_listings_feat, aes(x = beds)) +
  geom_histogram(binwidth = 1, fill = "deepskyblue", color = "white") +
  scale_x_continuous(breaks = seq(0, max(d.airbnb_listings_feat$beds, na.rm = TRUE), by = 2)) +
  labs(title = "Number of Beds", x = "Beds", y = "Count")

p.baths | p.beds


```
The distribution plots for the number of bathrooms and beds reveal a strong concentration at lower values, as expected. However, the 'Number of Beds' plot in particular shows several extreme outliers, with a few listings claiming exceptionally high bed counts (up to over 40). While these values are not entirely implausible—they could represent large hostels or group dormitories—a qualitative verification was deemed impractical. Manually cross-referencing each of these extreme cases directly on the Airbnb website would be excessively time-consuming, hence they remain in the dataset as potential noise.

## Utilisation

```{r, fig.show='hold', out.width='33%'}
# Minimum Nights (Log scale due to extreme outliers like 365+ nights)
p.nights <- ggplot(d.airbnb_listings_feat, aes(x = minimum_nights)) +
  geom_histogram(bins = 30, fill = "deepskyblue", color = "white") +
  scale_x_log10() +
  labs(title = "Minimum Nights $(Log_10$ Scale)", x = "Minimum Nights", y = "Count")

p.num_rev <- ggplot(d.airbnb_listings_feat, aes(x = number_of_reviews)) +
  geom_histogram(bins = 30, fill = "deepskyblue", color = "white") +
  labs(title = "Total Number of Reviews", x = "Reviews", y = "Count")

p.occ <- ggplot(d.airbnb_listings_feat, aes(x = estimated_occupancy_l365d)) +
  geom_histogram(bins = 30, fill = "deepskyblue", color = "white") +
  labs(title = "Estimated Occupancy", x = "Occupied Days", y = "Count")

p.nights | p.num_rev | p.occ
```

The histograms illustrate the distributions of booking requirements, guest engagement, and listing utilization.

Minimum Nights ($Log_10$ Scale): To appropriately visualize this variable, a logarithmic transformation (base 10) was applied. This methodological choice was necessary due to severe right-skewness caused by extreme outliers, notably listings that enforce a 365-day minimum stay. Despite this transformation, the distribution clearly shows that the vast majority of hosts prefer short-term stays, with the highest frequencies concentrated around 1 to 3 nights.

Total Number of Reviews: The review count distribution is heavily right-skewed, indicating an L-shaped distribution. A large portion of listings has very few reviews, while a small minority of highly established listings accumulates a massive review volume. This suggests that the market is dominated by a few 'super-performers' alongside many newer or less frequently booked properties.

Estimated Occupancy: Similar to the review counts, the estimated occupancy shows a strong rightward skew. Most listings are occupied for only a small fraction of the year. However, a notable anomaly is the slight spike at the far right tail of the distribution. This peak likely represents highly commercialized listings or permanent rental units that operate at near full capacity throughout the year."

## Feature Engineered Data

```{r, fig.show='hold', out.width='49%'}

# High Demand (Binary)
p.demand <- ggplot(d.airbnb_listings_feat, aes(x = as.factor(high_demand))) +
  geom_bar(fill = "deepskyblue") +
  labs(title = "High Demand Classification", x = "0 = No, 1 = Yes", y = "Count")

# Price High (Binary)
p.price_high <- ggplot(d.airbnb_listings_feat, aes(x = as.factor(price_high))) +
  geom_bar(fill = "deepskyblue") +
  labs(title = "High Price Classification", x = "0 = No, 1 = Yes", y = "Count")

p.demand | p.price_high
```

The bar charts illustrate the count of observations for our classification targets:

*High Demand Classification*:
This variable exhibits a severe class imbalance. While the majority of listings are classified as "0" (low demand), only a small fraction meets the threshold for "1" (high demand). From a methodological standpoint, this imbalance suggests that Accuracy will be a misleading metric, as a model could achieve high accuracy by simply predicting the majority class. Consequently, we will prioritize the F1-Score and Precision-Recall curves to evaluate model performance for this target.

*High Price Classification*:
In contrast, the distribution for high price is nearly perfectly balanced. This is the intended result of our decision to split the data using the median price. Since exactly half of the listings fall above and half below this value, the classes are equally represented. This provides an ideal baseline for classification algorithms, as the models will not be inherently biased toward a majority class.

## Verifications

```{r, fig.show='hold'}
p.verifications <- d.airbnb_listings_feat %>%
  count(host_verifications) %>%
  top_n(5, n) %>%
  ggplot(aes(x = reorder(host_verifications, n), y = n)) +
  geom_bar(stat = "identity", fill = "deepskyblue") +
  coord_flip() +
  labs(title = "Top 5 Host Verification Combinations", x = "Verification Methods", y = "Count")

p.verifications
```

To complete the profile of host credibility, we examined the specific combinations of verification methods used. The chart "Top 5 Host Verification Combinations" reveals a highly standardized approach to trust within the platform:

- Platform Standards: The vast majority of hosts rely on the dual verification of email and phone. This combination represents the baseline requirement for establishing a functional and reachable profile on Airbnb.

- Professionalism Indicator: A smaller but notable segment includes a "work_email" in their verification set. This often serves as a proxy for professional or commercial hosts who manage listings as part of a business entity rather than as private individuals.

- Minimalist Profiles: Profiles with only one or no verification methods are extremely rare, suggesting that the community (and likely the platform's own ranking algorithms) enforces a minimum level of identity disclosure to maintain marketplace integrity.

# Data Quality

To ensure the integrity of our analysis, we performed a systematic check for missing values (NAs) within the dataset. We utilized the `naniar` and `visdat` packages, which allow for both a numerical summary and a spatial visualization of missingness patterns.

Visualization and Identification: Using `vis_dat()` and `gg_miss_var()`, we identified that the missing data is not uniformly distributed but concentrated in specific features.

Response and Acceptance Rates: A high volume of NAs was observed in the "Rate" columns (e.g., host response and acceptance rates). An initial hypothesis suggested that these missing values might represent zero values that were not correctly captured during the data import process. While imputing these as 0% was considered, we proceeded with caution to avoid distorting host performance metrics.

Review Ratings and Methodological Decision: For the review score variables, we determined that the data is likely Missing Not At Random (MNAR). A listing without reviews often indicates it is new or has not yet been booked, meaning a "rating" simply does not exist for that observation.

Final Handling: We decided to omit observations with missing ratings from our models rather than imputing them. Imputing a zero would incorrectly categorize these listings as "poorly rated," while imputing a mean would artificially reduce the variance of our predictors. By omitting these cases, we ensure that our model training is based on verified, existing performance data.

```{r}
vis_miss(d.airbnb_listings_feat)
gg_miss_var(d.airbnb_listings_feat)
gg_miss_upset(d.airbnb_listings_feat)
```
### Aus Understanding the Data
```{r}
# Calculate percentage of missing values per column
missing_data <- d.airbnb_listings_feat %>%
  summarise(across(everything(), ~ mean(is.na(.)) * 100)) %>%
  pivot_longer(everything(), names_to = "Variable", values_to = "Missing_Percentage") %>%
  filter(Missing_Percentage > 0) %>%
  arrange(desc(Missing_Percentage))

# Plot missing values
p.missing_values <- ggplot(missing_data, aes(x = reorder(Variable, Missing_Percentage), y = Missing_Percentage)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +
  labs(title = "Percentage of Missing Values per Variable",
       x = "Variables",
       y = "% Missing")
p.missing_values
```
### Missing Values (NA) Check

This step is less for the client and essential for your ML models. Most
modeling functions (like lm() or svm()) will throw errors or silently
drop entire rows if NAs are present.

```{r}
# Count the number of NAs for each key column in your dataset
na_summary <- d.airbnb_listings_feat %>%
  select(price, log_price, accommodates, number_of_reviews, host_is_superhost) %>%
  summarise(across(everything(), ~ sum(is.na(.)))) %>%
  t() %>% # Transpose for better readability
  as.data.frame() %>%
  rename(Missing_Values = V1)

kable(na_summary, caption = "Count of Missing Values per Variable")
```

# Exploratory Data Analysis (DAVIDE)

### Overall Market Overview (Prices & Capacity)

This is the absolute minimum, maximum, median, and mean for the most
important numeric variables.

```{r}
# Basic summary statistics for key numerical variables
summary_stats <- d.airbnb_listings_feat %>%
  select(price, log_price, accommodates, number_of_reviews) %>%
  summary()

print(summary_stats)
```


### Analysis by Room Type

Since we want to know how many different types of rooms there are, we
group the data and calculate the counts, percentages, and price ranges
for each type.

```{r}
# Detailed breakdown by Room Type
room_type_analysis <- d.airbnb_listings_feat %>%
  group_by(room_type) %>%
  summarise(
    Total_Listings = n(),
    Percentage_of_Market = round((n() / nrow(d.airbnb_listings_feat)) * 100, 1),
    Min_Price = min(price, na.rm = TRUE),
    Median_Price = median(price, na.rm = TRUE),
    Average_Price = round(mean(price, na.rm = TRUE), 1),
    Max_Price = max(price, na.rm = TRUE)
  ) %>%
  arrange(desc(Total_Listings)) # Sorts by the most common room type

print(room_type_analysis)
```

Client Interpretation: "Entire homes/apartments dominate the Vaud
market, making up 75% of all listings. As expected, the Hotel room
command the highest median price (264.5 CHF), compared to shared rooms
(47 CHF)."

### Detailed property_type Analysis

We already analyzed room_type, but Airbnb often has dozens of
property_types (e.g., "Entire chalet", "Private room in castle", etc.).
Summarizing the Top 10 property types in the Canton of Vaud provides
maybe a great market insight.

```{r}
# Top 10 Property Types by frequency and their median price
top_properties <- d.airbnb_listings_feat %>%
  group_by(property_type) %>%
  summarise(
    Total_Listings = n(),
    Median_Price = median(price, na.rm = TRUE)
  ) %>%
  arrange(desc(Total_Listings)) %>%
  slice_head(n = 10) # Keep only the top 10

kable(top_properties, caption = "Top 10 Property Types in the Canton of Vaud")
```


Client Interpretation: "While standard apartments like "entire rental
unit" are the most common, specialty properties like chalets, homes and
villas command a significant premium, which our model will need to
capture."


### Correlation Matrix (Crucial before Machine Learning)

Before we train our LM, GAM, or SVM, we need to know how our numerical
variables relate to each other. This helps us identify multicollinearity
(when two predictors provide the same information) and strong price
drivers.

```{r}
# Select relevant numerical variables (drop NAs just for the correlation calculation)
cor_data <- d.airbnb_listings_feat %>%
  select(log_price, accommodates, number_of_reviews) %>%
  drop_na()

# Calculate the correlation matrix
cor_matrix <- cor(cor_data)

# Render a nice table for your RMarkdown document
kable(round(cor_matrix, 2), caption = "Correlation Matrix of Key Numerical Drivers")
```


```{r correlation_heatmap, fig.height=8, fig.width=10}
# Select all numerical variables, drop binaries we already analysed separately
num_vars <- d.airbnb_listings_feat %>%
  select(where(is.numeric)) %>%
  select(-price_high, -high_demand)

# Correlation matrix with pairwise complete observations to handle NAs
cor_matrix_full <- cor(num_vars, use = "pairwise.complete.obs")

# Reshape for ggplot
cor_long <- as.data.frame(cor_matrix_full) %>%
  rownames_to_column("Var1") %>%
  pivot_longer(-Var1, names_to = "Var2", values_to = "Correlation")

ggplot(cor_long, aes(x = Var1, y = Var2, fill = Correlation)) +
  geom_tile(color = "white") +
  geom_text(aes(label = round(Correlation, 2)), size = 2.8) +
  scale_fill_gradient2(low = "darkred", mid = "white", high = "steelblue",
                       midpoint = 0, limits = c(-1, 1)) +
  labs(title = "Correlation Heatmap of Numerical Variables",
       x = NULL, y = NULL) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Interpretation:
The heatmap confirms what we already suspected: *accommodates, beds and bathrooms* **correlate strongly** with each other. This is a clear sign of multicollinearity, which we need to keep in mind for the LM. The variable *price* and *log_price* show the expected relationships with *capacity* variables, while review-related variables form their own cluster. *number_of_reviews* and *estimated_occupancy_l365d* are highly correlated — listings that get booked more often naturally collect more reviews.

### Log Price vs. Accommodates (Scatter)
The most important Relationship in the dataset from what we see in the correlation matrix above. 

```{r price_vs_accommodates}
ggplot(d.airbnb_listings_feat, aes(x = accommodates, y = log_price)) +
  geom_jitter(alpha = 0.3, width = 0.2, color = "steelblue") +
  geom_smooth(method = "lm", color = "darkred", se = TRUE) +
  scale_x_continuous(breaks = seq(1, max(d.airbnb_listings_feat$accommodates, na.rm = TRUE), 2)) +
  labs(title = "Log Price vs. Accommodates",
       x = "Number of Guests",
       y = "Log Price")
```

### Log Price by Review Score Group (H1)
We display review_scores_rating as a histogram, but never in relation to the price.

```{r price_by_rating}
d.airbnb_listings_feat %>%
  filter(!is.na(review_scores_rating)) %>%
  mutate(rating_group = cut(review_scores_rating,
                            breaks = c(0, 3, 4, 4.5, 4.8, 5),
                            labels = c("≤3", "3–4", "4–4.5", "4.5–4.8", "4.8–5"),
                            include.lowest = TRUE)) %>%
  ggplot(aes(x = rating_group, y = log_price, fill = rating_group)) +
  geom_boxplot(alpha = 0.7, show.legend = FALSE) +
  labs(title = "Log Price by Review Score Group",
       x = "Review Score Rating",
       y = "Log Price")
```

Interpretation:
The medians across rating groups are surprisingly close. Higher-rated listings do not automatically charge more — which is consistent with the pattern we later see in the Superhost analysis: well-rated hosts often compete on price rather than charge a premium.


### Spatial Data: Mapping the Listings by Log(Price)
```{r}
# Cleaned up the map to focus purely on location and price.
ggplot(data = d.airbnb_listings_feat, aes(x = longitude, y = latitude)) +
  geom_point(aes(color = log_price), alpha = 0.6, size = 1.5) +
  scale_color_viridis_c(option = "plasma") +
  labs(
    title = "Airbnb Listings in Vaud by Log(Price)",
    x = "Longitude",
    y = "Latitude",
    color = "Log(Price)"
  ) +
  theme_minimal()
```

### Price per Guest Distribution
We calculated price per guest as a feature and visualize it here. 

```{r price_per_guest}
p_ppg <- ggplot(d.airbnb_listings_feat, aes(x = price_per_guest)) +
  geom_histogram(bins = 50, fill = "darkorange", color = "white", alpha = 0.7) +
  scale_x_log10() +
  labs(title = "Price per Guest (Log Scale)",
       x = "Price per Guest (CHF)", y = "Count")

p_ppg_room <- ggplot(d.airbnb_listings_feat, aes(x = room_type, y = price_per_guest, fill = room_type)) +
  geom_boxplot(alpha = 0.7, show.legend = FALSE) +
  scale_y_log10() +
  labs(title = "Price per Guest by Room Type",
       x = "Room Type", y = "Price per Guest (Log)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

p_ppg + p_ppg_room
```

Interpretation:
Most listings settle between 30 and 100 CHF per guest. Hotel rooms charge the highest per-guest price, while shared rooms are clearly the cheapest option per head. Entire homes look more expensive overall but actually offer the best value when you split the cost across guests.


### Superhost vs. Regular Host Breakdown

This business case specifically mentions superhost status. So we took a
look to see how many there are and if they actually charge more or even
less.

```{r}
# Superhost statistics
superhost_analysis <- d.airbnb_listings_feat %>%
  group_by(host_is_superhost) %>%
  summarise(
    Total_Hosts = n(),
    Percentage = round((n() / nrow(d.airbnb_listings_feat)) * 100, 1),
    Median_Price = median(price, na.rm = TRUE),
    Avg_Reviews = round(mean(number_of_reviews, na.rm = TRUE), 1)
  ) %>%
  mutate(host_is_superhost = ifelse(host_is_superhost == 1, "Yes", "No"))

print(superhost_analysis)
```


```{r Binary Predictor: Superhost Status}
# Adding this because it's a key part of the business case and fulfills the binary variable requirement.
ggplot(data = d.airbnb_listings_feat, aes(x = as.factor(host_is_superhost), y = log_price, fill = as.factor(host_is_superhost))) +
  geom_boxplot(alpha = 0.7) +
  scale_x_discrete(labels = c("0" = "Regular Host", "1" = "Superhost")) +
  labs(
    title = "Price Difference: Superhosts vs. Regular Hosts",
    x = "Host Status",
    y = "Log(Price)"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```





# LM (TIMON - Davide translated to english and commments)

## Theory

With this initial model, we aim to examine the first hypothesis and investigate which host-centered indicators influence the price of a rental property. 

For this, we use the *review_scores_rating*. We
also include the factor *host_is_superhost* to distinguish between the
two groups. We measure willingness to pay here using the effective price
of the rental property. Additionally, we introduce two control variables. First, the size of the rental property, as this obviously influences and skews the rental price, and second, the type of room. This also significantly influences the price.

Since there is a high risk of multicollinearity in our data, this is
also examined.


**Hier muss noch ein Teil rein bezüglich den Vorgaben des Models und inwiefern diese erfüllt sind.**


## calculating Models

```{r lm controlled by Timon}
# Fitting a Multiple Linear Regression model to predict the log-transformed price.
# We use log(price) to account for the right-skewed distribution of Airbnb prices.
lm.main <- lm(log(price) ~ 
                review_scores_rating + 
                as.factor(host_is_superhost) +
                host_has_profile_pic + 
                accommodates + 
                as.factor(room_type), 
              data = d.airbnb_listings_feat)

summary(lm.main)


# Checking for multicollinearity using Variance Inflation Factor (VIF).
# Values range from 1 to infinity. A value of exactly 1 indicates no collinearity.
# Values > 5-10 suggest problematic multicollinearity among the predictor variables.
multicollinearity(lm.main)
```

### LM Controlled Plots (TIMON - Davide translated to english and commments)

```{r}
# --- LM controlled plots ---
# Visualizing the raw relationship between review scores, log(price), and superhost status.
ggplot(data = d.airbnb_listings_feat, mapping = aes(review_scores_rating, log(price), color = host_is_superhost)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "lm")

# Creating a grid of hypothetical data to predict and plot the fitted LM lines cleanly.
# We hold 'room_type' and 'accommodates' constant to isolate the specific effects of reviews and superhost status.
d.lm_predict <- expand.grid(
  review_scores_rating = seq(3, 5, length.out = 100),
  host_is_superhost = c("TRUE", "FALSE"),
  room_type = "Entire home/apt",   # Fixed reference category
  host_has_profile_pic = c("t", "f"),
  accommodates = mean(d.airbnb_listings_feat$accommodates, na.rm = TRUE) # Fixed at mean value
)

# Generating predictions based on the trained LM model
d.lm_predict$pred <- predict(lm.main, newdata = d.lm_predict)

# Plotting the actual data points alongside our model's predicted regression lines
ggplot(d.airbnb_listings_feat, 
       aes(review_scores_rating, log(price), color = host_is_superhost)) +
  geom_point(alpha = 0.3) +
  geom_line(data = d.lm_predict, aes(y = pred), size = 1.2) +
  labs(title = "LM Predictions: Log(Price) vs. Review Rating",
       x = "Review Scores Rating",
       y = "Log(Price)")
```

## Interpretation
The model shows that both the hosts’ rating and their Superhost status have a significant influence. The rating has a positive effect of *`r round(lm.main$coefficients[2], digits = 3)`*. In contrast, Superhost status has a negative effect of *`r round(lm.main$coefficients[3], digits = 3)`*. The assumption that tenants are willing to pay more when renting from a Superhost therefore does not necessarily seem to hold true. One possible explanation for this phenomenon is that Superhosts may have a better understanding of the market because they have been active for a longer time, and therefore set lower prices in order to achieve a higher occupancy rate.
This is relevant for Airbnb, as these results show that reviews on the host side are a key factor in price setting. It is also interesting that Superhosts apparently develop their own understanding for how prices can be set. The exact reasons underlying this difference cannot be examined within the scope of this paper. However, these findings provide Airbnb with clarified insights that can serve as a basis for further work.

# GLM Poisson (PASCAL - Davide translated to english and commments)

## Theory
With the GLM model with Poisson-family / quasi-poisson family is used to analyse count data. Count data is strictly positive and counted (not continuous). Using this model we want to see which variables lead to the highest occupancy, meaning most of the time the airbnb is rented out. As the data is probably overdispersed - meaning the data has not the same variance and much more cheap listings than expensive ones - we need to use a quasi-poisson model that can handle this kind of data much better. First we fit a normal poisson-family, then check for overdispersion, and if necessary we use a quasi-poisson-family.

## Model
```{r}

glm.poisson <- glm(estimated_occupancy_l365d ~ price + host_is_superhost + 
                     review_scores_rating + accommodates + room_type, 
                   data = d.airbnb_listings_feat, 
                   family = "poisson")

summary(glm.poisson)


# Checking for Overdispersion.
dispersion_ratio <- glm.poisson$deviance / glm.poisson$df.residual
print(paste("Dispersion Ratio:", dispersion_ratio))

glm.quasipoisson <- glm(estimated_occupancy_l365d ~ price + host_is_superhost + 
                          review_scores_rating + accommodates + room_type, 
                        data = d.airbnb_listings_feat, 
                        family = "quasipoisson")

summary(glm.quasipoisson)
AIC(glm.quasipoisson)
```

## Crossvalidation


```{r Crossvalidation Datasets 10-fold (TIMON)}
set.seed(1)

cv.mse.poisson <- c()
cv.mse.quasipoisson <- c()

for(i in 1:200){
  #dataprep
  n <- nrow(d.airbnb_listings_feat)
  cv.ind_permuted <- sample(1:n, replace = FALSE)
  d.airbnb_listings_feat_permuted <- d.airbnb_listings_feat[cv.ind_permuted,]
  
  #create folds
  cv.folds <- cut(1:n, breaks = 3, labels = FALSE)
  cv.ind_test <- which(cv.folds == 1)
  
  d.cv_test <- d.airbnb_listings_feat_permuted[cv.ind_test, ]
  d.cv_train <- d.airbnb_listings_feat_permuted[-cv.ind_test, ]
  
  #train model poisson
  cv.glm_poisson_train <- glm(formula = formula(glm.poisson), 
                          data = d.cv_train)
  cv.glm_poisson_predict <- predict(cv.glm_poisson_train, newdata = d.cv_test)
  
  cv.mse.poisson[i] <- mean((d.cv_test$estimated_occupancy_l365d - cv.glm_poisson_predict)^2, 
                           na.rm = TRUE)
  
  
  #train model quasipoisson
  cv.glm_quasipoisson_train <- glm(formula = formula(glm.quasipoisson), 
                          data = d.cv_train)
  cv.glm_quasipoisson_predict <- predict(cv.glm_quasipoisson_train, newdata = d.cv_test)
  
  cv.mse.quasipoisson[i] <- mean((d.cv_test$estimated_occupancy_l365d - cv.glm_quasipoisson_predict)^2, 
                           na.rm = TRUE)
  
}

boxplot(cv.mse.poisson, cv.mse.quasipoisson, names = c("Poisson-Model","Quasipoisson-Model"))

```



## Interpretation GLM-Poisson:

As expected, the GLM with a Poisson distribution produced overdispersion, since the variability in the Airbnb data was to be expected. The model calculated the standard errors too small, which is why all variables were highly significant in the first output. The dispersion was then calculated and found to be greater than 1, specifically 56.53. A quasi-Poisson model was used as a solution, which adjusts the standard errors upward.

## Interpretation GLM-Quasipoisson

The GLM with quasi-Poisson distribution has corrected the overdispersion. The standard errors were adjusted upwards (significantly) and now provide a realistic picture of which factors truly have an influence on occupancy. Interpretations: 
- Price -> highly significant negative. As expected: The higher the price, the lower the occupancy
- Superhost: highly significant positive: Attention, logged rates, therefore calculate the effect with the exponential function exp(0.9577) = 2.6. Being a Superhost increases occupancy by 2.6 times compared to a normal host.
- Reviews: Positive and significant. Better reviews lead to more bookings.
- Capacity: Positive and significant. The more people that fit, the more bookings. 
- Room types: In the intercept is "Entire home/apt". Private Room is booked significantly worse than the entire accommodation. Shared Room has a high p-value. Statistically, it cannot be said whether the occupancy of a shared room differs from the entire apartment.

## Interpretation Cross Validation
Interpretation: die beiden Modelle unterscheiden sich nicht. wir erhalten keine outlier in den vorhersagen sprich das Model ist sehr stabil. 

## Business Case

The *glm.quasipoisson model* directly supports our revenue management objectives by identifying the core demand-side drivers of estimated_occupancy_l365d. It strongly supports H1 by demonstrating that host-centered indicators drastically increase demand; for example, achieving host_is_superhost status boosts expected occupancy by approximately 2.6 times. The model also confirms H2, showing that physical attributes like larger accommodates capacities and renting an entire home over a Private room significantly improve booking rates. These findings provide hosts with actionable, non-pricing levers to optimize their overall yield. Ultimately, improving review_scores_rating or attaining Superhost status offers hosts a powerful "occupancy buffer," allowing them to command higher prices while maintaining strong booking volumes.



## Limitations

A primary limitation of this model is the potential for reverse causality, as higher occupancy naturally generates more reviews, making it easier to maintain high ratings and Superhost status. Additionally, the deletion of 714 observations due to missingness could introduce survivorship bias if these removed data points represent poorly managed or unsuccessful listings. Furthermore, our feature set omits critical demand drivers like precise neighborhood desirability, seasonality, and listing photo quality, meaning unobserved factors likely account for significant residual variance. Because the Quasi-Poisson approach uses a quasi-likelihood function, it does not produce an AIC value (shown as AIC: NA), which complicates direct statistical comparisons with non-quasi models. A potential next step could be exploring a Negative Binomial model as an alternative method to handle overdispersion while retaining standard likelihood metrics for easier comparison.


# GLM Binomial (DAVIDE)

## Theory

### Why GLM for Airbnb Pricing?
When the outcome we want to model is binary — yes/no, above/below — a standard linear regression no longer fits. A Binomial GLM with a logit link is the natural choice: it bounds predictions between 0 and 1 and returns coefficients we can read as log-odds. Exponentiating them gives odds ratios, which are easier to communicate to a non-technical audience.

### Justification for our Use Case
We want to know which listings end up in the upper half of the market. To do that, we created the binary variable *price_high*, which equals 1 if a listing's price is above the median and 0 otherwise. The split at the median gives us two equally sized classes, so the model is not biased toward a majority outcome.


### Link to Hypotheses
This model speaks mainly to H1: do host-centered indicators (Superhost status, ratings) push up the price? The control variables (*accommodates*, *room_type*) are not the focus but make sure we don't confuse host effects with listing characteristics.

## Model
Feature Engeneering

```{r glm_binomial_extended}
# Predict the probability that a listing is priced above the market median
# We control for the most obvious price drivers so the host coefficients can be read cleanly
glm.binomial <- glm(price_high ~ host_is_superhost + 
                                 accommodates + 
                                 room_type + 
                                 review_scores_rating,
                    data   = d.airbnb_listings_feat,
                    family = binomial(link = "logit"))

summary(glm.binomial)
```

```{r glm_binomial_results}
# Convert log-odds into odds ratios with 95% confidence intervals for easier reading
d.results_glm_binomial <- tidy(glm.binomial, conf.int = TRUE) %>%
  mutate(
    odds_ratio = round(exp(estimate), 3),
    or_lower   = round(exp(conf.low), 3),
    or_upper   = round(exp(conf.high), 3),
    p.value    = round(p.value, 4)
  ) %>%
  select(term, estimate, odds_ratio, or_lower, or_upper, p.value)

kable(d.results_glm_binomial, 
      caption = "GLM Binomial: Odds Ratios with 95% Confidence Intervals")
```

## Visualisation
A coefficient table is hard to read for hosts. The plot below shows the predicted probability of charging an above-median price across listing sizes, split by Superhost status. Everything else is held at its average so the two lines differ only in the variable we care about.

```{r glm_binomial_plot}
# Build a prediction grid: vary accommodates and superhost, hold the rest constant
d.glm_predict <- expand.grid(
  accommodates         = seq(1, 10, by = 1),
  host_is_superhost    = c(TRUE, FALSE),
  room_type            = "Entire home/apt",
  review_scores_rating = mean(d.airbnb_listings_feat$review_scores_rating, na.rm = TRUE)
)

# Predict on the link scale to also get standard errors, then transform back
glm.pred <- predict(glm.binomial, newdata = d.glm_predict, type = "link", se.fit = TRUE)
d.glm_predict$prob       <- plogis(glm.pred$fit)
d.glm_predict$prob_lower <- plogis(glm.pred$fit - 1.96 * glm.pred$se.fit)
d.glm_predict$prob_upper <- plogis(glm.pred$fit + 1.96 * glm.pred$se.fit)

p.glm_binomial <- ggplot(d.glm_predict, aes(x = accommodates, y = prob, 
                                            color = host_is_superhost, fill = host_is_superhost)) +
  geom_ribbon(aes(ymin = prob_lower, ymax = prob_upper), alpha = 0.2, color = NA) +
  geom_line(size = 1.2) +
  scale_x_continuous(breaks = 1:10) +
  scale_color_manual(values = c("tomato", "seagreen"),
                     labels = c("Regular Host", "Superhost")) +
  scale_fill_manual(values = c("tomato", "seagreen"),
                    labels = c("Regular Host", "Superhost")) +
  labs(title    = "Probability of Charging an Above-Median Price",
       subtitle = "By guest capacity and Superhost status (entire home, average rating)",
       x        = "Accommodates (Guests)",
       y        = "Predicted Probability",
       color    = NULL, fill = NULL) +
  theme(legend.position = "top")

p.glm_binomial
```


## Interpretation GLM Binomial: Calculating Odds Ratios (DAVIDE)

After controlling for size, room type and rating, Superhosts have **31% lower odds** of charging an above-median price (OR = 0.69, CI [0.57, 0.84], p = 0.0002). 

The control variables behave as expected: each additional guest nearly doubles the odds of being in the upper price half (OR = 1.83), private and shared rooms are far less likely to land above the median, and an one-point increase in rating raises the odds by about 62%. The hotel-room coefficient is significant but its wide confidence interval (1.26 to 40.6) reflects how few hotel listings are in the dataset.

This is the opposite of what H1 would predict. Superhost status does not raise willingness to pay, it correlates with charging less. Combined with Pascal's Quasi-Poisson result (Superhosts get 2.6× more bookings), the picture is consistent: Superhosts compete on price and earn through volume, not through a price premium. The practical takeaway for hosts aiming for the badge is to price competitively and let occupancy do the work.


## Limitations
A few things to keep in mind:

- Median split is rough. A listing one franc above the median counts the same as one 500 francs above. The LM and GAM handle this much better.
- No location. We left location out to keep the focus on host variables, but in Vaud, location is a major price driver. Some of what we attribute to room type or capacity is probably picking up location indirectly.
- Snapshot only. No seasonality, no weekday effects, no events.
- Correlation, not causation. Superhosts charge less on average but we can't say the badge causes it. It's just as likely that hosts who already price competitively end up qualifying for the badge.
- The cutoff moves with the market. The median is just our sample's median. If the market shifts, so does it.


# GAM Model (DAVIDE)

## Theory
### Why GAM for Airbnb Pricing?
A standard linear model assumes that every predictor moves the price at a constant rate. For Airbnb data that's almost never true. A Generalized Additive Model (GAM) replaces the fixed coefficients with smooth functions that the model learns from the data itself. Each predictor gets its own curve instead of a single slope, which lets the model bend where the data actually bends.

The result sits between a linear regression and a black-box model: more flexible than the LM, more interpretable than a Random Forest or Neural Network. Each smooth term can still be plotted on its own and read like a coefficient.

### Justification for our Use Case
Three things in our data make a GAM the right fit:

1. Price effects are non-linear. Going from 2 to 4 guests roughly doubles the price; going from 8 to 10 barely moves it. Review scores show a similar pattern. A drop from 4.0 to 3.5 hurts much more than the difference between 4.8 and 5.0. A linear model averages these curves and gets both ends wrong.
2. Location is non-linear by nature. Vaud combines lakeshore towns, urban centres, alpine resorts, and rural areas, each with their own price logic. A 2D smoother over longitude and latitude captures this surface without us having to engineer postcode dummies or distance-to-lake variables. This is the main reason we picked GAM over LM.
3. Interpretability stays. Hosts should be able to see why a price is recommended, not just get a number from a black box. The partial effect plots make every smooth term readable on its own.

### Link to Hypotheses
The GAM speaks mainly to **H2** through the spatial spline *s(longitude, latitude)*, while controlling for **H1** via *host_is_superhost* and *review_scores_rating*. It is our main tool for checking whether location effects on willingness to pay are non-linear and clustered, something a linear model cannot answer.
  
## Model

```{r gam_data_split}
# Data preparation: keep only variables used in the GAM, drop rows with NAs
d.airbnb_listings_feat_gam <- d.airbnb_listings_feat %>%
  select(price, accommodates, review_scores_rating,
         number_of_reviews, longitude, latitude,
         room_type, host_is_superhost, bathrooms) %>%
  na.omit()

# Train/test split BEFORE fitting the model.
# This matters for the SVM later, which uses GAM labels — splitting first prevents leakage.
set.seed(42)
n <- nrow(d.airbnb_listings_feat_gam)
train.idx <- sample(seq_len(n), size = floor(0.8 * n))

df.train <- d.airbnb_listings_feat_gam[ train.idx, ]
df.test  <- d.airbnb_listings_feat_gam[-train.idx, ]

cat("Train:", nrow(df.train), " | Test:", nrow(df.test), "\n")
```

We model *log(price)* rather than the raw price. The price distribution is heavily right-skewed (see EDA-Part), and the log transform stabilises variance and brings the residuals close to normal. Coefficients on the log scale can be read multiplicatively: a coefficient of β corresponds to roughly a (exp(β) − 1) × 100% price change.
We use *method = "REML"* instead of the default. REML is less prone to undersmoothing and gives more reliable confidence intervals

```{r gam_fit}
gam.main <- gam(log(price) ~
                  s(accommodates) +
                  s(review_scores_rating) +
                  s(log1p(number_of_reviews)) +
                  s(longitude, latitude) +
                  room_type +
                  host_is_superhost +
                  bathrooms,
                family = gaussian(),
                method = "REML",
                data   = df.train)

summary(gam.main)
```

## Visualisation
The partial effect plots are the main reason to choose a GAM in the first place. Each smooth term is shown on its own, with the shaded area representing the 95% confidence band.

```{r gam_partial_effects}
plot(gam.main, pages = 1, shade = TRUE,
     col = "steelblue", shade.col = "lightblue", seWithMean = TRUE)

edf.table <- as.data.frame(summary(gam.main)$s.table) %>% round(3)
kable(edf.table, caption = "Smooth Terms: Estimated Degrees of Freedom (EDF)")

cat("R-squared:         ", round(summary(gam.main)$r.sq,  3), "\n")
cat("Deviance explained:", round(summary(gam.main)$dev.expl, 3), "\n")
```

To make the comparison fair, we fit a linear model on the exact same data, target, and predictors. The only difference is the absence of smooth terms.

```{r gam_lm_benchmark}
# Fair LM benchmark (same target, same training data)
lm.gam_comparison <- lm(log(price) ~ accommodates + review_scores_rating +
                          log1p(number_of_reviews) + longitude + latitude +
                          room_type + host_is_superhost + bathrooms,
                        data = df.train)

cat("LM  R²:", round(summary(lm.gam_comparison)$r.squared, 3), "\n")
cat("GAM R²:", round(summary(gam.main)$r.sq, 3), "\n")
AIC(lm.gam_comparison, gam.main)
```


## Traffic Light Map
The GAM predicts an expected price for every listing. Comparing actual to expected gives hosts a quick verdict — underpriced, fair, or overpriced.
We set the threshold at ±15%, anchored in the residual standard deviation of the GAM on the log scale (exp(σ) − 1 ≈ 15%). Listings outside this band deviate by more than one standard deviation from the model's expected price. The threshold is data-driven, but it remains a choice. A stricter or looser band would shift the proportions accordingly. This is one of the limitations we discuss below.

```{r traffic_light_labels}
# Predictions on log scale
df.train$fitted_logprice <- as.numeric(predict(gam.main, newdata = df.train))
df.test$fitted_logprice  <- as.numeric(predict(gam.main, newdata = df.test))

# Smearing correction (Duan 1983) for back-transforming log predictions
sigma2   <- summary(gam.main)$scale
smearing <- exp(sigma2 / 2)
df.train$fitted_price <- exp(df.train$fitted_logprice) * smearing
df.test$fitted_price  <- exp(df.test$fitted_logprice)  * smearing

# Apply traffic-light rule consistently to train and test
make.traffic_light <- function(price, fitted, threshold = 15) {
  pct <- (as.numeric(price) - as.numeric(fitted)) / as.numeric(fitted) * 100
  case_when(
    pct <= -threshold ~ "Underpriced",
    pct >=  threshold ~ "Overpriced",
    TRUE              ~ "Fair Price"
  ) %>% as.factor()
}

df.train$pct_diff      <- (df.train$price - df.train$fitted_price) / df.train$fitted_price * 100
df.test$pct_diff       <- (df.test$price  - df.test$fitted_price)  / df.test$fitted_price  * 100
df.train$traffic_light <- make.traffic_light(df.train$price, df.train$fitted_price)
df.test$traffic_light  <- make.traffic_light(df.test$price,  df.test$fitted_price)

# Sanity check: distribution should look similar across train and test
cat("\nTrain:\n"); print(prop.table(table(df.train$traffic_light)))
cat("\nTest:\n");  print(prop.table(table(df.test$traffic_light)))
```


## Interactive Map of Canton Vaude with the GAM
The map below plots every listing in the dataset, coloured by its traffic-light verdict. Each point shows the actual price, the GAM-recommended price, and the percentage deviation when clicked.

```{r gam_leaflet_map}
df.map <- bind_rows(df.train, df.test)
color_map <- c("Underpriced" = "#EF9F27",
               "Fair Price"  = "#1D9E75",
               "Overpriced"  = "#D85A30")
df.map$marker_color <- color_map[as.character(df.map$traffic_light)]

leaflet() %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  setView(lng = 6.65, lat = 46.52, zoom = 9) %>%
  addCircleMarkers(
    data = df_full,
    lng = ~longitude, lat = ~latitude,
    radius = 4, fillColor = ~marker_color,
    fillOpacity = 0.85, stroke = FALSE,
    popup = ~paste0(
      "<b>Host Verdict:</b> ", traffic_light, "<br>",
      "<b>Actual Price:</b> CHF ", round(price), "<br>",
      "<b>GAM Recommended:</b> CHF ", round(fitted_price), "<br>",
      "<b>Difference:</b> ", round(pct_diff, 1), "%"
    )
  ) %>%
  addLegend(
    position = "bottomright",
    colors   = color_map,
    labels   = c("Underpriced (Increase)", "Fair Price (±15%)", "Overpriced (Decrease)"),
    title    = "Pricing Strategy vs. Market (GAM)",
    opacity  = 0.85
  )

# Note: df_train and df_test are passed on to the SVM block.
```

## Interpretation
The GAM explains 66.4% of the deviance (adjusted R² = 0.658) on 2,527 training listings. That's a clear improvement over the LM benchmark (R² = 0.607) on the same data, and the AIC drops from 2,599 to 2,279. A substantial gain that confirms the smooth terms earn their additional degrees of freedom.

Parametric terms
Room type is the strongest categorical driver. Compared to entire homes, hotel rooms charge roughly 1.9× more, private rooms are about 30% cheaper, and shared rooms drop another step to around 55% below entire homes. The hotel-room result rests on a small number of listings, so we shouldn't over-read it.
Bathrooms show a clean linear effect. Each additional bathroom adds roughly 25% to the price. Besides room type and capacity, this is the strongest structural driver in the model.
Superhost status is the most telling result. After controlling for everything else, the coefficient is essentially zero and not significant. The badge has no measurable effect on the nightly price. This lines up with the GLM Binomial and the Neural Network. Three very different models converge on the same answer.

Smooth terms
All four smooth terms are highly significant (p < 0.001) and every EDF is well above 1, which means a linear approximation would have been wrong for each of them.

- *s(accommodates)* (EDF = 4.58): the curve rises steeply from 1 to about 6 guests and then flattens. Doubling capacity from 8 to 16 adds far less than doubling from 2 to 4, which is exactly the diminishing-returns pattern we expected.
- *s(review_scores_rating)* (EDF = 3.00): the effect is real but modest. The curve trends slightly upward across the range, with a wide confidence band — ratings matter, but they're not a major price lever.
- *s(log1p(number_of_reviews))* (EDF = 2.63): a slight downward trend. Listings with more reviews tend to be priced lower — consistent with the volume-over-margin pattern we keep seeing.
- *s(longitude, latitude)* (EDF = 24.93): by far the most flexible term in the model. The 2D surface shows clear high-price clusters along the Lake Geneva shore and around Lausanne, with sharper drops in rural and alpine areas. This single term is the strongest empirical support for H2 — location is a major, non-linear driver of willingness to pay.

Traffic-light distribution
The traffic-light split is consistent across train and test (Fair ≈ 33%, Underpriced ≈ 41%, Overpriced ≈ 25%), so the model generalises well. The slight overweight of "Underpriced" listings is interesting: it suggests that a notable share of hosts are leaving money on the table compared to what the model considers fair given their listing's features and location.

Take-aways for the pricing tool:
1. Structural features come first. Room type, bathrooms, and capacity together account for most of the explainable price variation that hosts can directly influence. The pricing interface should highlight these.
2. Location effects are real but non-linear. Coordinates as a continuous input clearly outperform postcode dummies — the spatial spline alone justifies the GAM approach for a regional tool.
3. The Superhost badge is an occupancy strategy, not a pricing one. Hosts aiming for the badge should price competitively and let bookings build up.


## Limitations
The GAM is the strongest model in this study, but a few things are worth keeping in mind:
Snapshot only. The model is trained on a single point in time. No seasonality, no weekday effects, no Montreux Jazz Festival, no ski season. Recommendations represent an average market price, not a dynamic one.

Listed price isn't willingness to pay. We use the listed price as a proxy, but it's what hosts charge, not what guests would actually pay. Listings that sit unbooked at their current price are still in the training data, which probably tilts the model towards slightly optimistic price levels.

±15% is a chosen threshold. It's anchored in the residual SD, which makes it defensible, but it's still a choice. Using prediction intervals from the GAM directly, or empirical residual quantiles, would be a more rigorous alternative.

Missing predictors. Amenity counts (pool, parking, Wi-Fi quality), photo quality, view, distance to public transport, noise. None of this is in the dataset. Some of the unexplained variance lives there rather than in the model's flexibility.

Spatial spline describes, doesn't explain. It fits the price surface very well but doesn't tell us why certain locations are expensive. For a pricing tool that's enough; for causal claims (does proximity to the lake cause higher prices, or just correlate with better properties?) it isn't.



# SVM (TIMON - Davide reviewed)

## Theory

## Model

```{r generating Split for SVM (TIMON) - Entfernen sobald OK von Timon da ist}

# Setting a seed ensures reproducibility. 
set.seed(42) 

# Define the relevant columns for the SVM model.
# We include 'price' because the use case requires evaluating if a 
# *given* price for a specific set of features is Overpriced, Fair, or Underpriced.
svm.coloums <- c("traffic_light", "price", "accommodates", "review_scores_rating", 
                 "number_of_reviews", "room_type", "host_is_superhost")

# Creating the 80% Training and 20% Test datasets
d.svm_train <- df.train[, svm.coloums]
d.svm_test  <- df.test[, svm.coloums]

# Verifying the split visually
cat("Training set size:", nrow(d.svm_train), "\n")
cat("Test set size:", nrow(d.svm_test), "\n")

```

```{r tuning model for optimal cost}
# We use 10-fold cross-validation to find the optimal 'cost' and 'gamma' parameters 
# for the radial basis function (RBF) kernel.

tuned_svm <- tune(
  svm,
  traffic_light ~ ., # 'traffic_light' was already converted to a factor in the GAM step
  data = d.svm_train,
  kernel = "radial",
  ranges = list(
    cost = 2^(-6:10),
    gamma = 2^(-5:2)
  )
)

# Extracting the best parameters found during cross-validation
svm.model_cost <- as.numeric(tuned_svm$best.parameters["cost"])
svm.model_gamma <- as.numeric(tuned_svm$best.parameters["gamma"])


#displaying Parameters for visual confirmation
cat("Optimal Cost:", svm.model_cost, "\n")
cat("Optimal Gamma:", svm.model_gamma, "\n")
cat("Best Cross-Validation Performance (Error Rate):", round(tuned_svm$best.performance, 4), "\n")

```

```{r TIMON}
# --- Final SVM Model & Evaluation ---

# Training the final Support Vector Machine using the optimized parameters
svm.main <- svm(traffic_light ~ ., 
                data = d.svm_train, 
                kernel = "radial", 
                cost = svm.model_cost, 
                gamma = svm.model_gamma, 
                scale = TRUE) # Scaling to treat all features equally

# Predicting the traffic light categories on test-data
svm.test_pred <- predict(svm.main, d.svm_test)

# Generating the Confusion Matrix to evaluate classification performance
svm.conf_matrix <- table(Predicted = svm.test_pred, Actual = d.svm_test$traffic_light)
print("Confusion Matrix:")
print(svm.conf_matrix)

# Calculating overall Accuracy
svm_accuracy <- sum(diag(svm.conf_matrix)) / sum(svm.conf_matrix)
cat("\nSVM Test Accuracy:", round(svm_accuracy * 100, 2), "%\n")
```
## Crossvalidation

```{r}
svm.cv1 <- svm(traffic_light ~ ., 
                data = d.svm_train, 
                kernel = "linear", 
                cost = svm.model_cost, 
                gamma = svm.model_gamma, 
                scale = TRUE) # Scaling to treat all features equally

# Predicting the traffic light categories on test-data
svm.test_pred_cv1 <- predict(svm.cv1, d.svm_test)

# Generating the Confusion Matrix to evaluate classification performance
svm.conf_matrix_cv1 <- table(Predicted = svm.test_pred_cv1, Actual = d.svm_test$traffic_light)
print("Confusion Matrix:")
print(svm.conf_matrix_cv1)

# Calculating overall Accuracy
svm_accuracy_cv1 <- sum(diag(svm.conf_matrix_cv1)) / sum(svm.conf_matrix_cv1)
cat("\nSVM Test Accuracy:", round(svm_accuracy_cv1 * 100, 2), "%\n")

```
```{r}
svm.cv2 <- svm(traffic_light ~ ., 
                data = d.svm_train, 
                kernel = "polynomial", 
                cost = svm.model_cost, 
                gamma = svm.model_gamma, 
                scale = TRUE) # Scaling to treat all features equally

# Predicting the traffic light categories on test-data
svm.test_pred_cv2 <- predict(svm.cv2, d.svm_test)

# Generating the Confusion Matrix to evaluate classification performance
svm.conf_matrix_cv2 <- table(Predicted = svm.test_pred_cv2, Actual = d.svm_test$traffic_light)
print("Confusion Matrix:")
print(svm.conf_matrix_cv2)

# Calculating overall Accuracy
svm_accuracy_cv2 <- sum(diag(svm.conf_matrix_cv2)) / sum(svm.conf_matrix_cv2)
cat("\nSVM Test Accuracy:", round(svm_accuracy_cv2 * 100, 2), "%\n")

```


```{r}
svm.cv3 <- svm(traffic_light ~ ., 
                data = d.svm_train, 
                kernel = "radial", 
                cost = 128, 
                gamma = 0.5, 
                scale = TRUE) # Scaling to treat all features equally

# Predicting the traffic light categories on test-data
svm.test_pred_cv3 <- predict(svm.cv3, d.svm_test)

# Generating the Confusion Matrix to evaluate classification performance
svm.conf_matrix_cv3 <- table(Predicted = svm.test_pred_cv3, Actual = d.svm_test$traffic_light)
print("Confusion Matrix:")
print(svm.conf_matrix_cv3)

# Calculating overall Accuracy
svm_accuracy_cv3 <- sum(diag(svm.conf_matrix_cv3)) / sum(svm.conf_matrix_cv3)
cat("\nSVM Test Accuracy:", round(svm_accuracy_cv3 * 100, 2), "%\n")

```


## Interpretion

Support Vector Machine (SVM) Classification To operationalize our
pricing decision-support system, we trained a Support Vector Machine
(SVM) using a radial basis function (RBF) kernel. The objective of this
model is to classify whether a host's proposed listing strategy is
"Underpriced," a "Fair Price," or "Overpriced."

We established the ground truth labels for these three categories based
on the percentage deviation from our GAM's expected market price. The
data was split into an 80% training set and a 20% test set. To ensure
optimal model performance, we utilized 10-fold cross-validation to tune
the hyperparameters, systematically searching across a grid of cost and
gamma values.

By feeding the SVM both the listing characteristics (e.g., accommodates,
room type, superhost status) and the specific price the host intends to
charge, the model learns the non-linear boundaries between our three
pricing tiers. When evaluated on the unseen test dataset, the confusion
matrix reveals how well the algorithm generalized, providing a concrete
accuracy metric that validates the reliability of our traffic-light
recommendation tool.

Using the variables contained in the dataset and some targeted tuning,
we’ve managed to achieve an accuracy of 62%. The model’s main challenge
lies in determining whether a price is fair or not, where it only
performed little better than randomness. This is evident not only in the
“Fair Price” category but also in the other two, where the largest group
of incorrect predictions is found in the “Fair Price” group.

However, the data itself represents a clear shortcoming for the model.
Prediction is virtually impossible without operationalizing the
location. While we have the coordinates, these alone are hardly
meaningful. For example, the distance to city centers or landmarks would
be important to enable better categorization.



# ANN (DAVIDE) Identisch mit Pascal

## Theory


## Model

```{r}
# Selecting basic variables. We omit missing values for this simplified approach.
d.ann_naive <- d.airbnb_listings_feat %>%
  select(
    price,                  
    review_scores_rating,   
    number_of_reviews,      
    host_is_superhost,      
    accommodates            
  ) %>%
  na.omit() 

# Neural Networks perform best (and converge faster) when all input features are scaled to a similar range, typically [0, 1].

normalize <- function(x) {
  return ((x - min(x)) / (max(x) - min(x)))
}

# Apply min-max normalization to all columns
d.ann_naive_scaled <- as.data.frame(lapply(d.ann_naive, normalize))

# Train the basic Neural Network
# We use 1 hidden layer with 3 neurons. linear.output = TRUE is used for regression (predicting a continuous price).
set.seed(42) 
nn.naive <- neuralnet(
  price ~ review_scores_rating + number_of_reviews + host_is_superhost + accommodates,
  data = d.ann_naive_scaled,
  hidden = 3,
  linear.output = TRUE,
  stepmax = 1e6         
)

plot(nn.naive)

# Calculate the mean for our features to represent a "statistically average" Airbnb
mean_rating <- mean(d.ann_naive_scaled$review_scores_rating)
mean_reviews <- mean(d.ann_naive_scaled$number_of_reviews)
mean_acc <- mean(d.ann_naive_scaled$accommodates)

# Create two identical fictional hosts: one is a regular host (0), one is a Superhost (1)
test_data_h3_naive <- data.frame(
  review_scores_rating = c(mean_rating, mean_rating),
  number_of_reviews = c(mean_reviews, mean_reviews),
  host_is_superhost = c(0, 1), # The only difference between the two instances
  accommodates = c(mean_acc, mean_acc)
)

# Predict the scaled prices
predictions_h3_naive <- compute(nn.naive, test_data_h3_naive)

# Reverse the min-max scaling to interpret the results in real CHF
min_price <- min(d.ann_naive$price)
max_price <- max(d.ann_naive$price)
real_prices_naive <- predictions_h3_naive$net.result * (max_price - min_price) + min_price

print("Predicted Prices (Naive Model): Regular Host vs. Superhost")
print(real_prices_naive)
```

```{r}

# Data Preparation & Dimensionality Reduction
d_model_ready <- d.airbnb_listings_feat %>%
  select(
    price,                  
    review_scores_rating,   
    number_of_reviews,      
    host_is_superhost,      
    accommodates,
    property_type,          
    room_type               
  ) %>%
  na.omit() %>%
  # put rare property types into an "Other" category to keep the variable tidy
  mutate(property_type = fct_lump_n(property_type, n = 4, other_level = "Other"))

# One-Hot Encoding: We convert categorical variables into binary dummy variables.
d_dummies <- as.data.frame(model.matrix(~ . - 1, data = d_model_ready))

# Sanitize column names to prevent crashes
colnames(d_dummies) <- make.names(colnames(d_dummies))

d_scaled <- as.data.frame(lapply(d_dummies, normalize))

set.seed(42) 
nn.expanded <- neuralnet(
  price ~ ., 
  data = d_scaled,
  hidden = 3,
  linear.output = TRUE, 
  stepmax = 1e6         
)

plot(nn.expanded)

# To conduct a fair "what-if" analysis (Partial Dependence), we take the mean of ALL variables to represent an absolutely "average" listing.
test_data_h3_exp <- as.data.frame(lapply(d_scaled, mean))

# Duplicate the row to create two identical hypothetical listings
test_data_h3_exp <- rbind(test_data_h3_exp, test_data_h3_exp)

# Isolate the Superhost effect: Row 1 = Regular Host, Row 2 = Superhost
test_data_h3_exp$host_is_superhostFALSE <- c(1, 0) 
test_data_h3_exp$host_is_superhostTRUE <- c(0, 1)

# Predict prices (excluding the 'price' column from the input)
predictions_h3_exp <- compute(nn.expanded, test_data_h3_exp %>% select(-price))

# Reverse scaling using the min/max from the unscaled dataset
min_price_exp <- min(d_model_ready$price)
max_price_exp <- max(d_model_ready$price)

real_prices_expanded <- predictions_h3_exp$net.result * (max_price_exp - min_price_exp) + min_price_exp

print("Predicted Prices (Expanded Model): Regular Host vs. Superhost")
print(real_prices_expanded)
```

## Interpretation ANN

Our initial Neural Network predicted price using only review metrics, listing capacity, and Superhost status. Surprisingly, it priced Superhosts lower (132 CHF) than regular hosts (160 CHF). This was a clear case of omitted variable bias: Superhosts often use competitive pricing to maintain the high occupancy required for the badge, while expensive luxury rentals are usually managed by agencies that don't hold Superhost status.

We corrected this in a second model by adding room and property types, using One-Hot Encoding to convert these categories into numerical inputs. Once we controlled for these structural factors to compare similar listings, the price gap practically vanished (482.64 CHF for Superhosts vs. 483.29 CHF for regular hosts). This refutes our initial hypothesis: when comparing apples to apples, Superhost status alone does not significantly increase a guest's willingness to pay.

## Business Case
The nn_expanded model serves as the analytical core of our pricing decision-support system by accurately predicting a listing's market value based on combined structural and host-centric features. By comparing this to the nn_naive model, we successfully refute our central hypothesis, demonstrating that achieving host_is_superhost status does not command a higher nightly price from guests. Instead, the data reveals that Superhosts often achieve their high status through competitive pricing strategies designed to maximize occupancy rather than nightly margins. Ultimately, a guest's willingness to pay is dictated almost entirely by physical listing attributes like accommodates, room_type, and property_type. Therefore, we advise hosts to anchor their pricing strictly to their property's structural features, leveraging their Superhost badge purely as a marketing tool to capture market share and drive booking volume.

## Limitations

A primary limitation of the nn_expanded model is the inherent "black box" nature of Artificial Neural Networks, which drastically reduces the interpretability of individual price drivers compared to transparent linear models. Because neural networks require mathematical normalization, calculating the exact monetary impact of isolated variables like review_scores_rating is complex and requires continuous reverse-scaling. Furthermore, while we corrected for some omitted variable bias by using One-Hot Encoding for property categories, the model still lacks precise geospatial data, meaning the massive price impact of specific, highly desirable neighborhoods remains uncaptured. Testing our hypothesis also relied on creating a mathematically "average" hypothetical listing, which may not perfectly reflect the nuanced reality of distinct luxury or budget segments in the Canton of Vaud. Finally, applying the na.omit() function removed listings with incomplete profiles, which could introduce bias if those missing data points systematically represented a specific market segment, such as poorly managed properties.


# Results
Wir können sagen, dass für Hosts keine Motivation gibt, 
# Conclusion

# Limitations



# Use of GenAI


# Annex