---
title: "Main Group Work_ML"
author: "Pascal Ingold, Timon Reichmuth und Davide Gualdani"
date: "2026-03-22"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, comment = NA)

```

Next Steps: 


Pascal: 
- a Generalised Linear Model with family set to Poisson --> glm(Price ~ reviews) Achtung Price kann nicht mit Poisson gerechnet werden! Sind keine Zähldaten. Neu Auslastung genutzt, um Zähldaten zu haben. -> DONE
- NA Handling untersuchen -> DONE
- Neural Network --> Ev. auch Kategorisierung analog SVM --> DONE


Timon: 
- a Linear Model --> lm(price ~ rewview_rating + Superhost T/F) --> für unterscheidung und kontrolle. -> INPROGRESS


- NA Handling untersuchen
  |-> NA vorwiegend bei ratings und rates vor. Verdacht: wenn nie eine Vermietung gemacht wurde erhalten wir einen NA in den Ratings
  TODO: schauen ob wenn wir mit den Variablen rechnen, NA's einen negativen Einfluss haben. Wie zählen NA in ein Modell hinein?
        --> beim lm keinen Einfluss, da diese Zeilen einfach weggelassen werden. 
        --> bei anderen Modellen muss aufgepasst werden. 
        
        
– a Support Vector Machine --> Overpriced & Underpriced & Right Price (Davide und Timon)


Davide:
- – a Generalised Linear Model with family set to Binomial --> glm(Price ~ Superhost T/F) -> DONE
– a Generalised Additive Model --> gam.main <- gam(price ~  --> allenfalls aufbauend auf dem lm()
                                    s(accommodates) +
                                    s(review_scores_rating) +
                                    s(number_of_reviews) +
                                    s(longitude, latitude) +
                                    room_type +
                                    host_is_superhost +
                                    bathrooms,
                                  family = gaussian(link = "log"),
                                  data   = d.airbnb_listings_feat_gam)


Infos zu bedenken: 
ABGABEDATUM: 12 Juni

- Outlier Handling definieren.
--> falsche Datenpunkte sollten gelöscht werden. Extremwerte im Zweifel drin lassen. Ansonsten kann das als p-hacking verstanden werden. 

Coding Conventions. 
• data sets are named with the prefix “d.” (e.g., d.cats).
• linear model objects are named with the prefix “lm.” (e.g., lm.cats)
• within the text R-objects are written using code formatting (e.g., lm.cats)
14
• variable names are written using italic (e.g., Body.weight)
• these conventions are arbitrary, but hopefully useful to the reader


# Research Questions

What factors, which can be modified by Hosts on AirBnb, affect guest willingness to pay more for a certain room?

What are factors that influence the guest willingness to pay (WTP)

What are factors on which the guest can adjust their pricing model?

## Concepts: Willingness to pay (WTP)

WTP is the hypothetical value, how much a customer is willing to pay for a location or a room for a certain standard. This can rise, if for example the host provides the ability to wash clothes in the apartment. So for the same location, once with and once without a washing machine, the customers WTP might increase with the washer. In a more abstract example, if the host is a super host, the customer might trust the host more and his WTP is higher. 

## Hypothesis
H1: If the review-score of a host is better, the WTP of a guest increases. 
H2: If a host has more reviews, the WTP of a guest increases. 
H3: If a host is e superhost, the WTP of a guest increases. 

WTP is measured in the Price displayed. We assume, that the mean of prices is similar to the WTP and residuals even out. Therefore we can work with the effective prices displayed on AirBnb. 



# Business Case
We propose a data-driven pricing decision-support system for Airbnb hosts in the
Canton of Vaud, providing market-aligned price recommendations based on listing
characteristics. The tool targets both new and existing hosts, using inputs such
as location, room type, guest capacity, availability, ratings, and superhost status.
Machine learning models, including LM, GLM, GAM, SVM, and Neural
Networks, will be applied to identify key price drivers and generate accurate
predictions. The objective is to balance revenue and occupancy, helping hosts
understand their market value.
A key feature is interpretability, enabling users to see why a certain price is
recommended. From a business perspective, the system aims to increase booking
rates, reduce price dispersion, and improve host satisfaction.
The model is tailored to the Canton of Vaud to capture regional economic effects.


# Editing of Dataset
First step: Loading Packages

```{r Libraries, results = "hide"}
library(tidyverse)
library(janitor)
library(mgcv)
library(e1071)
library(nnet)
library(pROC)
library(Metrics)
library(performance)
library(broom)
library(stringr)
library(ggplot2)
library(patchwork)
library(knitr)
library(performance)
library(naniar)
library(visdat)
library(neuralnet)
library(leaflet)
library(leaflet.extras)

```



Second step: Reading all available Datasets
```{r reading Dataset}
path_listings <- "./Data_Vaud_all/listings-2.csv"
d.raw_data_listings <- read.csv(path_listings, header = TRUE)
```

Cleaning the Airbnb-Listings Data. We keep only the variables needed for our project. We redouce the amount of variables to exact 20, according to the requirements by our lecturers.

```{r cleaning Dataset}
d.airbnb_listings <- d.raw_data_listings %>%
  clean_names()

d.listings_clean <- d.airbnb_listings %>%
  mutate(
    price = str_remove_all(price, "\\$|,"),
    price = as.numeric(price),
    host_is_superhost = ifelse(host_is_superhost %in% c("t", "TRUE", "1"), 1, 0),
    room_type = as.factor(room_type),
    property_type = as.factor(property_type)
  ) %>%
  filter(!is.na(price), price > 0)

d.listings_clean <- d.listings_clean %>%
  select(-id, -host_id, -host_listings_count, -host_total_listings_count, -listing_url, -scrape_id, -last_scraped,
         -source, -name, -description, -maximum_nights, -first_review, -last_review, -calculated_host_listings_count, -host_since, -availability_365, -estimated_revenue_l365d, 
         -neighborhood_overview, -picture_url, -host_url, -host_name, -host_location,
         -host_about, -host_thumbnail_url, -host_picture_url, -host_neighbourhood,
         -neighbourhood, -neighbourhood_cleansed, -bathrooms_text, -amenities, 
         -minimum_minimum_nights, -maximum_minimum_nights, -minimum_maximum_nights,
         -maximum_maximum_nights, -minimum_nights_avg_ntm, -maximum_nights_avg_ntm,
         -calendar_updated, -has_availability, -availability_30, -availability_60,
         -availability_90, -calendar_last_scraped, -number_of_reviews_ltm, -number_of_reviews_l30d,
         -availability_eoy, -number_of_reviews_ly, -review_scores_accuracy, -review_scores_cleanliness,
         -review_scores_checkin, -review_scores_communication, -review_scores_location,
         -license, -instant_bookable, -calculated_host_listings_count_entire_homes, -calculated_host_listings_count_private_rooms,
         -calculated_host_listings_count_shared_rooms, -reviews_per_month, -bedrooms, -review_scores_value)
```

This is our final, cleaned Dataset.
```{r final Call cleaned Dataset}
str(d.listings_clean)
```


# Feature Engineering

First Feature Engineering for the airbnb_listings data

```{r FE for XXX (XXX)}

#Für Model XXXX 
d.airbnb_listings_feat <- d.listings_clean %>%
  mutate(
    log_price = log(price),
    price_per_guest = price / pmax(accommodates, 1),
    high_demand = ifelse(estimated_occupancy_l365d > 200, 1, 0)
  )
```


```{r FE for lm (TIMON)}

#removing % from response rate and make it binomial
d.airbnb_listings_feat$host_response_rate <- as.numeric(sub("%", "", d.airbnb_listings_feat$host_response_rate)) / 100 

#removing % from response rate and make it binomial
d.airbnb_listings_feat$host_acceptance_rate <- as.numeric(sub("%", "", d.airbnb_listings_feat$host_acceptance_rate)) / 100 

#change the values of host_is_superhost to TRUE/False
d.airbnb_listings_feat$host_is_superhost <- ifelse(d.airbnb_listings_feat$host_is_superhost==1,
                                                   TRUE, FALSE
                                                   )


```


```{r FE for SVM (TIMON)}
d.airbnb_listings_feat$estimated_revenue_365 <- d.airbnb_listings_feat$price * d.airbnb_listings_feat$estimated_occupancy_l365d

summary(d.airbnb_listings_feat$estimated_revenue_365)
hist(d.airbnb_listings_feat$estimated_revenue_365)
print(quantile(d.airbnb_listings_feat$estimated_revenue_365))


```
wie gehen wir bei diesen Variablen mit den NA um?

# Data Quality

Wir prüfen ob und wieviele NA's vorhanden sind. Hierzu nutzen wir naniar und visdat, um die NA's auch visualisieren zu können. Die meisten NA-Werte sind bei den Rates, daher wir von Pascal angenommen, dass die Nullen nicht importiert wurden und deshalb es zu dieser grossen Anzahl NAs kam. Ich würde hier die NA durch 0 ersetzen.
Entscheid: Da Ratings fehlen werden sie in Modellen omitted. Daten fehlen aus einem Grund und nicht random. 

```{r}
vis_miss(d.airbnb_listings_feat)
gg_miss_var(d.airbnb_listings_feat)
gg_miss_upset(d.airbnb_listings_feat)
```


#GLM set to binomial(Davide)
1. Schritt: Zielvariable binarisieren
Aus Preis eine 0/1-Variable machen
Der Median ist der sauberste Cutpoint — er ist robust gegenüber Ausreissern und teilt den Datensatz exakt in zwei Hälften
```{r FE for glm binomial (DAVIDE)}
d.airbnb_listings_feat <- d.airbnb_listings_feat %>%
  mutate(price_high = as.integer(price > median(price, na.rm = TRUE)))

# Checken ob Split sauber ist
table(d.airbnb_listings_feat$price_high)

```


```{r FE for glm poisson (PASCAL)}


```




```{r final call of FE-dataset}
str(d.airbnb_listings_feat)
```


# Exploratory Data Analysis

```{r}
hist(d.airbnb_listings_feat$price)
hist(d.airbnb_listings_feat$log_price)
ggplot(data= d.airbnb_listings_feat, aes(x = room_type, y = price)) +
  geom_boxplot() +
  scale_y_log10() +
  labs(
    title = "Preisverteilung nach Room Type (log-Skala)",
    x = "Room Type",
    y = "Log_Preis"
  ) +
  theme_minimal()

ggplot(data = d.airbnb_listings_feat, aes(x = accommodates, y = log_price)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "loess", se = FALSE) +
  labs(
    title = "Preis vs. Kapazität (log-Skala)",
    x = "Anzahl Gäste",
    y = "Log_Preis"
  ) +
  theme_minimal()
  
ggplot(data = d.airbnb_listings_feat, aes(x = log_price, y = number_of_reviews)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "loess", se = FALSE) +
  labs(
    title = "Anzahl Reviews vs. Preis",
    x = "Log_Preis",
    y = "Anzahl Reviews"
  ) +
  theme_minimal()

ggplot(data = d.airbnb_listings_feat, aes(x = longitude, y = latitude)) +
  geom_point(aes(color = log_price), alpha = 0.6) +
  scale_color_viridis_c() +
  labs(
    title = "Airbnb Listings (Farbe = Log-Preis)",
    x = "Longitude",
    y = "Latitude",
    color = "Log_Preis"
  ) +
  theme_minimal()

ggplot(data = d.airbnb_listings_feat, aes(x = longitude, y = latitude)) +
  geom_point(aes(color = log_price, size = number_of_reviews), alpha = 0.5) +
  scale_color_viridis_c() +
  labs(
    title = "Listings: Preis und Nachfrage",
    x = "Longitude",
    y = "Latitude",
    color = "Log_Preis",
    size = "Reviews"
  ) +
  theme_minimal()
```

*Important to mention*
During our exploratory data analysis, we identified one 'Private room' listing with an unrealistically high price of 2'147'483'647 CHF. We excluded this single observation from our dataset as it is an extreme outlier that does not reflect the true market conditions and would negatively skew our predictive models.
```{r}
# This keeps everything EXCEPT the row that is a Private room AND costs more than 5000
d.airbnb_listings_feat <- d.airbnb_listings_feat %>%
  filter(!(room_type == "Private room" & price > 100000))
```


# Davide's EDA
```{r}
# --- 1. Distribution of the Target Variable (Price) ---
# Upgrading base R hist() to ggplot2 for a more professional look.
# Showing both raw and log side-by-side helps visually justify the transformation to the client.
p_raw <- ggplot(d.airbnb_listings_feat, aes(x = price)) +
  geom_histogram(bins = 50, fill = "steelblue", color = "white") +
  labs(title = "Raw Price Distribution", x = "Price", y = "Count") +
  theme_minimal()

p_log <- ggplot(d.airbnb_listings_feat, aes(x = log_price)) +
  geom_histogram(bins = 50, fill = "darkgreen", color = "white") +
  labs(title = "Log-transformed Price Distribution", x = "Log(Price)", y = "Count") +
  theme_minimal()
p_raw + p_log # Displays them side-by-side

# --- 2. Categorical Predictor: Room Type vs. Price ---
# Your original code was good, just cleaned up the aesthetics.
ggplot(data = d.airbnb_listings_feat, aes(x = room_type, y = log_price, fill = room_type)) +
  geom_boxplot(alpha = 0.7) +
  labs(
    title = "Price Distribution by Room Type",
    x = "Room Type",
    y = "Log(Price)"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

# --- 3. Continuous Predictor: Capacity vs. Price ---
ggplot(data = d.airbnb_listings_feat, aes(x = accommodates, y = log_price)) +
  geom_point(alpha = 0.3, color = "darkgray") +
  geom_smooth(method = "loess", color = "blue", se = FALSE) +
  labs(
    title = "Price vs. Capacity",
    subtitle = "Smoother applied to identify potential non-linear trends",
    x = "Number of Guests (Accommodates)",
    y = "Log(Price)"
  ) +
  theme_minimal()

# --- 4. Continuous Predictor: Reviews vs. Price ---
# Correction: The response variable (log_price) should generally be on the y-axis 
# when plotting against a predictor[cite: 255].
ggplot(data = d.airbnb_listings_feat, aes(x = number_of_reviews, y = log_price)) +
  geom_point(alpha = 0.3, color = "darkgray") +
  geom_smooth(method = "loess", color = "red", se = FALSE) +
  labs(
    title = "Does the Number of Reviews impact Price?",
    x = "Total Number of Reviews",
    y = "Log(Price)"
  ) +
  theme_minimal()

# --- 5. Binary Predictor: Superhost Status ---
# Adding this because it's a key part of your business case and fulfills the binary variable requirement.
ggplot(data = d.airbnb_listings_feat, aes(x = as.factor(host_is_superhost), y = log_price, fill = as.factor(host_is_superhost))) +
  geom_boxplot(alpha = 0.7) +
  scale_x_discrete(labels = c("0" = "Regular Host", "1" = "Superhost")) +
  labs(
    title = "Price Difference: Superhosts vs. Regular Hosts",
    x = "Host Status",
    y = "Log(Price)"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

# --- 6. Spatial Data: Mapping the Listings ---
# Cleaned up the map to focus purely on location and price.
ggplot(data = d.airbnb_listings_feat, aes(x = longitude, y = latitude)) +
  geom_point(aes(color = log_price), alpha = 0.6, size = 1.5) +
  scale_color_viridis_c(option = "plasma") +
  labs(
    title = "Airbnb Listings in Vaud by Price",
    x = "Longitude",
    y = "Latitude",
    color = "Log(Price)"
  ) +
  theme_minimal()
```
# Davide's overall Data Analysis
## 1. Overall Market Overview (Prices & Capacity)
This is the absolute minimum, maximum, median, and mean for the most important numeric variables.
```{r}
# Basic summary statistics for key numerical variables
summary_stats <- d.airbnb_listings_feat %>%
  select(price, log_price, accommodates, number_of_reviews) %>%
  summary()

print(summary_stats)
```
Client Interpretation: "The average Airbnb in Vaud costs X CHF per night, but the median is Y CHF, indicating that a few extremely expensive luxury listings might be skewing the average upward. Prices range from a minimum of A CHF to a maximum of B CHF."

## 2. Analysis by Room Type
Since we want to know how many different types of rooms there are, we group the data and calculate the counts, percentages, and price ranges for each type.
```{r}
# Detailed breakdown by Room Type
room_type_analysis <- d.airbnb_listings_feat %>%
  group_by(room_type) %>%
  summarise(
    Total_Listings = n(),
    Percentage_of_Market = round((n() / nrow(d.airbnb_listings_feat)) * 100, 1),
    Min_Price = min(price, na.rm = TRUE),
    Median_Price = median(price, na.rm = TRUE),
    Average_Price = round(mean(price, na.rm = TRUE), 1),
    Max_Price = max(price, na.rm = TRUE)
  ) %>%
  arrange(desc(Total_Listings)) # Sorts by the most common room type

print(room_type_analysis)
```
Client Interpretation: "Entire homes/apartments dominate the Vaud market, making up 75% of all listings. As expected, the Hotel room command the highest median price (264.5 CHF), compared to shared rooms (47 CHF)."

## 3. Superhost vs. Regular Host Breakdown
This business case specifically mentions superhost status. So we took a look to see how many there are and if they actually charge more or even less.
```{r}
# Superhost statistics
superhost_analysis <- d.airbnb_listings_feat %>%
  group_by(host_is_superhost) %>%
  summarise(
    Total_Hosts = n(),
    Percentage = round((n() / nrow(d.airbnb_listings_feat)) * 100, 1),
    Median_Price = median(price, na.rm = TRUE),
    Avg_Reviews = round(mean(number_of_reviews, na.rm = TRUE), 1)
  ) %>%
  mutate(host_is_superhost = ifelse(host_is_superhost == 1, "Yes", "No"))

print(superhost_analysis)
```


## 1. Correlation Matrix (Crucial before Machine Learning)
Before we train our LM, GAM, or SVM, we need to know how our numerical variables relate to each other. This helps us identify multicollinearity (when two predictors provide the same information) and strong price drivers.

```{r}
# Select relevant numerical variables (drop NAs just for the correlation calculation)
cor_data <- d.airbnb_listings_feat %>%
  select(log_price, accommodates, number_of_reviews) %>%
  drop_na()

# Calculate the correlation matrix
cor_matrix <- cor(cor_data)

# Render a nice table for your RMarkdown document
kable(round(cor_matrix, 2), caption = "Correlation Matrix of Key Numerical Drivers")
```
Client Interpretation: "As expected, accommodates shows the strongest positive correlation with price. It is also interesting to note the relationship between availability and number_of_reviews..."

## 2. Detailed property_type Analysis
We already analyzed room_type, but Airbnb often has dozens of property_types (e.g., "Entire chalet", "Private room in castle", etc.). Summarizing the Top 10 property types in the Canton of Vaud provides maybe a great market insight.
```{r}
# Top 10 Property Types by frequency and their median price
top_properties <- d.airbnb_listings_feat %>%
  group_by(property_type) %>%
  summarise(
    Total_Listings = n(),
    Median_Price = median(price, na.rm = TRUE)
  ) %>%
  arrange(desc(Total_Listings)) %>%
  slice_head(n = 10) # Keep only the top 10

kable(top_properties, caption = "Top 10 Property Types in the Canton of Vaud")
```

Client Interpretation: "While standard apartments like "entire rental unit" are the most common, specialty properties like chalets, homes and villas command a significant premium, which our model will need to capture."

## 3. Missing Values (NA) Check
This step is less for the client and essential for your ML models. Most modeling functions (like lm() or svm()) will throw errors or silently drop entire rows if NAs are present. You need to know exactly where your data is missing!

```{r}
# Count the number of NAs for each key column in your dataset
na_summary <- d.airbnb_listings_feat %>%
  select(price, log_price, accommodates, number_of_reviews, host_is_superhost) %>%
  summarise(across(everything(), ~ sum(is.na(.)))) %>%
  t() %>% # Transpose for better readability
  as.data.frame() %>%
  rename(Missing_Values = V1)

kable(na_summary, caption = "Count of Missing Values per Variable")
```



#LM (Timon)


```{r lm controlled by Timon}
lm.main <- lm(log(price) ~ review_scores_rating + as.factor(host_is_superhost) + as.factor(room_type) + accommodates, data = d.airbnb_listings_feat)
summary(lm.main)

multicollinearity(lm.main) #kann werte von 1 - unendlich haben. 1 ist perfekt nicht kollinear

```

```{r lm controlled plot}

ggplot(data = d.airbnb_listings_feat, mapping = aes(review_scores_rating, log(price), color = host_is_superhost))+
  geom_point()+
  geom_smooth(method = "lm")




d.lm_predict <- expand.grid(
  review_scores_rating = seq(3, 5, length.out = 100),
  host_is_superhost = c("TRUE", "FALSE"),
  room_type = "Entire home/apt",   # fixiert
  accommodates = mean(d.airbnb_listings_feat$accommodates, na.rm = TRUE)
)

d.lm_predict$pred <- predict(lm.main, newdata = d.lm_predict)


ggplot(d.airbnb_listings_feat, 
       aes(review_scores_rating, log(price), color = host_is_superhost)) +
  geom_point(alpha = 0.3) +
  geom_line(data = d.lm_predict, aes(y = pred), size = 1.2)
```



# GLM set to poisson (Pascal)
```{r, Pascals GLM}
# Poisson-Modell zum Auslastung vorherzusagen

glm_poisson <- glm(estimated_occupancy_l365d ~ price + host_is_superhost + 
                     review_scores_rating + accommodates + room_type, 
                   data = d.airbnb_listings_feat, 
                   family = "poisson")

summary(glm_poisson)

# Gibt es Overdispersion? Poisson-Annahme ist, dass Mittelwert exakt gleich der Varianz ist. Airbnbs sind viele günstig und wenige sehr teuer!
dispersion_ratio <- glm_poisson$deviance / glm_poisson$df.residual
print(paste("Dispersion Ratio:", dispersion_ratio))

glm_quasipoisson <- glm(estimated_occupancy_l365d ~ price + host_is_superhost + 
                          review_scores_rating + accommodates + room_type, 
                        data = d.airbnb_listings_feat, 
                        family = "quasipoisson")

summary(glm_quasipoisson)
```
## Interpretation 
### GLM-Poisson:

Das GLM mit Poissonverteilung hat wie erwartet eine Überdispersion produziert, da die Streuung der Airbnb-Daten zu erwarten war. Das Modell hat die Standardfehler zu klein berechnet, wweshalb im ersten Output alle Variablen hochsignifikant waren. Danach wurde die Dispersion berechnet und sie war grösser als 1, nämlich 56.53. Als Lösung wurde ein Quasipoissonmodell genutzt, welches die Standardfehler nach oben anpasst.

### GLM-Quasipoisson
Das GLM mit Quasipoissonverteilung hat die Überdispersion korrigiert. Die Standardfehler wurden nach oben angepasst (deutlich) und gibt nun ein realistisches Bild, welche Faktoren wirklich einen Einfluss auf die Auslastung haben.
Interpretationen:
- Price -> hochsignifikant negativ. Wie erwartet: Je höher der Preis, desto geringer die Auslastung
- Superhost: hochsignifikant positiv: Achtung, logarithmierte Raten, darum Effekt mit Exponentialfunktion berechnen exp(0.9577) = 2.6. Ein Superhost zu sein steigert die Auslastung um das 2,6-fache im Vergleich zum normalen Host.
- Bewertungen: Positiv und signifikant. Bessere Bewertungen führen zu mehr Buchungen.
- Kapazität: Positiv und signifikant. Je mehr Leute reinpassen, desto mehr Buchungen.
- Zimmertypen: Im Intercept ist "Entire home/apt". Private Room ist signifikant schlechter gebucht als ganze Unterkunft. Shared Room hat hohen P-Wert. Statistisch kann man nicht sagen, ob sich die Auslastung eines geteilten Zimmers von der ganzen Wochnung unterscheidet.

## GLM Quasi-Poisson Ergebnisse für Businesscase

```{r, Pascal}

# Extract model results and calculate the percentage change for each variable
d.results_glm_quasipoisson <- tidy(glm_quasipoisson) %>%
  # Calculate the percentage
  mutate(
    pct_change = (exp(estimate) - 1) * 100,
    # Round the percentage change to 2 
    pct_change = round(pct_change, 2)
  ) %>%
  # Select only the relevant columns for business case
  select(term, estimate, pct_change, p.value)

print(d.results_glm_quasipoisson)

```



# GLM Binomial (DAVIDE)

```{r}
glm_binomial <- glm(price_high ~ host_is_superhost,
                    data   = d.airbnb_listings_feat,
                    family = binomial(link = "logit"))

summary(glm_binomial)
```
# Interpretation GLM Binomial - Odds Ratio berechnen (DAVIDE)
Genau wie Pascal exp() für seinen Quasipoisson gebraucht hat, brauchst du das hier auch. Bei Binomial sind es Odds Ratios:

```{r}
d.results_glm_binomial <- tidy(glm_binomial) %>%
  mutate(
    odds_ratio = round(exp(estimate), 3),
    p.value    = round(p.value, 4)
  ) %>%
  select(term, estimate, odds_ratio, p.value)

print(d.results_glm_binomial)
```
Was der Output sagt
Der Koeffizient für host_is_superhost ist negativ (-0.2217) und signifikant (p = 0.002). Das Odds Ratio von 0.801 bedeutet: Superhosts haben ~20% geringere Odds, über dem Medianpreis zu liegen — *somit sind Superhosts tendenziell günstiger, nicht teurer.*
Das klingt zuerst wie ein Widerspruch zu H3. Ist es aber nicht.
Die eigentlich spannende Erkenntnis: kombiniere das mit Pascals Ergebnis

Pascals Quasipoisson hat gezeigt: 
Superhost-Status steigert die Auslastung um Faktor 2.6. Dein Binomial zeigt: Superhosts liegen eher unter dem Medianpreis.
Das ergibt zusammen ein kohärentes Bild der Superhost-Strategie:

Superhost = niedrigerer Preis  +  viel höhere Auslastung
Normal Host = höherer Preis    +  deutlich weniger Buchungen

*Superhosts optimieren auf Volumen statt Marge.* Das macht auch betriebswirtschaftlich Sinn: um den Superhost-Status zu halten, braucht man viele Buchungen und gute Reviews — das funktioniert besser mit kompetitiven Preisen.

H3 wird nicht bestätigt im Sinne von "Superhosts können mehr verlangen".
Stattdessen zeigt sich das Gegenteil: *Superhosts preisen strategisch*
günstiger, erzielen dafür aber massiv mehr Buchungen (Pascal's Modell).
Das ist ein stärkerer Befund als eine simple Bestätigung von H3 — es zeigt eine echte Preisstrategie.

Ein Host der Superhost-Status anstrebt, sollte seinen Preis nicht erhöhen, sondern durch kompetitive Preise Buchungen und Reviews sammeln. Der Superhost-Status zahlt sich dann indirekt über Auslastung aus, nicht über den Einzelpreis.


# GAM Modell (DAVIDE)

```{r}

# Select only the necessary variables and remove NAs
# GAM cannot handle NA's in df → na.omit() neccessary
d.airbnb_listings_feat_gam <- d.airbnb_listings_feat %>%
  select(price, accommodates, review_scores_rating,
         number_of_reviews, longitude, latitude,
         room_type, host_is_superhost, bathrooms) %>%
  na.omit()

cat("Observations after na.omit():", nrow(d.airbnb_listings_feat_gam), "\n")
```

```{r}
# GAM Modell
gam.main <- gam(price ~
                  s(accommodates) +
                  s(review_scores_rating) +
                  s(number_of_reviews) +
                  s(longitude, latitude) +
                  room_type +
                  host_is_superhost +
                  bathrooms,
                family = gaussian(link = "log"),
                data   = d.airbnb_listings_feat_gam)

summary(gam.main)
```

Model Overview
The Generalised Additive Model (GAM) was fitted with a Gaussian family and log link, using 3,160 observations after listwise deletion of missing values. The model explains 52.9% of deviance (adj. R² = 0.521), which is considered strong performance for hedonic price modelling in the short-term rental market.

Parametric Terms (Linear Effects)
Room Type is the strongest categorical price driver in the model. Using Entire home/apt as the reference category:

Hotel room: significant positive premium (β = 0.720, p < 0.001). Translating from the log scale: exp(0.720) ≈ 2.05, meaning hotel rooms are associated with roughly double the price of comparable entire homes.
Private room: significant negative effect (β = −0.363, p < 0.001), corresponding to approximately a 30% price reduction (exp(−0.363) = 0.696).
Shared room: not statistically significant (p = 0.180), likely due to the small number of shared room listings in the Canton of Vaud.

Bathrooms show a highly significant positive effect (β = 0.181, p < 0.001). Each additional bathroom is associated with approximately a 20% price increase (exp(0.181) = 1.198), reflecting the importance of physical listing characteristics for guest willingness to pay.
Superhost status shows a small negative coefficient (β = −0.049, p = 0.078), which is only marginally significant at the 10% level. This finding is consistent across all three models estimated in this study — the GLM Binomial, the Neural Network, and the GAM all converge on the same conclusion: superhost status does not yield a measurable price premium. This is discussed in detail in the cross-model comparison below.

Smooth Terms (Non-linear Effects)
All four smooth terms are highly significant (p < 2e-16) and show substantial non-linearity, as indicated by their estimated degrees of freedom (EDF)

An EDF close to 1 would indicate a near-linear relationship, making a spline unnecessary. All EDF values here are far above 1, confirming that non-linear smoothing is statistically justified for every continuous predictor in the model. A standard linear model would have failed to capture these relationships adequately.
The *2D spatial spline* s(longitude, latitude) deserves particular attention with an EDF of 27.75. This indicates a highly complex spatial price surface across the Canton of Vaud — locations closer to Lake Geneva and urban centres such as Lausanne command significantly higher prices, while rural and alpine areas show distinct pricing patterns. This spatial component alone justifies the GAM approach for a regionally scoped pricing tool.

Cross-Model Comparison: Superhost Effect
One of the most consistent findings across all models in this study is the absence of a superhost price premium
This convergence across fundamentally different model architectures — parametric, non-parametric, and connectionist — constitutes strong evidence against H3 as originally formulated. Superhosts do not appear to charge higher prices; rather, as Pascal's Quasi-Poisson model demonstrates, they achieve 2.6× higher occupancy through competitive pricing. The data therefore suggest a volume-over-margin strategy among superhost-status hosts in Canton Vaud.




# GAM Visualisation
```{r}
# Partial effect plots—this is the main advantage of GAM over LM
par(mfrow = c(2, 3))
plot(gam.main, 
     shade      = TRUE, 
     col        = "steelblue", 
     shade.col  = "lightblue",
     seWithMean = TRUE)

# EDF table: shows whether splines are truly nonlinear
# edf ≈ 1 → nearly linear (spline hardly necessary)
# edf > 2 → clearly nonlinear (spline justified)
edf_table <- as.data.frame(summary(gam.main)$s.table) %>%
  round(3)
kable(edf_table, caption = "Smooth Terms: Estimated Degrees of Freedom (EDF)")

# Model quality
cat("R-squared:         ", round(summary(gam.main)$r.sq, 3), "\n")
cat("Deviance explained:", round(summary(gam.main)$dev.expl, 3), "\n")

# Comparison with Timon's LM (AIC — lower = better) -> unfair comparison
AIC(lm.main, gam.main)

# LM with identical Variables as in GAM (fair comparison)
lm.gam_comparison <- lm(log(price) ~ accommodates + review_scores_rating +
                           number_of_reviews + longitude + latitude +
                           room_type + host_is_superhost + bathrooms,
                         data = d.airbnb_listings_feat_gam)

cat("LM  R-squared:", summary(lm.gam_comparison)$r.squared, "\n")
cat("GAM R-squared: 0.521\n")
AIC(lm.gam_comparison, gam.main)
```
Business Case Implications
The GAM results provide three actionable insights for the pricing decision-support tool:
*Firstly*, physical listing characteristics (room type, bathroom count, guest capacity) are the primary price drivers a host can directly control. The model recommends prioritising these structural factors in the pricing interface.
*Secondly*, location matters non-linearly. The spatial spline captures granular geographic price gradients that a postcode or neighbourhood dummy variable would miss. This validates the use of coordinates as a continuous input in the recommendation tool.
*Thirdly*, pursuing superhost status should be framed as an occupancy strategy, not a pricing strategy. Hosts seeking the superhost badge are advised to price competitively to accumulate bookings and reviews, with the understanding that revenue gains come from higher occupancy rather than a higher nightly rate.


#Interactive Map of Canton Vaude with the GAM
```{r}
# ── Echte Listings mit GAM-Fitted Values ──

d.airbnb_listings_feat_gam$fitted_price <- fitted(gam.main)

# Farbpalette
pal_pts <- colorNumeric(
  palette = "viridis",
  domain  = d.airbnb_listings_feat_gam$fitted_price
)

# ── KARTE ──
leaflet() %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  setView(lng = 6.65, lat = 46.52, zoom = 9) %>%
  addCircleMarkers(
    data        = d.airbnb_listings_feat_gam,
    lng         = ~longitude,
    lat         = ~latitude,
    radius      = 4,
    color       = ~pal_pts(fitted_price),
    fillOpacity = 0.8,
    stroke      = FALSE,
    popup       = ~paste0(
      "<b>Actual Price:</b> CHF ", round(price), "<br>",
      "<b>GAM Fitted:</b> CHF ", round(fitted_price), "<br>",
      "<b>Room Type:</b> ", room_type, "<br>",
      "<b>Superhost:</b> ", host_is_superhost, "<br>",
      "<b>Accommodates:</b> ", accommodates
    )
  ) %>%
  addLegend(
    position = "bottomright",
    pal      = pal_pts,
    values   = d.airbnb_listings_feat_gam$fitted_price,
    title    = "Fitted price<br>per listing (CHF)"
  )
```


```{r}
# ── Echte Listings mit Ampelsystem (%-Abweichung vom GAM) ──

d.airbnb_listings_feat_gam$fitted_price <- fitted(gam.main)

# Prozentuale Abweichung: (Actual - Fitted) / Fitted * 100
d.airbnb_listings_feat_gam$pct_diff <- 
  (d.airbnb_listings_feat_gam$price - d.airbnb_listings_feat_gam$fitted_price) / 
  d.airbnb_listings_feat_gam$fitted_price * 100

# Ampel-Kategorien
d.airbnb_listings_feat_gam$traffic_light <- case_when(
  d.airbnb_listings_feat_gam$pct_diff <= -10 ~ "Good Deal (10%+ underpriced)",
  d.airbnb_listings_feat_gam$pct_diff >=  10 ~ "Overpriced (10%+ overpriced)",
  TRUE                                        ~ "Fair Price (within ±10%)"
)

# Farbzuweisung
color_map <- c(
  "Good Deal (10%+ underpriced)" = "#1D9E75",  # grün
  "Fair Price (within ±10%)"     = "#EF9F27",  # gelb/amber
  "Overpriced (10%+ overpriced)" = "#D85A30"   # rot
)

d.airbnb_listings_feat_gam$marker_color <- 
  color_map[d.airbnb_listings_feat_gam$traffic_light]

# ── KARTE ──
leaflet() %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  setView(lng = 6.65, lat = 46.52, zoom = 9) %>%
  addCircleMarkers(
    data        = d.airbnb_listings_feat_gam,
    lng         = ~longitude,
    lat         = ~latitude,
    radius      = 4,
    fillColor   = ~marker_color,
    fillOpacity = 0.85,
    stroke      = FALSE,
    popup       = ~paste0(
      "<b>Verdict:</b> ", traffic_light, "<br>",
      "<b>Actual Price:</b> CHF ", round(price), "<br>",
      "<b>GAM Fitted:</b> CHF ", round(fitted_price), "<br>",
      "<b>Difference:</b> ", round(pct_diff, 1), "%", "<br>",
      "<hr>",
      "<b>Room Type:</b> ", room_type, "<br>",
      "<b>Superhost:</b> ", host_is_superhost, "<br>",
      "<b>Accommodates:</b> ", accommodates
    )
  ) %>%
  addLegend(
    position = "bottomright",
    colors   = c("#1D9E75", "#EF9F27", "#D85A30"),
    labels   = c("Good Deal (10%+ underpriced)",
                 "Fair Price (within ±10%)",
                 "Overpriced (10%+ overpriced)"),
    title    = "Price vs. GAM Estimate",
    opacity  = 0.85
  )
```


# SVM


```{r}
s <- sample(3160,2528) #80% of data is in the test-dataset
coloums <- c("host_is_superhost",
             "review_scores_rating", "host_acceptance_rate")
d.airbnb_listings_feat_train <- d.airbnb_listings_feat[s, coloums]
d.airbnb_listings_feat_test <- d.airbnb_listings_feat[-s, coloums]


print("traindata:") 
length(d.airbnb_listings_feat_train$host_is_superhost)
print("testdata:")  
length(d.airbnb_listings_feat_test$host_is_superhost)
```

```{r}
svm.tune_price <- tune(svm, as.factor(host_is_superhost) ~ ., data = d.airbnb_listings_feat_train, kernel = "radial", ranges = list(cost = c(0.001, 0.01, 0.1, 1, 10, 100)))
summary(svm.tune_price)

```

```{r}
svm.model_price <- svm(as.factor(host_is_superhost) ~ ., data = d.airbnb_listings_feat_train, kernel = "radial", cost = 0.1, scale = TRUE)

print(svm.model_price)
plot(svm.model_price, data = d.airbnb_listings_feat_test)

svm.test_pred_price <- predict(svm.model_price, d.airbnb_listings_feat_test)
plot(svm.test_pred_price)
svm.conf_matrix_price <- table(svm.test_pred_price, d.airbnb_listings_feat_test[,1]) %>% print()

```







### SVM auf das GAM-DAtaset

```{r generating Split for SVM (TIMON)}
s <- sample(3160,2528) #80% of data is in the test-dataset
coloums <- c(
  "traffic_light",
  "accommodates",
  "longitude") #here we can set the relevant coloums for the svm model out uf the dataset
d.airbnb_listings_feat_gam_train <- d.airbnb_listings_feat_gam[s, coloums]
d.airbnb_listings_feat_gam_test <- d.airbnb_listings_feat_gam[-s, coloums]

```

```{r tuning model for optimal cost}
svm.tune <- tune(svm, as.factor(traffic_light) ~ ., data = d.airbnb_listings_feat_gam_train, kernel = "linear", ranges = list(cost = c(0.001, 0.01, 0.1, 1, 10, 100)))
summary(svm.tune)

```
> summary(svm.tune)

Parameter tuning of ‘svm’:

- sampling method: 10-fold cross validation 

- best parameters:

- best performance: 0.5399555 

- Detailed performance results:



```{r}
svm.model <- svm(as.factor(traffic_light) ~ ., data = d.airbnb_listings_feat_gam_train, kernel = "radial", cost = 0.1, scale = TRUE)

print(svm.model)
#plot(svm.model, data = d.airbnb_listings_feat_gam_test)

svm.test_pred <- predict(svm.model, d.airbnb_listings_feat_gam_test)
plot(svm.test_pred)
svm.conf_matrix <- table(svm.test_pred, d.airbnb_listings_feat_gam_test[,1]) %>% print()
```




# ANN: Hilfe durch AI, sehr kompliziert!

```{r}
d.airbnb_listings_feat_ANN <- d.airbnb_listings_feat %>%
  select(
    price,                  
    review_scores_rating,   
    number_of_reviews,      
    host_is_superhost,      
    accommodates            
  ) %>%
  na.omit() # Für diesen simplen Ansatz werfen wir fehlende Werte einfach raus

# 3. 
# Neuronale Netze funktionieren am besten, wenn alle Werte zwischen 0 und 1 liegen.
normalize <- function(x) {
  return ((x - min(x)) / (max(x) - min(x)))
}
d.airbnb_listings_feat_ANN_scaled <- as.data.frame(lapply(d.airbnb_listings_feat_ANN, normalize))

# 4. Das Neuronale Netz trainieren
# Wir nutzen 1 Hidden Layer mit 3 Neuronen (sehr simpel)
set.seed(42) # Für reproduzierbare Ergebnisse
nn_simple <- neuralnet(
  price ~ review_scores_rating + number_of_reviews + host_is_superhost + accommodates,
  data = d.airbnb_listings_feat_ANN_scaled,
  hidden = 3,
  linear.output = TRUE,
  stepmax = 1e6         
)

# 5. Das Netz visualisieren
plot(nn_simple)


# Durchschnittswerte für unser fiktives Airbnb berechnen
mean_rating <- mean(d.airbnb_listings_feat_ANN_scaled$review_scores_rating)
mean_reviews <- mean(d.airbnb_listings_feat_ANN_scaled$number_of_reviews)
mean_acc <- mean(d.airbnb_listings_feat_ANN_scaled$accommodates)

# Test für H3: Macht Superhost einen Unterschied?
# Wir erfinden zwei Hosts: Exakt gleich, nur einer ist Superhost (1), der andere nicht (0)
test_data_h3 <- data.frame(
  review_scores_rating = c(mean_rating, mean_rating),
  number_of_reviews = c(mean_reviews, mean_reviews),
  host_is_superhost = c(0, 1), # <-- Hier Unterschied!
  accommodates = c(mean_acc, mean_acc)
)

# Das Netz den Preis vorhersagen lassen
predictions_h3 <- compute(nn_simple, test_data_h3)

# Skalierung rückgängig machen, um echte CHF-Preise zu sehen
min_price <- min(d.airbnb_listings_feat_ANN$price)
max_price <- max(d.airbnb_listings_feat_ANN$price)
echte_preise <- predictions_h3$net.result * (max_price - min_price) + min_price

print(echte_preise)
```


```{r}

# --- 1. Datenaufbereitung & Reduzierung ---
d_model_ready <- d.airbnb_listings_feat %>%
  select(
    price,                  
    review_scores_rating,   
    number_of_reviews,      
    host_is_superhost,      
    accommodates,
    property_type,          
    room_type               
  ) %>%
  na.omit() %>%
  mutate(property_type = fct_lump_n(property_type, n = 4, other_level = "Other"))

# --- 2. Dummy-Variablen erstellen (One-Hot-Encoding) ---

d_dummies <- as.data.frame(model.matrix(~ . - 1, data = d_model_ready))

# Leer- oder Sonderzeichen in Spaltennamen bringen neuralnet zum Absturz. 
colnames(d_dummies) <- make.names(colnames(d_dummies))


# --- 3. Skalierung ---
normalize <- function(x) {
  return ((x - min(x)) / (max(x) - min(x)))
}
# Wir wenden normalize auf alle Spalten an. DummyVariabeln bleiben gelich
d_scaled <- as.data.frame(lapply(d_dummies, normalize))


# --- 4. Das Neuronale Netz trainieren ---
set.seed(42) 


nn_simple <- neuralnet(
  price ~ .,
  data = d_scaled,
  hidden = 3,
  linear.output = TRUE, 
  stepmax = 1e6         
)

# 5. Das Netz visualisieren
plot(nn_simple)


# --- 6. Vorhersage & Test für H3 ---
# Um eine faire "Was-wäre-wenn"-Analyse zu machen, nehmen wir einfach von ALLEN 
# Variablen im Datensatz den Durchschnitt (das repräsentiert das "durchschnittliche Airbnb").
test_data_h3 <- as.data.frame(lapply(d_scaled, mean))
# Wir verdoppeln diese Zeile, damit wir zwei identische "durchschnittliche" Hosts haben
test_data_h3 <- rbind(test_data_h3, test_data_h3)

# Jetzt verändern wir NUR den Superhost-Status:
# Zeile 1 = Kein Superhost (0), Zeile 2 = Superhost (1)
test_data_h3$host_is_superhostFALSE <- c(1, 0) 
test_data_h3$host_is_superhostTRUE <- c(0, 1)

# Das Netz den Preis vorhersagen lassen (Wir schließen die Preis-Spalte für den Test aus)
predictions_h3 <- compute(nn_simple, test_data_h3 %>% select(-price))

# Skalierung rückgängig machen mit den Werten aus d_model_ready (Originaldaten!)
min_price <- min(d_model_ready$price)
max_price <- max(d_model_ready$price)

echte_preise <- predictions_h3$net.result * (max_price - min_price) + min_price

print(echte_preise)
```
# Interpretation ANN
Ausgangslage und Hypothesen
Ziel der Untersuchung war es, eine datengetriebene Pricing-Strategie für Airbnb-Hosts im Kanton Waadt (Vaud) zu entwickeln. Im Fokus stand die Frage, welche Faktoren die Zahlungsbereitschaft (Willingness to Pay, WTP) der Gäste beeinflussen und wie Hosts ihre Preise entsprechend anpassen können. Eine zentrale Hypothese (H3) nahm an, dass der "Superhost"-Status eines Vermieters einen signifikanten Preisaufschlag rechtfertigt, da er höheres Vertrauen und eine bessere Servicequalität suggeriert. Als Proxy für die WTP wurden die tatsächlich aufgerufenen Preise der Inserate verwendet.

Modell 1: Das naive Neuronale Netz (Omitted Variable Bias)
In einem ersten, vereinfachten Ansatz wurde ein Neuronales Netz (Multilayer Perceptron) mit nur vier Variablen trainiert: Preis (Zielvariable), Bewertung (Review Score), Anzahl der Bewertungen und Superhost-Status (ja/nein), kontrolliert durch die maximale Gästeanzahl (accommodates).

Ergebnis Modell 1: Entgegen der Erwartungen sagte das Netz für einen Superhost einen niedrigeren Preis (ca. 132 CHF) voraus als für einen regulären Host (ca. 160 CHF).

Interpretation: Dieses Ergebnis ist ein klassischer Fall von "Omitted Variable Bias" (Verzerrung durch weggelassene Variablen). Superhosts erzielen ihren Status oft durch hohe Auslastung (viele Buchungen), was in der Praxis häufig über wettbewerbsfähige, tendenziell niedrigere Preise erreicht wird. Zudem fehlen in diesem Modell entscheidende Kontextinformationen wie die Art der Unterkunft: Luxuriöse, sehr teure Objekte werden oft von professionellen Agenturen verwaltet, die seltener über einen individuellen Superhost-Badge verfügen.

Modell 2: Das erweiterte Modell (One-Hot-Encoding)
Um diese Verzerrung zu korrigieren, wurde das Modell im zweiten Schritt um kategoriale Variablen erweitert: die Art der Unterkunft (room_type, z.B. Privatzimmer vs. ganze Wohnung) und den Immobilientyp (property_type, z.B. Apartment vs. Chalet). Da Neuronale Netze nur numerische Inputs verarbeiten können, wurden diese Kategorien mittels "One-Hot-Encoding" in binäre Dummy-Variablen übersetzt.

Ergebnis Modell 2: Beim Test eines fiktiven, absolut durchschnittlichen Inserats (gleicher Raumtyp, gleiche Kapazität etc.) unterschied sich der Preis nur noch marginal: 483,29 CHF (kein Superhost) zu 482,64 CHF (Superhost).

Interpretation: Wenn wir durch Dummy-Variablen sicherstellen, dass wir "Äpfel mit Äpfeln" vergleichen, verschwindet der vermeintliche Preisabschlag für Superhosts fast vollständig. Gleichzeitig wird unsere Ausgangshypothese (H3) widerlegt: Der reine Superhost-Status führt nicht zu einer signifikant höheren Zahlungsbereitschaft der Gäste.

Fazit für den Business Case
Für das Decision-Support-System bedeutet dies, dass Pricing-Empfehlungen primär auf strukturellen Faktoren (Größe, Lage, Art der Immobilie) basieren müssen. Sich den Superhost-Status zu erarbeiten, ist für Hosts aus betriebswirtschaftlicher Sicht zwar ratsam, um die Sichtbarkeit und Buchungswahrscheinlichkeit (Occupancy) zu erhöhen, rechtfertigt aber per se keinen spürbaren Preisaufschlag im Markt.
